### 2024-10-19 paper/paper.qmd
update paper

---
title: "Prediction of US election with linear model"
author: 
  - Colin Sihan Yang
  - Lexun Yu
  - Siddharth Gowda
thanks: "Code and data are available at: [https://github.com/yulexun/uselection](https://github.com/yulexun/uselection)."
date: today
date-format: long
abstract: "This paper forecast the winner of the upcoming US presidential election using “poll-of-polls” by building a linear model."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction {#sec-intro}

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-19 scripts/02-download_data.R
update download r

@@ -9,18 +9,12 @@
 
 
 #### Workspace setup ####
-library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# Specify the download URL from FiveThirtyEight
+url <- "https://projects.fivethirtyeight.com/polls/data/president_polls.csv"
 
+# Save the data csv
+download.file(url, "data/01-raw_data/president_polls.csv")
 
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         

### 2024-10-20 scripts/02-download_data.R
update download data and omit NA

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from fivethirtyeight.com.
+# Author: Lexun Yu
+# Date: 19 October 2024
+# Contact: lx.yu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 
 #### Workspace setup ####

### 2024-10-20 scripts/03-clean_data.R

@@ -11,34 +11,29 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/01-raw_data/president_polls.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+# Delete NA columns, 
+# Note: I am not sure if I deleting state here is appropriate, the half of the state column are missing.
+cleaned_data <- raw_data |>
+  select(-sponsor_ids, 
+    -sponsors, 
+    -sponsor_candidate_id, 
+    -sponsor_candidate, 
+    -sponsor_candidate_party, 
+    -state, 
+    -endorsed_candidate_id, 
+    -endorsed_candidate_name, 
+    -endorsed_candidate_party, 
+    -subpopulation, -tracking, 
+    -notes, -url_article, -url_topline, -url_crosstab,
+    -source, -internal, -partisan, -seat_name, 
+    -ranked_choice_round) |> na.omit()
+
+# Keep data from one pollster for Appendix
+morningconsult_data <- cleaned_data %>%
+  filter(pollster == "Morning Consult")
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")
+write_csv(morningconsult_data, "data/02-analysis_data/mc_data.csv")

### 2024-10-21 scripts/03-clean_data.R
filter data by national or state

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Purpose: Cleans the raw  data from FiveThirtyEight
 # Author: Rohan Alexander [...UPDATE THIS...]
 # Date: 6 April 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
@@ -13,27 +13,40 @@ library(tidyverse)
 #### Clean data ####
 raw_data <- read_csv("data/01-raw_data/president_polls.csv")
 
-# Delete NA columns, 
-# Note: I am not sure if I deleting state here is appropriate, the half of the state column are missing.
+
 cleaned_data <- raw_data |>
-  select(-sponsor_ids, 
-    -sponsors, 
-    -sponsor_candidate_id, 
-    -sponsor_candidate, 
-    -sponsor_candidate_party, 
-    -state, 
-    -endorsed_candidate_id, 
-    -endorsed_candidate_name, 
-    -endorsed_candidate_party, 
-    -subpopulation, -tracking, 
-    -notes, -url_article, -url_topline, -url_crosstab,
-    -source, -internal, -partisan, -seat_name, 
-    -ranked_choice_round) |> na.omit()
+  filter(
+    numeric_grade >= 2.7
+  )
+
+cleaned_data_national <- cleaned_data |>
+  filter(is.na(state))
+
+cleaned_data_state <- cleaned_data |>
+  filter(!is.na(state))
+
+# Delete NA columns
+# cleaned_data <- raw_data |>
+#   select(-sponsor_ids, 
+#     -sponsors, 
+#     -sponsor_candidate_id, 
+#     -sponsor_candidate, 
+#     -sponsor_candidate_party, 
+#     -endorsed_candidate_id, 
+#     -endorsed_candidate_name, 
+#     -endorsed_candidate_party, 
+#     -subpopulation, -tracking, 
+#     -notes, -url_article, -url_topline, -url_crosstab,
+#     -source, -internal, -partisan, -seat_name, 
+#     -ranked_choice_round) |> na.omit()
+
 
 # Keep data from one pollster for Appendix
-morningconsult_data <- cleaned_data %>%
-  filter(pollster == "Morning Consult")
+# morningconsult_data <- cleaned_data %>%
+#   filter(pollster == "Morning Consult")
 
 #### Save data ####
[…]

### 2024-10-21 paper/paper.qmd
update introduction

@@
 # Introduction {#sec-intro}
+Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States.
+Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result.
+As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
+However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
+This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+
+This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
+We incorporate aggregation by filtering the polls by numeric grade that indicates pollster’s reliability, prediction that incorporates historical data and economic indicators, and hybrid approaches that leverages Bayesian techniques, allowing for a dynamic prediction of the U.S.
+presidential election.
+
+The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump.
+The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.
+
+The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]
+
+The remainder of this paper is structured as follows: [update this …]
 
@@
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
 Talk way more about it.
@@
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+us
 
@@
 Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
 
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-10-21 paper/paper.qmd
add literature review to introduction

@@
 This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
+As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
+Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics.
+Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
 We incorporate aggregation by filtering the polls by numeric grade that indicates pollster’s reliability, prediction that incorporates historical data and economic indicators, and hybrid approaches that leverages Bayesian techniques, allowing for a dynamic prediction of the U.S.

### 2024-10-21 paper/paper.qmd
update chean data, 538 cc license

@@
 Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
-We incorporate aggregation by filtering the polls by numeric grade that indicates pollster’s reliability, prediction that incorporates historical data and economic indicators, and hybrid approaches that leverages Bayesian techniques, allowing for a dynamic prediction of the U.S.
+We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
 presidential election.
@@
   coord_flip()
+
 ```
@@
 
+# FiveThirtyEight Licenses
+[FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).
+
 \newpage

### 2024-10-21 scripts/03-clean_data.R

@@ -1,17 +1,21 @@
 #### Preamble ####
 # Purpose: Cleans the raw  data from FiveThirtyEight
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Colin Sihan Yang, Siddharth Gowda, Lexun Yu
+# Date: 21 October 2024
+# Contact: lx.yu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
 raw_data <- read_csv("data/01-raw_data/president_polls.csv")
+head(raw_data %>% mutate(
+  state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+  end_date = mdy(end_date),
+  start_date = mdy(start_date)
+) %>% select(state))
 
 
 cleaned_data <- raw_data |>
@@ -19,25 +23,36 @@ cleaned_data <- raw_data |>
     numeric_grade >= 2.7
   )
 
-cleaned_data_national <- cleaned_data |>
-  filter(is.na(state))
+cleaned_data <- cleaned_data |>
+  mutate(
+    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    end_date = mdy(end_date),
+    start_date = mdy(start_date),
+    election_date = mdy(election_date)
+  ) |>
+  filter(start_date >= as.Date("2024-07-21")) # When Harris declared
 
-cleaned_data_state <- cleaned_data |>
-  filter(!is.na(state))
+# Election date is 5 Nov 2024
+cleaned_data <- cleaned_data %>%
+  filter(cycle == 2024, office_type == "U.S. President", stage == "general") %>%
+  mutate(days_taken_from_election = as.numeric(mdy("11/5/2024") - start_date)) %>%
+  select(poll_id, pollster_id, pollster, question_id,
+         sample_size, pollscore, methodology, days_taken_from_election,
+         end_date, start_date, state, answer, pct)
 
 # Delete NA columns
 # cleaned_data <- raw_data |>
-#   select(-sponsor_ids, 
-#     -sponsors, 
-#     -sponsor_candidate_id, 
[…]

### 2024-10-21 scripts/02-download_data.R
update style and simulate data

@@ -16,4 +16,3 @@ url <- "https://projects.fivethirtyeight.com/polls/data/president_polls.csv"
 
 # Save the data csv
 download.file(url, "data/01-raw_data/president_polls.csv")
-

### 2024-10-21 scripts/03-clean_data.R

@@ -11,11 +11,11 @@ library(tidyverse)
 
 #### Clean data ####
 raw_data <- read_csv("data/01-raw_data/president_polls.csv")
-head(raw_data %>% mutate(
-  state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
-  end_date = mdy(end_date),
-  start_date = mdy(start_date)
-) %>% select(state))
+# head(raw_data %>% mutate(
+#   state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+#   end_date = mdy(end_date),
+#   start_date = mdy(start_date)
+# ) %>% select(state))
 
 
 cleaned_data <- raw_data |>
@@ -36,9 +36,11 @@ cleaned_data <- cleaned_data |>
 cleaned_data <- cleaned_data %>%
   filter(cycle == 2024, office_type == "U.S. President", stage == "general") %>%
   mutate(days_taken_from_election = as.numeric(mdy("11/5/2024") - start_date)) %>%
-  select(poll_id, pollster_id, pollster, question_id,
-         sample_size, pollscore, methodology, days_taken_from_election,
-         end_date, start_date, state, answer, pct)
+  select(
+    poll_id, pollster_id, pollster, question_id,
+    sample_size, pollscore, methodology, days_taken_from_election,
+    end_date, start_date, state, answer, pct
+  )
 
 # Delete NA columns
 # cleaned_data <- raw_data |>
@@ -62,6 +64,6 @@ cleaned_data <- cleaned_data %>%
 
 #### Save data ####
 write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")
-#write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_national.csv")
-#write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_state.csv")
+# write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_national.csv")
+# write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_state.csv")
 # write_csv(morningconsult_data, "data/02-analysis_data/mc_data.csv")

### 2024-10-21 scripts/00-simulate_data.R
modified:   scripts/00-simulate_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the
-# state and party that won each division.
+# Purpose: Simulates a dataset of Australian electoral divisions, including the 
+  #state and party that won each division.
 # Author: Rohan Alexander
 # Date: 26 September 2024
 # Contact: rohan.alexander@utoronto.ca
@@ -17,22 +17,69 @@ set.seed(853)
 #### Simulate data ####
 # State names
 states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+  "National",
+  "Colorado",
+  "Oklahoma",
+  "Wisconsin",
+  "Washington",
+  "Oregon",
+  "Florida",
+  "Arizona",
+  "Georgia",
+  "Michigan",
+  "Nevada",
+  "North Carolina",
+  "Pennsylvania",
+  "South Carolina",
+  "Texas",
+  "Tennessee",
+  "Massachusetts",
+  "Maryland",
+  "Ohio",
+  "California",
+  "Minnesota",
+  "New Mexico",
+  "New Hampshire",
+  "Missouri",
+  "Alaska",
+  "Montana",
+  "Nebraska",
+  "Utah",
+  "Illinois",
+  "Virginia",
+  "New Jersey",
+  "North Dakota",
+  "New York",
+  "Indiana",
+  "Iowa",
+  "Nebraska CD-2",
+  "Delaware",
+  "Rhode Island",
+  "Connecticut",
+  "Maine",
+  "Maine CD-1",
+  "Maine CD-2",
+  "Arkansas",
+  "West Virginia",
+  "Vermont",
+  "South Dakota",
+  "Louisiana",
+  "Hawaii",
+  "Alabama",
+  "Idaho",
+  "Kansas",
+  "Kentucky",
+  "Mississippi",
+  "Wyoming"
 )
+
 
 # Political parties
 parties <- c("Labor", "Liberal", "Greens", "National", "Other")
 
 # Create a dataset by randomly assigning states and parties to divisions
 analysis_data <- tibble(
-  division = paste("Division", 1:151), # Add "Division" to make it a character
+  division = paste("Division", 1:151),  # Add "Division" to make it a character
   state = sample(
     states,
     size = 151,

### 2024-10-21 scripts/00-simulate_data.R
update simulated data

@@ -16,82 +16,20 @@ set.seed(853)
 
 #### Simulate data ####
 # State names
-states <- c(
-  "National",
-  "Colorado",
-  "Oklahoma",
-  "Wisconsin",
-  "Washington",
-  "Oregon",
-  "Florida",
-  "Arizona",
-  "Georgia",
-  "Michigan",
-  "Nevada",
-  "North Carolina",
-  "Pennsylvania",
-  "South Carolina",
-  "Texas",
-  "Tennessee",
-  "Massachusetts",
-  "Maryland",
-  "Ohio",
-  "California",
-  "Minnesota",
-  "New Mexico",
-  "New Hampshire",
-  "Missouri",
-  "Alaska",
-  "Montana",
-  "Nebraska",
-  "Utah",
-  "Illinois",
-  "Virginia",
-  "New Jersey",
-  "North Dakota",
-  "New York",
-  "Indiana",
-  "Iowa",
-  "Nebraska CD-2",
-  "Delaware",
-  "Rhode Island",
-  "Connecticut",
-  "Maine",
-  "Maine CD-1",
-  "Maine CD-2",
-  "Arkansas",
-  "West Virginia",
-  "Vermont",
-  "South Dakota",
-  "Louisiana",
-  "Hawaii",
-  "Alabama",
-  "Idaho",
-  "Kansas",
-  "Kentucky",
-  "Mississippi",
-  "Wyoming"
-)
+states <- state.name
+answer <- c("Trump", "Harris")
+pollster <- c("Marist", "YouGov", "Ipsos", "Emerson", "SurveyUSA")
 
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
 
 # Create a dataset by randomly assigning states and parties to divisions
 analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
   state = sample(
     states,
-    size = 151,
+    size = 1500,
     replace = TRUE,
-    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
   ),
-  party = sample(
-    parties,
-    size = 151,
-    replace = TRUE,
-    prob = c(0.40, 0.40, 0.05, 0.1, 0.05) # Rough party distribution
-  )
+
 )
 
 

### 2024-10-21 paper/paper.qmd
update data and measurement

@@
 library(palmerpenguins)
+library(here)
+library(tibble)
+library(knitr)
 ```
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
+- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
+- **numeric predictor:**\
+ sample size(sample_size: The total number of respondents participating in the poll) \
+ timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
+ pollscore(A numeric value representing the score or reliability of the pollster in question) \
+- **categorical predictor**
+state(The U.S. state where the poll was conducted or focused) \
+methodology(The method used to conduct the poll)\
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+In this dataset, each row represents a polling question that records the variables of interest.
+Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
+This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
 
-## Outcome variables
+## Clean Data
+The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
+First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable.
+Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S.
+election date.
+Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy.
+Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-cleaned-data
+#| tbl-cap: Sample of cleaned US election data 
 #| echo: false
+#| warning: false
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-
+cleaned_data |>
+  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, answer) |>
+  head(6) |>
+  kable(
+    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "answer"),
+    booktabs = TRUE
+    ) 
 ```
 
-Talk more about it.
+## Basic Statistics Summary for Data
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+```{r}
+#| label: fig-clean-data-histogram-Harris-Trump
+#| fig-cap: the average PCT vs State for Harris and Trump
+#| echo: false
+#| warning: false
+#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
+#| layout-nrow: 2
+#| fig-width: 40
+#| fig-height: 20
+Harris_data = cleaned_data[cleaned_data$answer == "Harris", ]
+Harris_data |>
+  group_by(state) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  ggplot(mapping = aes(x = state, y = avg_pct)) + 
+  geom_col()  +
+  theme_minimal() +
+  labs(
+    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
+  )
 
-Talk way more about it.
+Trump_data = cleaned_data[cleaned_data$answer == "Trump", ]
+Trump_data |>
+  group_by(state) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  ggplot(mapping = aes(x = state, y = avg_pct)) + 
+  geom_col()  +
+  theme_minimal() +
+  labs(
+    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
+  )
 
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+# cleaned_data %>%
+#   mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
+#                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
+#   ggplot(mapping = aes(x = month, y = total_decedents)) + 
+#   geom_bar(stat = "identity") +
+#   theme_minimal() +
+#   labs(
+#     x = "Month", 
+#     y = "Total Decedents"
+#   )
+```
 

### 2024-10-21 scripts/03-clean_data.R

@@ -20,7 +20,7 @@ raw_data <- read_csv("data/01-raw_data/president_polls.csv")
 
 cleaned_data <- raw_data |>
   filter(
-    numeric_grade >= 2.7
+    numeric_grade >= 2.5
   )
 
 cleaned_data <- cleaned_data |>

### 2024-10-21 scripts/03-clean_data.R
forced data to be after harris declaration and to be only general election

@@ -12,32 +12,45 @@ library(tidyverse)
 
 #### Clean data ####
 raw_data <- read_csv("data/01-raw_data/president_polls.csv")
-
-
 cleaned_data <- raw_data |>
   filter(
-    numeric_grade >= 2.7
+    numeric_grade >= 2.5
   )
 
-cleaned_data_national <- cleaned_data |>
-  filter(is.na(state))
 
-cleaned_data_state <- cleaned_data |>
-  filter(!is.na(state))
+cleaned_data <- cleaned_data |>
+  mutate(
+    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    end_date = mdy(end_date),
+    start_date = mdy(start_date),
+    election_date = mdy(election_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) # When Harris declared
+
+# Election date is 5 Nov 2024
+cleaned_data <- cleaned_data %>%
+  filter(cycle == 2024, office_type == "U.S. President", stage == "general") %>%
+  mutate(days_taken_from_election = as.numeric(mdy("11/5/2024") - start_date)) %>%
+  select(poll_id, pollster_id, pollster, question_id,
+         sample_size, pollscore, methodology,
+         days_taken_from_election, transparency_score, party,
+         end_date, start_date, state, answer, pct)
+
+cleaned_data %>% select(state) %>% unique() %>% data.frame()
 
 # Delete NA columns
 # cleaned_data <- raw_data |>
-#   select(-sponsor_ids, 
-#     -sponsors, 
-#     -sponsor_candidate_id, 
-#     -sponsor_candidate, 
-#     -sponsor_candidate_party, 
-#     -endorsed_candidate_id, 
-#     -endorsed_candidate_name, 
-#     -endorsed_candidate_party, 
-#     -subpopulation, -tracking, 
+#   select(-sponsor_ids,
+#     -sponsors,
+#     -sponsor_candidate_id,
+#     -sponsor_candidate,
+#     -sponsor_candidate_party,
+#     -endorsed_candidate_id,
+#     -endorsed_candidate_name,
+#     -endorsed_candidate_party,
+#     -subpopulation, -tracking,
 #     -notes, -url_article, -url_topline, -url_crosstab,
-#     -source, -internal, -partisan, -seat_name, 
+#     -source, -internal, -partisan, -seat_name,
[…]

### 2024-10-21 paper/paper.qmd
made first multi linear model with residuals

--- 2024-10-21 paper/paper.qmd
@@
 library(tidyverse)
-library(palmerpenguins)
+library(modelsummary)
 ```
@@
 # Introduction {#sec-intro}
-Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States.
-Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result.
-As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
-However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
-This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
-
-This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
-As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
-Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics.
-Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
-We incorporate aggregation by filtering the polls by numeric grade that indicates pollster’s reliability, prediction that incorporates historical data and economic indicators, and hybrid approaches that leverages Bayesian techniques, allowing for a dynamic prediction of the U.S.
-presidential election.
-
-The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump.
-The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.
-
-The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]
-
-The remainder of this paper is structured as follows: [update this …]
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
-
-## Measurement
-
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
-
-## Outcome variables
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Model
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+#| include: false
+#| warning: false
+#| message: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# ELECTION DATE: 11 Nov 24
+# Day Kamala announced she was running: 21 July 2024
+# will start looking at polling data from August
+
+# data manipulation
+clean_data <- read_csv("../data/02-analysis_data/cleaned_data.csv") %>% 
+  filter(answer %in% c("Harris", "Trump", "Kennedy", "Stein"))
 ```
 
-Talk more about it.
+```{r}
+#| label: tbl-multilinearmodel
+#| tbl-cap: "Coefficents for Multilinear Regression Model"
+#| echo: false
+#| warning: false
+#| message: false
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+multi_lm <- lm(pct ~ transparency_score + pollscore + sample_size + answer + days_taken_from_election, clean_data)
 
-Talk way more about it.
+modelsummary(multi_lm)
+```
 
-## Predictor variables
+\begin{table}[htbp]
+\centering
+\caption{Coefficients for Multilinear Regression Model}
+\label{tbl-multilinearmodel}
+\begin{tabular}{lc}
+\hline
+Variable & Coefficient \\
+\hline
+Intercept & 51.371 \\
+Transparency Score & -0.218 \\
+Poll Score & 0.321 \\
+Sample Size & 0.000 \\
+Candidate Name & Varies \\
+Days from Election & Not shown \\
+\hline
+\end{tabular}
+\end{table}
 
-Add graphs, tables and text.
+\begin{equation}
+\begin{split}
+\text{candidate\_percentage} = & -0.218 \times \text{transparency\_score} + 0.321 \times \text{pollscore} \\
+& + 0 \times \text{sample\_size} + \text{candidate\_intercept} \times \text{candidate\_name\_vector} \\
+& + 51.371
+\end{split}
+\end{equation}
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+From @tbl-multilinearmodel, candiate name is the biggest factor in terms of support.
+The name was added to the model to give a baseline not to show any correlation.
+After that, poll score and transparency score are the biggest factors.
+These scores are two baselines of how reliable and trustable the pollsters are.
+The sample size had a coefficent of zero, meaning it had no affect on the model's output.
 
-# Model
+The model also has a very high R^2 value in 0.962 Most of this can be attributed to using the canidate name (anwser) as a variable.
 
-The goal of our modelling strategy is twofold. Firstly,...
+## Residuals
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+```{r}
+#| label: fig-multilinearmodel
+#| fig-cap: "Multilinear Regression Model Residuals"
+#| echo: false
+#| warning: false
+#| message: false
 
-## Model set-up
+ggplot(multi_lm, aes(x = .fitted, y = .resid, color = answer)) +
+  geom_point() +
+  geom_hline(yintercept = 0) +
+  labs(y = "Residual", x = "Predicted Candiate Vote Percentage")
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-us
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+```
+Based on @fig-multilinearmodel, there appears to be no patterns in the residuals.
+This means that a linear model like this could be appropriate.
+However, there does seem to be a big difference in terms of residual variability based on candidate name.
 

### 2024-10-22 paper/paper.qmd
update simulate data

--- 2024-10-21 paper/paper.qmd
@@
 ---
-title: "Prediction of US election with linear model"
+title: "Prediction of 2024 US election ..."
 author: 
@@
 date-format: long
-abstract: "This paper forecast the winner of the upcoming US presidential election using “poll-of-polls” by building a linear model."
+abstract: "We forecast the winner of the 2024 US presidential election using “poll-of-polls” by building a linear model."
 format: pdf

### 2024-10-22 scripts/00-simulate_data.R

@@ -17,19 +17,44 @@ set.seed(853)
 #### Simulate data ####
 # State names
 states <- state.name
-answer <- c("Trump", "Harris")
+answer <- c("Trump", "Harris", "Oliver", "West")
 pollster <- c("Marist", "YouGov", "Ipsos", "Emerson", "SurveyUSA")
 
+# Generate Poll IDs and States
+unique_poll_ids <- sample(10000:99999, 50)
+poll_ids <- rep(unique_poll_ids, each = 4)
+state_column <- rep(states, each = 4)
+state_column <- rep(state_column, length.out = 200)
+sample_size <- sample(100:20000, 50)
+sample_size <- rep(sample_size, each = 4)
+pollster_column <- sample(pollster, 50, replace = TRUE)
+pollster_column <- rep(pollster_column, each = 4)
 
+# Function to generate a set of 4 random percentages that sum to 100%
+generate_percentages <- function() {
+  # Generate 4 random numbers
+  rand_values <- runif(4)
+  # Normalize them so they sum to 100
+  percentages <- (rand_values / sum(rand_values)) * 100
+  return(percentages)
+}
+
+# Initialize an empty vector to store percentages
+pct_column <- c()
+
+# Generate percentages for 200 rows (50 groups of 4 rows)
+for (i in 1:50) {
+  pct_column <- c(pct_column, generate_percentages())
+}
 
 # Create a dataset by randomly assigning states and parties to divisions
 analysis_data <- tibble(
-  state = sample(
-    states,
-    size = 1500,
-    replace = TRUE,
-  ),
-
+  pollster = pollster_column,
+  poll_ids = poll_ids,
+  state = state_column,
+  sample_size = sample_size,
+  answer = rep(answer, length.out = 200),
+  pct = pct_column,
 )
 
 

### 2024-10-22 scripts/00-simulate_data.R
fix style

@@ -1,6 +1,6 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
+# Purpose: Simulates a dataset of Australian electoral divisions, including the
+# state and party that won each division.
 # Author: Rohan Alexander
 # Date: 26 September 2024
 # Contact: rohan.alexander@utoronto.ca

### 2024-10-22 scripts/00-simulate_data.R
update simulation style

@@ -1,12 +1,11 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the
-# state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of Australian electoral divisions, including the 
+  #state and party that won each division.
+# Author: Lexun Yu
+# Date: 22 October 2024
+# Contact: lx.yu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
 
 
 #### Workspace setup ####

### 2024-10-30 scripts/03-clean_data.R
update cleaned data and youguv methodology

@@ -23,6 +23,10 @@ cleaned_data <- raw_data |>
     numeric_grade >= 2.5
   )
 
+# Keep data from one pollster for Appendix
+morningconsult_data <- cleaned_data %>%
+  filter(pollster == "Morning Consult")
+
 cleaned_data <- cleaned_data |>
   mutate(
     state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
@@ -42,6 +46,11 @@ cleaned_data <- cleaned_data %>%
     end_date, start_date, state, answer, pct
   )
 
+  cleaned_data_national <- cleaned_data |>
+    filter(state == "National")
+  cleaned_data_state <- cleaned_data |>
+    filter(state != "National")
+
 # Delete NA columns
 # cleaned_data <- raw_data |>
 #   select(-sponsor_ids,
@@ -57,13 +66,8 @@ cleaned_data <- cleaned_data %>%
 #     -source, -internal, -partisan, -seat_name,
 #     -ranked_choice_round) |> na.omit()
 
-
-# Keep data from one pollster for Appendix
-# morningconsult_data <- cleaned_data %>%
-#   filter(pollster == "Morning Consult")
-
 #### Save data ####
 write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")
-# write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_national.csv")
-# write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_state.csv")
-# write_csv(morningconsult_data, "data/02-analysis_data/mc_data.csv")
+write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_national.csv")
+write_csv(cleaned_data_state, "data/02-analysis_data/cleaned_data_state.csv")
+write_csv(morningconsult_data, "data/02-analysis_data/mc_data.csv")

### 2024-10-30 scripts/03-clean_data.R
add swing states article

@@ -51,6 +51,8 @@ cleaned_data <- cleaned_data %>%
   cleaned_data_state <- cleaned_data |>
     filter(state != "National")
 
+
+
 # Delete NA columns
 # cleaned_data <- raw_data |>
 #   select(-sponsor_ids,

### 2024-10-30 scripts/05-exploratory_data_analysis.R

@@ -15,6 +15,8 @@ library(rstanarm)
 #### Read data ####
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
+# Plot poll score and sample size
+
 ### Model data ####
 first_model <-
   stan_glm(

### 2024-10-30 scripts/03-clean_data.R
add weight based on pollscore

@@ -43,15 +43,21 @@ cleaned_data <- cleaned_data %>%
   select(
     poll_id, pollster_id, pollster, question_id,
     sample_size, pollscore, methodology, days_taken_from_election,
-    end_date, start_date, state, answer, pct
+    end_date, start_date, state, candidate_name, pct
   )
 
+# add weight to pct
+cleaned_data$adjusted_weight <- 1 / abs(cleaned_data$pollscore)
+
+cleaned_data |>
+  mutate(weight = sum(pct * adjusted_weight) / sum(adjusted_weight))
+
+# store national and state data seperately
   cleaned_data_national <- cleaned_data |>
     filter(state == "National")
   cleaned_data_state <- cleaned_data |>
     filter(state != "National")
-
-
+  
 
 # Delete NA columns
 # cleaned_data <- raw_data |>

### 2024-10-30 scripts/03-clean_data.R
update clean data

@@ -49,15 +49,13 @@ cleaned_data <- cleaned_data %>%
 # add weight to pct
 cleaned_data$adjusted_weight <- 1 / abs(cleaned_data$pollscore)
 
-cleaned_data |>
-  mutate(weight = sum(pct * adjusted_weight) / sum(adjusted_weight))
 
 # store national and state data seperately
   cleaned_data_national <- cleaned_data |>
     filter(state == "National")
   cleaned_data_state <- cleaned_data |>
     filter(state != "National")
-  
+
 
 # Delete NA columns
 # cleaned_data <- raw_data |>

### 2024-10-30 scripts/03-clean_data.R
update file format to parquet

@@ -8,6 +8,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
 #### Clean data ####
 raw_data <- read_csv("data/01-raw_data/president_polls.csv")
@@ -73,7 +74,7 @@ cleaned_data$adjusted_weight <- 1 / abs(cleaned_data$pollscore)
 #     -ranked_choice_round) |> na.omit()
 
 #### Save data ####
-write_csv(cleaned_data, "data/02-analysis_data/cleaned_data.csv")
-write_csv(cleaned_data_national, "data/02-analysis_data/cleaned_data_national.csv")
-write_csv(cleaned_data_state, "data/02-analysis_data/cleaned_data_state.csv")
-write_csv(morningconsult_data, "data/02-analysis_data/mc_data.csv")
+write_parquet(cleaned_data, "data/02-analysis_data/cleaned_data.parquet")
+write_parquet(cleaned_data_national, "data/02-analysis_data/cleaned_data_national.parquet")
+write_parquet(cleaned_data_state, "data/02-analysis_data/cleaned_data_state.parquet")
+write_parquet(morningconsult_data, "data/02-analysis_data/mc_data.parquet")

### 2024-10-30 paper/paper.qmd
colin code update2

@@
 library(knitr)
+library(brms)
+library(rstanarm)
 ```
@@
 Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
+
+In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model.
 We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
@@
 methodology(The method used to conduct the poll)\
+
+## Explore the data
 
@@
 cleaned_data |>
-  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, answer) |>
+  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
   head(6) |>
   kable(
-    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "answer"),
+    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "candidate_name"),
     booktabs = TRUE
@@
 #| fig-height: 20
-Harris_data = cleaned_data[cleaned_data$answer == "Harris", ]
+Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
 Harris_data |>
@@
 
-Trump_data = cleaned_data[cleaned_data$answer == "Trump", ]
+Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
 Trump_data |>
@@
 us
+
+## Basic Model
+```{r}
+just_harris_data = Harris_data |> na.omit()
+model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology, data = just_harris_data)
+summary(model_MLR)
+just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
+
+predictions = predict(model_MLR, just_harris_data)
+
+ggplot(just_harris_data, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
+  facet_wrap(vars(methodology)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
+```
+
+```{r}
+baye_model_data = just_harris_data
+baye_model_data$pct = as.factor(baye_model_data$pct)
+baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
+baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
+baye_model_data$methodology = as.factor(baye_model_data$methodology)
+baye_model_data$pct <- as.numeric(baye_model_data$pct)
+# Define the Bayesian model with brms
+
+formula <- pct ~ pollscore + days_taken_from_election + (1 | methodology) + (1 | state)
+
+priors = normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = formula,
+  data = just_harris_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+pp_check(bayesian_model_1)
+summary(bayesian_model_1)
+
+# Plot random effects
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+
+```
+
+```{r}
+
+```
+
+```{r}
+
+# Transform Biden and Trump’s vote shares to fit a beta distribution
+# Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
+total_votes <- 100
+biden_shape1 <- 51.3 / total_votes * 10
+biden_shape2 <- (1 - 51.3 / total_votes) * 10
+trump_shape1 <- 46.8 / total_votes * 10
+trump_shape2 <- (1 - 46.8 / total_votes) * 10
+
+# Set up priors using beta distributions based on 2020 vote shares
+prior <- c(
+  set_prior(paste0("beta(", biden_shape1, ", ", biden_shape2, ")"), class = "b", coef = "biden"),
+  set_prior(paste0("beta(", trump_shape1, ", ", trump_shape2, ")"), class = "b", coef = "trump")
+)
+
+# Define the model formula
+formula <- pct ~ pollscore + days_taken_from_election + methodology
+
+# Run the Bayesian model
+model <- brm(
+  formula = formula,
+  data = baye_model_data,
+  family = categ0orical(),
+  prior = prior,
+  chains = 4,
+  iter = 2000,
+  warmup = 1000,
+  cores =4
+)
+
+# Summarize the model
+summary(model)
+
+```
 

### 2024-10-31 paper/paper.qmd
update methodology

--- 2024-10-22 paper/paper.qmd
@@
 
+# Methodology of YouGov
+YouGov's methodology documentations are seperated in two articles.
+The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
+
+## Population, Frame, and Sample
+As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
+The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
+Voter files indicates whether asomeone voted in a given election, thus YouGov's population covers all voters in previous US elections.
+
+YouGov’s sampling frame consists of its online panel members.
+These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
+YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
+
+YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
+YouGov interviews nearly 100,000 people in the first set of estimates.
+For the second set of estimates, YouGov didn't just start over with a new sample.
+They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
+
+## Sample Recruitment
+Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
+They must provide demographic details upon joining, which helps in selecting representative samples for each survey.
+When respondents complete a survey, they are awarded points that can be exchanged for money.
+
+## Sampling Approach and Trade-offs
+YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
+This method allows quick and cost-effective data collection.
+However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
+YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
+
+## Non-response Handling
+YouGov apply statistical weighting to adjust for the differences between the sample and target population.
+The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
+Additionally, quality control measures exclude unreliable responses to improve data accuracy.
+The respondents are offered a small incentive to decrease the non-response and increase participation.
+
+## Strengths and Weaknesses of the Questionnaire
+YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
+The pollster can recruit a large amount of panelists because of the online format.
+Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
+
+As a non-probability sample, it might miss certain demographic groups not covered by the online population.
+While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
+Additionaly, the categories in the surbey is oversimplified with bias.
+For instance, in the poll result published by YouGov, gender is divided into Male and Female.
+Race is divided into White, Black, Hispanic and Other.
+This indicates a lack of representation.
+
 # FiveThirtyEight Licenses

### 2024-11-01 paper/paper.qmd
update paper

@@
 library(knitr)
+library(arrow)
+library(rstanarm)
+library(brms)
 ```
@@
 However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
-This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+A more adaptable model is required to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
 
-This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
+We build a hybrid election forecasting model following the strategies mentioned by @pasek.
 As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
@@
 
-In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model.
-We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
+In this paper, we predict the 2024 us election result with the hybrid election forcasting model.
+We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that uses Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
 presidential election.
@@
 
+## Explore the data
+
 ## Measurement
@@
 #| warning: false
-cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
+cleaned_data = read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
 
@@
 #   )
+```
+
+# Model
+
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+us
+
+## Basic Model
+```{r}
+just_harris_data = Harris_data |> na.omit()
+model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology, data = just_harris_data)
+summary(model_MLR)
+just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
+
+predictions = predict(model_MLR, just_harris_data)
+
+ggplot(just_harris_data, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
+  facet_wrap(vars(methodology)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
+```
+
+```{r}
+baye_model_data = just_harris_data
+baye_model_data$pct = as.factor(baye_model_data$pct)
+baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
+baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
+baye_model_data$methodology = as.factor(baye_model_data$methodology)
+baye_model_data$pct <- as.numeric(baye_model_data$pct)
+# Define the Bayesian model with brms
+
+formula <- pct ~ pollscore + days_taken_from_election + (1 | methodology) + (1 | state)
+
+priors = normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = formula,
+  data = just_harris_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+pp_check(bayesian_model_1)
+summary(bayesian_model_1)
+
+# Plot random effects
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+
+```
+
+```{r}
+
+# Transform Biden and Trump’s vote shares to fit a beta distribution
+# Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
+total_votes <- 100
+biden_shape1 <- 51.3 / total_votes * 10
+biden_shape2 <- (1 - 51.3 / total_votes) * 10
+trump_shape1 <- 46.8 / total_votes * 10
+trump_shape2 <- (1 - 46.8 / total_votes) * 10
+
+# Set up priors using beta distributions based on 2020 vote shares
+# prior <- c(
+#   set_prior(paste0("beta(", biden_shape1, ", ", biden_shape2, ")"), class = "b", coef = "biden"),
+#   set_prior(paste0("beta(", trump_shape1, ", ", trump_shape2, ")"), class = "b", coef = "trump")
+# )
+
+prior <- c(
+  set_prior(beta(biden_shape1, biden_shape2), class = "b", coef = "biden"),
+  set_prior(beta(trump_shape1, trump_shape2), class = "b", coef = "trump"))
+
+# Define the model formula
+
+formula1 <- pct ~ pollscore + days_taken_from_election + methodology
+
+# Run the Bayesian model
+model <- brm(
+  formula = formula1,
+  data = baye_model_data,
+  family = categorical(),
+  prior = prior,
+  chains = 4,
+  iter = 2000,
+  warmup = 1000,
+  cores =4
+)
+
+# Summarize the model
+summary(model)
+
 ```

### 2024-11-01 paper/paper.qmd
update paper

@@
 library(brms)
+library(dplyr)
+library(lubridate)
 ```
@@
 The remainder of this paper is structured as follows: [update this …]
+
+The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here] and lubridate [@lubridate].
 
@@
 ```{r}
-#| label: fig-clean-data-histogram-Harris-Trump
-#| fig-cap: the average PCT vs State for Harris and Trump
+#| label: fig-clean-data-histogram-Harris
+#| fig-cap: the average PCT vs State for Harris
 #| echo: false
 #| warning: false
-#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
-#| layout-nrow: 2
-#| fig-width: 40
-#| fig-height: 20
+#| height: 40
+
 Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
@@
   summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = state, y = avg_pct)) + 
-  geom_col()  +
-  theme_minimal() +
-  labs(
-    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
-  )
-
-Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
-Trump_data |>
-  group_by(state) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = state, y = avg_pct)) + 
+  ggplot(mapping = aes(x = avg_pct, y = state)) + 
   geom_col()  +
@@
 #   )
+```
+
+```{r}
+#| label: fig-clean-data-histogram-trump
+#| fig-cap: the average PCT vs State for Trump
+#| echo: false
+#| warning: false
+#| height: 40
+
+Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
+Trump_data |>
+  group_by(state) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  ggplot(mapping = aes(x = avg_pct, y = state)) + 
+  geom_col()  +
+  theme_minimal() +
+  labs(
+    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
+  )
+
 ```
@@
 plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
-
-```
-
-```{r}
-
-# Transform Biden and Trump’s vote shares to fit a beta distribution
-# Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
-total_votes <- 100
-biden_shape1 <- 51.3 / total_votes * 10
-biden_shape2 <- (1 - 51.3 / total_votes) * 10
-trump_shape1 <- 46.8 / total_votes * 10
-trump_shape2 <- (1 - 46.8 / total_votes) * 10
-
-# Set up priors using beta distributions based on 2020 vote shares
-# prior <- c(
-#   set_prior(paste0("beta(", biden_shape1, ", ", biden_shape2, ")"), class = "b", coef = "biden"),
-#   set_prior(paste0("beta(", trump_shape1, ", ", trump_shape2, ")"), class = "b", coef = "trump")
-# )
-
-prior <- c(
-  set_prior(beta(biden_shape1, biden_shape2), class = "b", coef = "biden"),
-  set_prior(beta(trump_shape1, trump_shape2), class = "b", coef = "trump"))
-
-# Define the model formula
-
-formula1 <- pct ~ pollscore + days_taken_from_election + methodology
-
-# Run the Bayesian model
-model <- brm(
-  formula = formula1,
-  data = baye_model_data,
-  family = categorical(),
-  prior = prior,
-  chains = 4,
-  iter = 2000,
-  warmup = 1000,
-  cores =4
-)
-
-# Summarize the model
-summary(model)
 

### 2024-11-01 scripts/05-exploratory_data_analysis.R

@@ -1,33 +1,23 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Purpose: Visualize the relationship between variables and overall trend of response
+# Author: Lexun Yu
+# Date: 1 November 2024
+# Contact: lx.yu@mail.utoronto.ca
+# Pre-requisites: None
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 
 # Plot poll score and sample size
 
+pairs(analysis_data)
+
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
 
 
 #### Save model ####

### 2024-11-02 paper/paper.qmd
update measurement

@@
 
-In this dataset, each row represents a polling question that records the variables of interest.
-Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
-This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
+FiveThirtyEight’s polling measurement approach transforms raw polling data from multiple sources into one dataset by following a structured, multi-step process.
+First, it collects polling data from a variety of sources that meet its methodological and ethical standards.
+According to @radcliffe_538s_2023, polls included must disclose key details, such as the pollster’s name, survey dates, and the sampled population.
+FiveThirtyEight also excludes any hypothetical matchups or polls asking for general support of a party.
+The average result is reflected in the 'pct' column.
+
+Next, FiveThirtyEight applies a series of adjustments to the data to account for sample size, timing, and pollster quality.
+As stated by @silver_polls_2008, the polls are weighted more heavily if they associate with larger sample.
+The sample size is documented in the data set as 'sample_size'.
+Poll recency is another critical factor, with newer polls weighted more to reflect current opinion trends.
+The 'pollscore' in the dataset is an indicator of pollster transparency and reliability in their methodology, according to @morris_trump_2024.
+
+Further adjustments address polling biases, such as “house effects” (systematic leanings toward a particular candidate) and the mode of polling.
+Polls from partisan sources are included, but FiveThirtyEight applies a correction to counteract inherent biases [@morris_trump_2024].
 

### 2024-11-02 paper/paper.qmd
colin code update3

--- 2024-11-01 paper/paper.qmd
@@
 library(knitr)
-library(arrow)
+library(brms)
 library(rstanarm)
-library(brms)
+library(gridExtra)
+```
+
+```{r}
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
 ```
@@
 However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
-A more adaptable model is required to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
 
-We build a hybrid election forecasting model following the strategies mentioned by @pasek.
+This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
 As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
@@
 
-In this paper, we predict the 2024 us election result with the hybrid election forcasting model.
-We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that uses Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
+In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model.
+We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
 presidential election.
@@
 For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
-- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
+- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
 - **numeric predictor:**\
- sample size(sample_size: The total number of respondents participating in the poll) \
- timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
- pollscore(A numeric value representing the score or reliability of the pollster in question) \
+ `sample size`(sample_size: The total number of respondents participating in the poll) \
+ `timegap`(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
+ `pollscore`(A numeric value representing the score or reliability of the pollster in question) \
 - **categorical predictor**
-state(The U.S. state where the poll was conducted or focused) \
-methodology(The method used to conduct the poll)\
+`state`(The U.S. state where the poll was conducted or focused) \
+`methodology`(The method used to conduct the poll)\
 
 ## Explore the data
+```{r}
+#| label: tbl-explore-data
+#| echo: false
+#| warning: false
+a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'sample size', y='pct', 
+       title = 'pct vs sample size')
+
+b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'pollscore', y='pct', 
+       title = 'pct vs pollscore')
+
+c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'wind speed', y='pct', 
+       title = 'pct vs methodology')
+
+d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'max gust speed', y='pct', 
+       title = 'pct vs days taken from election')
+
+e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'total snow', y='pct', 
+       title = 'pct vs state')
+
+grid.arrange(a, b, c, d, e, nrow=2)
+```
+
+- pct vs sample size: This scatter plot shows the pct against the sample size, with a fitted trend line indicating a slight positive relationship.
+The data points are denser for lower sample sizes, suggesting that smaller sample sizes are more common in the dataset.
+
+- pct vs pollscore: This scatter plot illustrates the pct against pollscore.
+The fitted trend line suggests a weak negative relationship between pollscore and pct.
+The points are scattered without a strong linear pattern.
+
+- pct vs methodology: A boxplot comparing pct for different polling methodologies.
+The pct distribution varies across methodologies, with some showing greater spread or median differences.
+This suggests that the polling methodology may influence pct outcomes.
+
+- pct vs days taken from election: A scatter plot displaying pct versus the number of days before the election.
+The trend line indicates a slight negative relationship, suggesting that as the election date approaches, pct may decrease slightly.
+
+- pct vs state: A boxplot depicting pct across different states.
+The pct distribution varies by state, with some states showing wider variability or different median values, implying state-specific effects on pct.
+
+```{r}
+# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
+numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
+
+# Create the pairs plot
+pairs(numeric_data)
+```
+
+The pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
+The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
+The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
+No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
 
@@
 #| warning: false
-cleaned_data = read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
 
@@
 ```{r}
+#| echo: false
+#| warning: false
 just_harris_data = Harris_data |> na.omit()
-model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology, data = just_harris_data)
+lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+just_harris_data = Harris_data |> na.omit()
+model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
 summary(model_MLR)
@@
   labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
+```
+
+```{r}
+#| echo: false
+#| warning: false
+ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
 ```
@@
 summary(bayesian_model_1)
+```
 
+This plot displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.
+
+The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
+The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
+However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
+
+The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
+Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
+Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.
+
+```{r}
 # Plot random effects
 plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+```
 
-```
+This plot displays the estimated posterior distribution for the intercept parameter in the Bayesian model.
+The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S.
+election data used in the analysis.
+The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.
+
+The range of this credible interval spans from approximately 50 to 54 percent, suggesting that the model estimates the baseline level of support for Kamala Harris to be around this range.
+The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
+
+In conclusion, the estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
+election data analyzed, is centered around 52%, with a credible interval spanning approximately from 50% to 54%.
+This suggests that, according to the Bayesian model, the underlying support for Kamala Harris.
 
@@
 
-# Transform Biden and Trump’s vote shares to fit a beta distribution
-# Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
-total_votes <- 100
-biden_shape1 <- 51.3 / total_votes * 10
-biden_shape2 <- (1 - 51.3 / total_votes) * 10
-trump_shape1 <- 46.8 / total_votes * 10
-trump_shape2 <- (1 - 46.8 / total_votes) * 10
+# # Transform Biden and Trump’s vote shares to fit a beta distribution
+# # Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
+# total_votes <- 100
+# biden_shape1 <- 51.3 / total_votes * 10
+# biden_shape2 <- (1 - 51.3 / total_votes) * 10
+# trump_shape1 <- 46.8 / total_votes * 10
+# trump_shape2 <- (1 - 46.8 / total_votes) * 10
 
-# Set up priors using beta distributions based on 2020 vote shares
+# # Set up priors using beta distributions based on 2020 vote shares
 # prior <- c(
@@
 
-prior <- c(
-  set_prior(beta(biden_shape1, biden_shape2), class = "b", coef = "biden"),
-  set_prior(beta(trump_shape1, trump_shape2), class = "b", coef = "trump"))
+# # Define the model formula
+# formula <- pct ~ pollscore + days_taken_from_election + methodology
 
-# Define the model formula
+# # Run the Bayesian model
+# model <- brm(
+#   formula = formula,
+#   data = baye_model_data,
+#   family = categ0orical(),
+#   prior = prior,
+#   chains = 4,
+#   iter = 2000,
+#   warmup = 1000,
+#   cores =4
+# )
 
-formula1 <- pct ~ pollscore + days_taken_from_election + methodology
-
-# Run the Bayesian model
-model <- brm(
-  formula = formula1,
-  data = baye_model_data,
-  family = categorical(),
-  prior = prior,
-  chains = 4,
-  iter = 2000,
-  warmup = 1000,
-  cores =4
-)
-
-# Summarize the model
-summary(model)
+# # Summarize the model
+# summary(model)
 
 ```
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-us
 
@@
 
-# Methodology of YouGov
-YouGov's methodology documentations are seperated in two articles.
-The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
-
-## Population, Frame, and Sample
-As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
-The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
-Voter files indicates whether asomeone voted in a given election, thus YouGov's population covers all voters in previous US elections.
-
-YouGov’s sampling frame consists of its online panel members.
-These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
-YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
-
-YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
-YouGov interviews nearly 100,000 people in the first set of estimates.
-For the second set of estimates, YouGov didn't just start over with a new sample.
-They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
-
-## Sample Recruitment
-Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
-They must provide demographic details upon joining, which helps in selecting representative samples for each survey.
-When respondents complete a survey, they are awarded points that can be exchanged for money.
-
-## Sampling Approach and Trade-offs
-YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
-This method allows quick and cost-effective data collection.
-However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
-YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
-
-## Non-response Handling
-YouGov apply statistical weighting to adjust for the differences between the sample and target population.
-The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
-Additionally, quality control measures exclude unreliable responses to improve data accuracy.
-The respondents are offered a small incentive to decrease the non-response and increase participation.
-
-## Strengths and Weaknesses of the Questionnaire
-YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
-The pollster can recruit a large amount of panelists because of the online format.
-Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
-
-As a non-probability sample, it might miss certain demographic groups not covered by the online population.
-While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
-Additionaly, the categories in the surbey is oversimplified with bias.
-For instance, in the poll result published by YouGov, gender is divided into Male and Female.
-Race is divided into White, Black, Hispanic and Other.
-This indicates a lack of representation.
-
 # FiveThirtyEight Licenses

### 2024-11-02 scripts/03-clean_data.R

@@ -42,6 +42,8 @@ cleaned_data <- cleaned_data %>%
     end_date, start_date, state, candidate_name, pct
   )
 
+cleaned_data = cleaned_data |> na.omit()
+
 # Delete NA columns
 # cleaned_data <- raw_data |>
 #   select(-sponsor_ids,

### 2024-11-02 paper/paper.qmd
update paper
added data graphs

--- 2024-11-01 paper/paper.qmd
@@
 ---
-title: "Prediction of 2024 US election ..."
+title: "Prediction of US election with linear model"
 author: 
@@
 date-format: long
-abstract: "We forecast the winner of the 2024 US presidential election using “poll-of-polls” by building a linear model."
+abstract: "This paper forecast the winner of the upcoming US presidential election using “poll-of-polls” by building a linear model."
 format: pdf
@@
 library(tidyverse)
-library(palmerpenguins)
-library(here)
-library(tibble)
-library(knitr)
-library(arrow)
-library(rstanarm)
-library(brms)
-library(dplyr)
-library(lubridate)
+library(modelsummary)
 ```
 
-# Introduction {#sec-intro}
-Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States.
-Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result.
-As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
-However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
-A more adaptable model is required to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+\newpage
 
-We build a hybrid election forecasting model following the strategies mentioned by @pasek.
-As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
-Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics.
-Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
-
-In this paper, we predict the 2024 us election result with the hybrid election forcasting model.
-We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that uses Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
-presidential election.
-
-The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump.
-The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.
-
-The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]
-
-The remainder of this paper is structured as follows: [update this …]
-
-The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here] and lubridate [@lubridate].
-
-# Data {#sec-data}
-
-## Overview
-
-For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
-- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
-- **numeric predictor:**\
- sample size(sample_size: The total number of respondents participating in the poll) \
- timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
- pollscore(A numeric value representing the score or reliability of the pollster in question) \
-- **categorical predictor**
-state(The U.S. state where the poll was conducted or focused) \
-methodology(The method used to conduct the poll)\
-
-## Explore the data
-
-## Measurement
-
-In this dataset, each row represents a polling question that records the variables of interest.
-Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
-This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
-
-## Clean Data
-The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
-First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable.
-Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.
-
-We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S.
-election date.
-Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy.
-Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.
+# Graphs and Tables
 
 ```{r}
-#| label: tbl-cleaned-data
-#| tbl-cap: Sample of cleaned US election data 
-#| echo: false
-#| warning: false
-cleaned_data = read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
 
-cleaned_data |>
-  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
-  head(6) |>
-  kable(
-    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "candidate_name"),
-    booktabs = TRUE
-    ) 
+clean_data <- read_csv("../data/02-analysis_data/cleaned_data.csv")
+head(clean_data)
+
+head(clean_data)
 ```
 
-## Basic Statistics Summary for Data
+## Relationship Pollscore and Sample Size
 
 ```{r}
-#| label: fig-clean-data-histogram-Harris
-#| fig-cap: the average PCT vs State for Harris
-#| echo: false
-#| warning: false
-#| height: 40
 
-Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
-Harris_data |>
-  group_by(state) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = avg_pct, y = state)) + 
-  geom_col()  +
-  theme_minimal() +
-  labs(
-    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
-  )
+# removing outliers and na's
 
-# cleaned_data %>%
-#   mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
-#                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
-#   ggplot(mapping = aes(x = month, y = total_decedents)) + 
-#   geom_bar(stat = "identity") +
-#   theme_minimal() +
-#   labs(
-#     x = "Month", 
-#     y = "Total Decedents"
-#   )
-```
+graph_data <- clean_data %>% 
+  filter(!is.na(pollscore) & !is.na(sample_size))
 
-```{r}
-#| label: fig-clean-data-histogram-trump
-#| fig-cap: the average PCT vs State for Trump
-#| echo: false
-#| warning: false
-#| height: 40
+outlier_top <- (quantile(graph_data$sample_size, 0.75) + 
+                  IQR(graph_data$sample_size)*1.5)
 
-Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
-Trump_data |>
-  group_by(state) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = avg_pct, y = state)) + 
-  geom_col()  +
-  theme_minimal() +
-  labs(
-    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
-  )
+outlier_bottom <- (quantile(graph_data$sample_size, 0.25) - 
+                  IQR(graph_data$sample_size)*1.5)
+
+graph_data <- clean_data %>%
+  filter(sample_size < outlier_top & sample_size > outlier_bottom)
+
+graph_data %>% filter(!is.na(pollscore) & !is.na(sample_size)) %>% 
+  ggplot(aes(x = pollscore, y = log(sample_size))) +
+  geom_point() + geom_smooth(method = "lm", se = FALSE, color = "red")
+
+# this model is really here just for explain, this realtionship should not
+# be treated as lienar
+lm_score_sample <- lm(log(sample_size) ~ pollscore, data = graph_data)
+summary(lm_score_sample)
 
@@
 
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-us
-
-## Basic Model
-```{r}
-just_harris_data = Harris_data |> na.omit()
-model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology, data = just_harris_data)
-summary(model_MLR)
-just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
-
-predictions = predict(model_MLR, just_harris_data)
-
-ggplot(just_harris_data, aes(x = end_date)) +
-  geom_point(aes(y = pct), color = "black") +
-  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
-  facet_wrap(vars(methodology)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
-```
+## Relationship Between Pollscore and Transparency Score
 
 ```{r}
-baye_model_data = just_harris_data
-baye_model_data$pct = as.factor(baye_model_data$pct)
-baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
-baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
-baye_model_data$methodology = as.factor(baye_model_data$methodology)
-baye_model_data$pct <- as.numeric(baye_model_data$pct)
-# Define the Bayesian model with brms
 
-formula <- pct ~ pollscore + days_taken_from_election + (1 | methodology) + (1 | state)
+clean_data %>% filter(!is.na(pollscore) & !is.na(transparency_score)) %>% 
+  ggplot(aes(x = pollscore, y = transparency_score)) +
+  geom_point() + geom_smooth(method = "lm", se = FALSE, color = "red")
 
-priors = normal(0, 2.5, autoscale = TRUE)
-
-bayesian_model_1 <- stan_glmer(
-  formula = formula,
-  data = just_harris_data,
-  family = gaussian(),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
-pp_check(bayesian_model_1)
-summary(bayesian_model_1)
-
-# Plot random effects
-plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+# this model is really here just for explain, this realtionship should not
+# be treated as lienar
+lm_score_sample <- lm(pollscore ~ transparency_score, data = clean_data)
+summary(lm_score_sample)
 
@@
 
-# Model
+## Relationship Between Days Taken From Electon and Support Percentage
 
-The goal of our modelling strategy is twofold. Firstly,...
+```{r}
+head(clean_data)
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+clean_data %>% 
+  filter(!is.na(days_taken_from_election) & !is.na(pct)) %>% 
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>% 
+  ggplot(aes(x = days_taken_from_election, y = pct)) +
+  geom_point() + 
+  geom_smooth(method = "lm", se = FALSE, color = "red") +
+  labs(x = "Days Before Election",
+       y = "Poll Percentage") + facet_wrap(~candidate_name)
 
-## Model set-up
+```
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+## Metholody And Support Percentage
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+```{r}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-us
+head(clean_data)
 
-### Model justification
+clean_data %>% 
+  filter(!is.na(methodology) & !is.na(pct)) %>% 
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>% 
+  ggplot(aes(x = methodology, y = pct, fill = candidate_name)) +
+  geom_boxplot(position = position_dodge(width = 0.9)) +
+  labs(x = "Methodology", y = "Percentage", 
+       title = "Poll Percentages by Methodology and Candidate",
+       fill = "Candidate") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "top") +
+  scale_fill_brewer(palette = "Set1")
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-# Methodology of YouGov
-YouGov's methodology documentations are seperated in two articles.
-The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
-
-## Population, Frame, and Sample
-As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
-The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
-Voter files indicates whether asomeone voted in a given election, thus YouGov's population covers all voters in previous US elections.
-
-YouGov’s sampling frame consists of its online panel members.
-These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
-YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
-
-YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
-YouGov interviews nearly 100,000 people in the first set of estimates.
-For the second set of estimates, YouGov didn't just start over with a new sample.
-They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
-
-## Sample Recruitment
-Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
-They must provide demographic details upon joining, which helps in selecting representative samples for each survey.
-When respondents complete a survey, they are awarded points that can be exchanged for money.
-
-## Sampling Approach and Trade-offs
-YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
-This method allows quick and cost-effective data collection.
-However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
-YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
-
-## Non-response Handling
-YouGov apply statistical weighting to adjust for the differences between the sample and target population.
-The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
-Additionally, quality control measures exclude unreliable responses to improve data accuracy.
-The respondents are offered a small incentive to decrease the non-response and increase participation.
-
-## Strengths and Weaknesses of the Questionnaire
-YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
-The pollster can recruit a large amount of panelists because of the online format.
-Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
-
-As a non-probability sample, it might miss certain demographic groups not covered by the online population.
-While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
-Additionaly, the categories in the surbey is oversimplified with bias.
-For instance, in the poll result published by YouGov, gender is divided into Male and Female.
-Race is divided into White, Black, Hispanic and Other.
-This indicates a lack of representation.
-
-# FiveThirtyEight Licenses
-[FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).
-
-\newpage
-
-# References
+```

### 2024-11-02 scripts/03-clean_data.R

@@ -34,9 +34,11 @@ cleaned_data <- cleaned_data %>%
   select(poll_id, pollster_id, pollster, question_id,
          sample_size, pollscore, methodology,
          days_taken_from_election, transparency_score, party,
-         end_date, start_date, state, answer, pct)
+         end_date, start_date, state, candidate_name, pct)
 
 cleaned_data %>% select(state) %>% unique() %>% data.frame()
+
+cleaned_data %>% select(state) %>% unique()
 
 # Delete NA columns
 # cleaned_data <- raw_data |>

### 2024-11-02 paper/paper.qmd
update unfinished test simulated data script

--- 2024-11-02 paper/paper.qmd
@@
 As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
-However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
-A more adaptable model is required to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+However, the traditional poll aggregation does not reflect dynamics of an election, especially with differences in quality of polls.
+A more adaptable model is required to predict the election result based on both polling data and additional variables, such as sample size and pollster credibility.
 
@@
 Further adjustments address polling biases, such as “house effects” (systematic leanings toward a particular candidate) and the mode of polling.
+Likely-voter adjustment are applied to all polls of registered voters and adults.
+Third party adjustment are applied to polls that only consider Biden and Trump as the candidates.
 Polls from partisan sources are included, but FiveThirtyEight applies a correction to counteract inherent biases [@morris_trump_2024].

### 2024-11-02 scripts/00-simulate_data.R

@@ -10,6 +10,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 set.seed(853)
 
 
@@ -58,4 +59,4 @@ analysis_data <- tibble(
 
 
 #### Save data ####
-write_csv(analysis_data, "data/00-simulated_data/simulated_data.csv")
+write_parquet(analysis_data, "data/00-simulated_data/simulated_data.parquet")

### 2024-11-02 scripts/01-test_simulated_data.R

@@ -1,20 +1,21 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian
-# electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated US
+# election poll dataset.
+# Author: Lexun Yu
+# Date: 2 November 2024
+# Contact: lx.yu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
 # - The `tidyverse` package must be installed and loaded
 # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
+library(testthat)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+analysis_data <- read_parquet("data/00-simulated_data/simulated_data.parquet")
 
 # Test if the data was successfully loaded
 if (exists("analysis_data")) {
@@ -23,69 +24,39 @@ if (exists("analysis_data")) {
   stop("Test Failed: The dataset could not be loaded.")
 }
 
-
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
-}
+# Check if the dataset has 200 rows
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
+test_that("Dataset row count is 200", {
+  expect_equal(nrow(analysis_data), 200, 
+               info = "The dataset does not have 200 rows.")
+})
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
-} else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
-}
+# Check if the dataset has 6 columns
+test_that("Dataset column count is 6", {
+  expect_equal(ncol(analysis_data), 6, 
+               info = "The dataset does not have 6 columns.")
+})
 
[…]

### 2024-11-02 paper/paper.qmd
update measurement in paper and simulation

@@
 
-FiveThirtyEight’s polling measurement approach transforms raw polling data from multiple sources into one dataset by following a structured, multi-step process.
-First, it collects polling data from a variety of sources that meet its methodological and ethical standards.
-According to @radcliffe_538s_2023, polls included must disclose key details, such as the pollster’s name, survey dates, and the sampled population.
-FiveThirtyEight also excludes any hypothetical matchups or polls asking for general support of a party.
-The average result is reflected in the 'pct' column.
+The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys.
+According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024].
+The dataset contains a list of questions in polls and surveys and their results.
+When a new poll is conducted, the poll is appended to the dataset.
+FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.
 
-Next, FiveThirtyEight applies a series of adjustments to the data to account for sample size, timing, and pollster quality.
-As stated by @silver_polls_2008, the polls are weighted more heavily if they associate with larger sample.
-The sample size is documented in the data set as 'sample_size'.
-Poll recency is another critical factor, with newer polls weighted more to reflect current opinion trends.
-The 'pollscore' in the dataset is an indicator of pollster transparency and reliability in their methodology, according to @morris_trump_2024.
+In each of the polls recorded by FiveThirtyEight, all options of each question are recorded in candidate_names, while the proportion of respondents choosing that option is recorded in pct as a percentage.
+The pct is our primary response variable.
 
-Further adjustments address polling biases, such as “house effects” (systematic leanings toward a particular candidate) and the mode of polling.
-Likely-voter adjustment are applied to all polls of registered voters and adults.
-Third party adjustment are applied to polls that only consider Biden and Trump as the candidates.
-Polls from partisan sources are included, but FiveThirtyEight applies a correction to counteract inherent biases [@morris_trump_2024].
+There are limitations in these measurements.
+The differences in sampling methods, wording in survey questions and systematic biases are all reflected in the outcome [@radcliffe_538s_2023].
+Thus, in the dataset, FiveThirtyEight includes other variables, such as pollscore and sample_size for their prediction model in addition to the polling results for transparency and accuracy.
+As an example, pollscore indicates the pollster’s reliability rating, a low pollscore indicates a higher reliability rating [@silver_polls_2008].
+
+These variables combined allow researchers to reliably analyze and predict the 2024 US election result over time.
 

### 2024-11-02 scripts/00-simulate_data.R

@@ -1,7 +1,8 @@
+#### Workspace setup ####
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
+# Purpose: Simulates a dataset of Australian electoral divisions, including the
   #state and party that won each division.
-# Author: Lexun Yu
+# Author: Lexun Yu, Siddharth Gowda
 # Date: 22 October 2024
 # Contact: lx.yu@mail.utoronto.ca
 # License: MIT
@@ -10,53 +11,97 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(arrow)
 set.seed(853)
 
 
 #### Simulate data ####
 # State names
-states <- state.name
-answer <- c("Trump", "Harris", "Oliver", "West")
+# Add National
+states <- c(state.name, "National")
+candidate_names <- c("Donald Trump", "Kamala Harris", "Jill Stein",
+                    "Robert F. Kennedy")
 pollster <- c("Marist", "YouGov", "Ipsos", "Emerson", "SurveyUSA")
 
+# Reliably blue states
+blue_states <- c("California", "Colorado", "Connecticut", "Delaware", "Hawaii",
+                 "Illinois", "Maine", "Maryland", "Massachusetts", "Minnesota",
+                 "New Jersey", "New Mexico", "New York", "Oregon", "Rhode Island",
+                 "Vermont", "Virginia", "Washington", "District of Columbia")
+
+# Reliably red states
+red_states <- c("Alabama", "Alaska", "Arkansas", "Florida", "Idaho", "Indiana",
+                "Iowa", "Kansas", "Kentucky", "Louisiana", "Mississippi",
+                "Missouri", "Montana", "Nebraska", "North Dakota", "Ohio",
+                "Oklahoma", "South Carolina", "South Dakota", "Tennessee",
+                "Texas", "Utah", "West Virginia", "Wyoming")
+
+# Swing states (from the search results)
+swing_states <- c("Arizona", "Georgia", "Michigan", "Nevada",
+                  "North Carolina", "Pennsylvania", "Wisconsin")
+
 # Generate Poll IDs and States
-unique_poll_ids <- sample(10000:99999, 50)
+unique_poll_ids <- sample(10000:99999, length(states))
 poll_ids <- rep(unique_poll_ids, each = 4)
 state_column <- rep(states, each = 4)
[…]

### 2024-11-03 scripts/00-simulate_data.R
update test

@@ -17,7 +17,7 @@ set.seed(853)
 #### Simulate data ####
 # State names
 # Add National
-states <- c(state.name, "National")
+states <- c(state.name, "National", "Maine CD-2", "Maine CD-1", "Nebraska CD-2")
 candidate_names <- c("Donald Trump", "Kamala Harris", "Jill Stein",
                     "Robert F. Kennedy")
 pollster <- c("Marist", "YouGov", "Ipsos", "Emerson", "SurveyUSA")
@@ -41,7 +41,9 @@ swing_states <- c("Arizona", "Georgia", "Michigan", "Nevada",
 
 # Generate Poll IDs and States
 unique_poll_ids <- sample(10000:99999, length(states))
+unique_question_ids <- sample(10000:999999, length(states))
 poll_ids <- rep(unique_poll_ids, each = 4)
+question_ids <- rep(unique_question_ids, each = 4)
 state_column <- rep(states, each = 4)
 state_column <- rep(state_column, length.out = 4*length(states))
 sample_size <- sample(100:20000, length(states))
@@ -82,9 +84,10 @@ generate_percentages <- function(state) {
 analysis_data <- tibble(
   pollster = pollster_column,
   poll_id = poll_ids,
+  question_ids = question_ids,
   state = state_column,
   sample_size = sample_size,
-  candidate_name = rep(candidate_names, length.out = 204),
+  candidate_name = rep(candidate_names, length.out = 216),
 )
 
 analysis_data <- analysis_data %>% mutate(pct = 0.0)
@@ -102,6 +105,8 @@ for (id_ in unique(analysis_data$poll_id)) {
       TRUE ~ pct  # keep original value if no match
     ))
 }
+
+analysis_data$pct <- analysis_data$pct * 100
 
 #### Save data ####
 write_parquet(analysis_data, "data/00-simulated_data/simulated_data.parquet")

### 2024-11-03 scripts/01-test_simulated_data.R

@@ -28,20 +28,20 @@ if (exists("analysis_data")) {
 
 # Check if the dataset has 200 rows
 
-test_that("Dataset row count is 200", {
-  expect_equal(nrow(analysis_data), 200, 
-               info = "The dataset does not have 200 rows.")
+test_that("Dataset row count is 216", {
+  expect_equal(nrow(analysis_data), 216, 
+               info = "The dataset does not have 216 rows.")
 })
 
-# Check if the dataset has 6 columns
-test_that("Dataset column count is 6", {
-  expect_equal(ncol(analysis_data), 6, 
-               info = "The dataset does not have 6 columns.")
+# Check if the dataset has 7 columns
+test_that("Dataset column count is 7", {
+  expect_equal(ncol(analysis_data), 7, 
+               info = "The dataset does not have 7 columns.")
 })
 
 # Check if all values in the 'state' column are a valid state name
 valid_states <- 
-  c(state.name, "National")
+  c(state.name, "National", "Maine CD-2", "Maine CD-1", "Nebraska CD-2")
 
 test_that("The 'state' column contains only valid US state names", {
   expect_true(all(analysis_data$state %in% valid_states), 
@@ -55,8 +55,22 @@ test_that("The dataset contains no missing values", {
               info = "The dataset contains missing values.")
 })
 
-# Check if the percentages are all above 0
-test_that("The percentage are positive values", {
-  expect_true(all(analysis_data$pct >= 0, 
-  info = ""))
+# Check if the percentages are all above 0 and below 100
+test_that("The percentages are in the range between 0 and 100", {
+  expect_true(all(analysis_data$pct >= 0 & analysis_data$pct <= 100), 
+              info = "The percentages are not in the range between 0 and 100")
 })
+
+
+# Check the variable type
+  test_that("pct columns are of type double", {
+    # Check numeric (float) columns
+    expect_type(analysis_data$pct, "double")
+  })
+  
+  test_that("Character columns are of type character", {
+    # Check character columns
+    expect_type(analysis_data$pollster, "character")
+    expect_type(analysis_data$state, "character")
+    expect_type(analysis_data$candidate_name, "character")
+  })

### 2024-11-03 scripts/03-clean_data.R

@@ -57,6 +57,8 @@ cleaned_data$adjusted_weight <- 1 / abs(cleaned_data$pollscore)
   cleaned_data_state <- cleaned_data |>
     filter(state != "National")
 
+cleaned_data <- cleaned_data |> na.omit()
+
 
 # Delete NA columns
 # cleaned_data <- raw_data |>

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -1,71 +1,78 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Purpose: Tests cleaned data
+# Author: Lexun Yu
+# Date:2 November 2024
+# Contact: lx.yu@mail.utoronto.ca
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 
+# Test if the data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+
+# Check if the dataset has 1732 rows
+
+test_that("Dataset row count is 1732", {
+  expect_equal(nrow(analysis_data), 1732, 
+               info = "The dataset does not have 1732 rows.")
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
+# Check if the dataset has 14 columns
+test_that("Dataset column count is 14", {
+  expect_equal(ncol(analysis_data), 14, 
+               info = "The dataset does not have 14 columns.")
 })
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
+# Check if all values in the 'state' column are a valid state name
+valid_states <- 
+  c(state.name, "National", "Maine CD-2", "Maine CD-1", "Nebraska CD-2")
+
+test_that("The 'state' column contains only valid US state names", {
+  expect_true(all(analysis_data$state %in% valid_states), 
+              info = "The 'state' column contains invalid state names.")
 })
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
[…]

### 2024-11-03 scripts/00-simulate_data.R
update style

@@ -1,7 +1,7 @@
 #### Workspace setup ####
 #### Preamble ####
 # Purpose: Simulates a dataset of Australian electoral divisions, including the
-  #state and party that won each division.
+# state and party that won each division.
 # Author: Lexun Yu, Siddharth Gowda
 # Date: 22 October 2024
 # Contact: lx.yu@mail.utoronto.ca
@@ -18,26 +18,34 @@ set.seed(853)
 # State names
 # Add National
 states <- c(state.name, "National", "Maine CD-2", "Maine CD-1", "Nebraska CD-2")
-candidate_names <- c("Donald Trump", "Kamala Harris", "Jill Stein",
-                    "Robert F. Kennedy")
+candidate_names <- c(
+  "Donald Trump", "Kamala Harris", "Jill Stein",
+  "Robert F. Kennedy"
+)
 pollster <- c("Marist", "YouGov", "Ipsos", "Emerson", "SurveyUSA")
 
 # Reliably blue states
-blue_states <- c("California", "Colorado", "Connecticut", "Delaware", "Hawaii",
-                 "Illinois", "Maine", "Maryland", "Massachusetts", "Minnesota",
-                 "New Jersey", "New Mexico", "New York", "Oregon", "Rhode Island",
-                 "Vermont", "Virginia", "Washington", "District of Columbia")
+blue_states <- c(
+  "California", "Colorado", "Connecticut", "Delaware", "Hawaii",
+  "Illinois", "Maine", "Maryland", "Massachusetts", "Minnesota",
+  "New Jersey", "New Mexico", "New York", "Oregon", "Rhode Island",
+  "Vermont", "Virginia", "Washington", "District of Columbia"
+)
 
 # Reliably red states
-red_states <- c("Alabama", "Alaska", "Arkansas", "Florida", "Idaho", "Indiana",
-                "Iowa", "Kansas", "Kentucky", "Louisiana", "Mississippi",
-                "Missouri", "Montana", "Nebraska", "North Dakota", "Ohio",
-                "Oklahoma", "South Carolina", "South Dakota", "Tennessee",
-                "Texas", "Utah", "West Virginia", "Wyoming")
+red_states <- c(
+  "Alabama", "Alaska", "Arkansas", "Florida", "Idaho", "Indiana",
+  "Iowa", "Kansas", "Kentucky", "Louisiana", "Mississippi",
[…]

### 2024-11-03 scripts/01-test_simulated_data.R

@@ -29,48 +29,53 @@ if (exists("analysis_data")) {
 # Check if the dataset has 200 rows
 
 test_that("Dataset row count is 216", {
-  expect_equal(nrow(analysis_data), 216, 
-               info = "The dataset does not have 216 rows.")
+  expect_equal(nrow(analysis_data), 216,
+    info = "The dataset does not have 216 rows."
+  )
 })
 
 # Check if the dataset has 7 columns
 test_that("Dataset column count is 7", {
-  expect_equal(ncol(analysis_data), 7, 
-               info = "The dataset does not have 7 columns.")
+  expect_equal(ncol(analysis_data), 7,
+    info = "The dataset does not have 7 columns."
+  )
 })
 
 # Check if all values in the 'state' column are a valid state name
-valid_states <- 
+valid_states <-
   c(state.name, "National", "Maine CD-2", "Maine CD-1", "Nebraska CD-2")
 
 test_that("The 'state' column contains only valid US state names", {
-  expect_true(all(analysis_data$state %in% valid_states), 
-              info = "The 'state' column contains invalid state names.")
+  expect_true(all(analysis_data$state %in% valid_states),
+    info = "The 'state' column contains invalid state names."
+  )
 })
 
 
 # Check if there are any missing values in the dataset
 test_that("The dataset contains no missing values", {
-  expect_true(all(!is.na(analysis_data)), 
-              info = "The dataset contains missing values.")
+  expect_true(all(!is.na(analysis_data)),
+    info = "The dataset contains missing values."
+  )
 })
 
 # Check if the percentages are all above 0 and below 100
 test_that("The percentages are in the range between 0 and 100", {
-  expect_true(all(analysis_data$pct >= 0 & analysis_data$pct <= 100), 
-              info = "The percentages are not in the range between 0 and 100")
+  expect_true(all(analysis_data$pct >= 0 & analysis_data$pct <= 100),
+    info = "The percentages are not in the range between 0 and 100"
+  )
 })
 
 
 # Check the variable type
-  test_that("pct columns are of type double", {
-    # Check numeric (float) columns
-    expect_type(analysis_data$pct, "double")
-  })
-  
-  test_that("Character columns are of type character", {
-    # Check character columns
-    expect_type(analysis_data$pollster, "character")
-    expect_type(analysis_data$state, "character")
-    expect_type(analysis_data$candidate_name, "character")
-  })
+test_that("pct columns are of type double", {
[…]

### 2024-11-03 scripts/03-clean_data.R

@@ -52,10 +52,10 @@ cleaned_data$adjusted_weight <- 1 / abs(cleaned_data$pollscore)
 
 
 # store national and state data seperately
-  cleaned_data_national <- cleaned_data |>
-    filter(state == "National")
-  cleaned_data_state <- cleaned_data |>
-    filter(state != "National")
+cleaned_data_national <- cleaned_data |>
+  filter(state == "National")
+cleaned_data_state <- cleaned_data |>
+  filter(state != "National")
 
 cleaned_data <- cleaned_data |> na.omit()
 

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -23,56 +23,61 @@ if (exists("analysis_data")) {
 # Check if the dataset has 1732 rows
 
 test_that("Dataset row count is 1732", {
-  expect_equal(nrow(analysis_data), 1732, 
-               info = "The dataset does not have 1732 rows.")
+  expect_equal(nrow(analysis_data), 1732,
+    info = "The dataset does not have 1732 rows."
+  )
 })
 
 # Check if the dataset has 14 columns
 test_that("Dataset column count is 14", {
-  expect_equal(ncol(analysis_data), 14, 
-               info = "The dataset does not have 14 columns.")
+  expect_equal(ncol(analysis_data), 14,
+    info = "The dataset does not have 14 columns."
+  )
 })
 
 # Check if all values in the 'state' column are a valid state name
-valid_states <- 
+valid_states <-
   c(state.name, "National", "Maine CD-2", "Maine CD-1", "Nebraska CD-2")
 
 test_that("The 'state' column contains only valid US state names", {
-  expect_true(all(analysis_data$state %in% valid_states), 
-              info = "The 'state' column contains invalid state names.")
+  expect_true(all(analysis_data$state %in% valid_states),
+    info = "The 'state' column contains invalid state names."
+  )
 })
 
 
 # Check if there are any missing values in the dataset
 test_that("The dataset contains no missing values", {
-  expect_true(all(!is.na(analysis_data)), 
-              info = "The dataset contains missing values.")
+  expect_true(all(!is.na(analysis_data)),
+    info = "The dataset contains missing values."
+  )
 })
 
 # Check if the percentages are all above 0 and below 100
 test_that("The percentages are in the range between 0 and 100", {
-  expect_true(all(analysis_data$pct >= 0 & analysis_data$pct <= 100), 
-              info = "The percentages are not in the range between 0 and 100")
+  expect_true(all(analysis_data$pct >= 0 & analysis_data$pct <= 100),
+    info = "The percentages are not in the range between 0 and 100"
+  )
 })
 
 # Check the variable type
-  test_that("pct columns are of type double", {
-    # Check numeric (float) columns
-    expect_type(analysis_data$pct, "double")
-  })
-  
-  test_that("Character columns are of type character", {
-    # Check character columns
-    expect_type(analysis_data$pollster, "character")
-    expect_type(analysis_data$state, "character")
-    expect_type(analysis_data$candidate_name, "character")
[…]

### 2024-11-03 scripts/04-test_analysis_data.R
delete result test

@@ -72,12 +72,3 @@ test_that("Character columns are of type character", {
   expect_type(analysis_data$state, "character")
   expect_type(analysis_data$candidate_name, "character")
 })
-
-
-result <- analysis_data %>%
-  group_by(question_id) %>%
-  summarize(sum_pct = sum(pct)) %>%
-  filter(abs(sum_pct - 1) > 10) # Adjust tolerance as needed
-
-# Display the irregular results
-result

### 2024-11-03 paper/paper.qmd
colin code update4

--- 2024-11-02 paper/paper.qmd
@@
 library(tidyverse)
-library(modelsummary)
+library(palmerpenguins)
+library(here)
+library(tibble)
+library(knitr)
+library(brms)
+library(rstanarm)
+library(gridExtra)
+library(Matrix)
+library(rstan)
+library(brms)
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
+```
+
+# Introduction {#sec-intro}
+Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States.
+Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result.
+As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
+However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
+This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+
+This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
+As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
+Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics.
+Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
+
+In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model.
+We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
+presidential election.
+
+The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump.
+The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.
+
+The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]
+
+The remainder of this paper is structured as follows: [update this …]
+
+# Data {#sec-data}
+
+## Overview
+
+For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
+- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
+- **numeric predictor:**\
+ `sample size`(sample_size: The total number of respondents participating in the poll) \
+ `timegap`(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
+ `pollscore`(A numeric value representing the score or reliability of the pollster in question) \
+- **categorical predictor**
+`state`(The U.S. state where the poll was conducted or focused) \
+`methodology`(The method used to conduct the poll)\
+
+## Data Exploration
+```{r}
+#| label: tbl-explore-data
+#| echo: false
+#| warning: false
+a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'sample size', y='pct', 
+       title = 'pct vs sample size')
+
+b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'pollscore', y='pct', 
+       title = 'pct vs pollscore')
+
+c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'methodology', y='pct', 
+       title = 'pct vs methodology')
+
+d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'max gust speed', y='pct', 
+       title = 'pct vs days taken from election')
+
+e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'state', y='pct', 
+       title = 'pct vs state')
+
+grid.arrange(a, b, c, d, e, nrow=2)
+```
+
+- pct vs sample size: This scatter plot shows the pct against the sample size, with a fitted trend line indicating a slight positive relationship.
+The data points are denser for lower sample sizes, suggesting that smaller sample sizes are more common in the dataset.
+
+- pct vs pollscore: This scatter plot illustrates the pct against pollscore.
+The fitted trend line suggests a weak negative relationship between pollscore and pct.
+The points are scattered without a strong linear pattern.
+
+- pct vs methodology: A boxplot comparing pct for different polling methodologies.
+The pct distribution varies across methodologies, with some showing greater spread or median differences.
+This suggests that the polling methodology may influence pct outcomes.
+
+- pct vs days taken from election: A scatter plot displaying pct versus the number of days before the election.
+The trend line indicates a slight negative relationship, suggesting that as the election date approaches, pct may decrease slightly.
+
+- pct vs state: A boxplot depicting pct across different states.
+The pct distribution varies by state, with some states showing wider variability or different median values, implying state-specific effects on pct.
+
+```{r}
+#| label: fig-pairs
+#| echo: false
+#| warning: false
+# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
+numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
+
+# Create the pairs plot
+pairs(numeric_data)
+```
+
+@fig-pairs The pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
+The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
+The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
+No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
+
+## Measurement
+
+In this dataset, each row represents a polling question that records the variables of interest.
+Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
+This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
+
+## Clean Data
+@tbl-cleaned-data The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
+First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable.
+Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.
+
+We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S.
+election date.
+Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy.
+Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.
+
+```{r}
+#| label: tbl-cleaned-data
+#| tbl-cap: Sample of cleaned US election data 
+#| echo: false
+#| warning: false
+
+cleaned_data |>
+  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
+  head(6) |>
+  kable(
+    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "candidate_name"),
+    booktabs = TRUE
+    ) 
+```
+
+## Basic Statistics Summary for Data
+
+```{r}
+#| label: fig-clean-data-histogram-Harris-Trump
+#| fig-cap: the average PCT vs State for Harris and Trump
+#| echo: false
+#| warning: false
+#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
+#| layout-nrow: 2
+#| fig-width: 40
+#| fig-height: 20
+Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+Harris_data |>
+  group_by(state) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  ggplot(mapping = aes(x = state, y = avg_pct)) + 
+  geom_col()  +
+  theme_minimal() +
+  labs(
+    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
+  )
+
+Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
+Trump_data |>
+  group_by(state) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  ggplot(mapping = aes(x = state, y = avg_pct)) + 
+  geom_col()  +
+  theme_minimal() +
+  labs(
+    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
+  )
+
+# cleaned_data %>%
+#   mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
+#                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
+#   ggplot(mapping = aes(x = month, y = total_decedents)) + 
+#   geom_bar(stat = "identity") +
+#   theme_minimal() +
+#   labs(
+#     x = "Month", 
+#     y = "Total Decedents"
+#   )
 ```
@@
 
-# Graphs and Tables
+@fig-clean-data-histogram-Harris-Trump-1 The histogram displays the average support percentage for Kamala Harris across different U.S.
+states.
+The data indicates that support for Harris varies significantly across states.
+Notable observations include relatively high average support in states such as California and New York, which are known for being more Democratic-leaning.
+On the other hand, there are states with lower average support percentages, particularly in more traditionally Republican or swing states.
+The distribution suggests regional variations in support, with some states showing consistent backing for Harris while others indicate a weaker performance.
+
+@fig-clean-data-histogram-Harris-Trump-2 The second histogram shows the average support percentage for Donald Trump across various states.
+It highlights substantial support in states such as Florida and Texas, which align with historical trends of strong Republican support.
+Trump's average support appears robust in many midwestern and southern states, which are known for their conservative voter base.
+However, in more liberal-leaning states such as California and New York, the average support is lower, reflecting these states' tendency to lean Democratic.
+
+# Model
+
+The goal of our modeling strategy is twofold.
+Firstly, we aim to accurately predict the support percentage (PCT) for Harris and Trump based on relevant poll data and key influencing factors.
+Secondly, we seek to evaluate the efficacy of different modeling approaches—from simple linear regression (SLR) to multiple linear regression (MLR) and Bayesian hierarchical models—to understand their predictive capabilities and assess the underlying relationships between variables.
+By comparing these models, we can determine which approach provides the most robust and reliable predictions, while considering the variability and potential uncertainty in the data.
+
+## Model set-up
+
+The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:
+
+- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$
+
+- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$
+
+- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+us
+
+## Basic Model
+```{r}
+#| label: fig-model-pct-pollscore
+#| echo: false
+#| warning: false
+just_harris_data = Harris_data |> na.omit()
+lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+# summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+@fig-model-pct-pollscore in this SLR model, the response variable is pct and the only one predictor is the pollscore.
+The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line).
+This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct.
+The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
 
 ```{r}
+#| echo: false
+#| warning: false
+just_harris_data = Harris_data |> na.omit()
+model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
+# summary(model_MLR)
+just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
 
-clean_data <- read_csv("../data/02-analysis_data/cleaned_data.csv")
-head(clean_data)
+predictions = predict(model_MLR, just_harris_data)
 
-head(clean_data)
+ggplot(just_harris_data, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
+  facet_wrap(vars(methodology)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
 ```
 
-## Relationship Pollscore and Sample Size
+```{r}
+#| label: fig-model-MLR
+#| echo: false
+#| warning: false
+ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+@fig-model-MLR, in this MLR model, we predicts pct using pollscore, days_taken_from_election, methodology, sample_size, and state as predictors.
+The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
+The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60).
+This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.
+
+## Bayesian Model
+```{r}
+#| label: fig-bayesian-model
+#| echo: false
+#| warning: false
+baye_model_data = just_harris_data
+baye_model_data$pct = as.factor(baye_model_data$pct)
+baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
+baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
+baye_model_data$methodology = as.factor(baye_model_data$methodology)
+baye_model_data$pct <- as.numeric(baye_model_data$pct)
+# Define the Bayesian model with brms
+# test and train dataset
+
+formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)
+
+priors = normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = formula,
+  data = just_harris_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+pp_check(bayesian_model_1)
+# summary(bayesian_model_1)
+```
+
+### Posterior Predictive Checks 
+@fig-bayesian-model This plot displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.
+
+The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
+The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
+However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
+
+The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
+Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
+Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.
+
+### Train Test Validation
 
 ```{r}
+#| warning: false
+# Split data into training and test sets
+set.seed(123)
+train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
+train_data <- just_harris_data[train_indices, ]
+test_data <- just_harris_data[-train_indices, ]
 
-# removing outliers and na's
+# Fit the model on the training data
+bayesian_model_train <- stan_glmer(
+  formula = formula,
+  data = train_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
 
-graph_data <- clean_data %>% 
-  filter(!is.na(pollscore) & !is.na(sample_size))
+# Predict on the test data and check performance
+predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
+```
 
-outlier_top <- (quantile(graph_data$sample_size, 0.75) + 
-                  IQR(graph_data$sample_size)*1.5)
+```{r}
+#| warning: false
+ss_total <- sum((test_data$pct - mean(test_data$pct))^2)
+ss_residual <- sum((test_data$pct - colMeans(predictions))^2)
+r_squared <- 1 - (ss_residual / ss_total)
+print(paste("R-squared:", r_squared))
+# "R-squared: 0.700563202695799"
+```
 
-outlier_bottom <- (quantile(graph_data$sample_size, 0.25) - 
-                  IQR(graph_data$sample_size)*1.5)
+\begin{align}
+SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
+SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
+R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
+\text{R-squared} = 0.700563202695799
+\end{align}
 
-graph_data <- clean_data %>%
-  filter(sample_size < outlier_top & sample_size > outlier_bottom)
+Where:\
+\begin{itemize}
+    \item \( y_i \) represents the actual values,
+    \item \( \bar{y} \) is the mean of the actual values,
+    \item \( \hat{y}_i \) are the predicted values,
+    \item \( SS_{\text{total}} \) is the total sum of squares,
+    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
+\end{itemize}
 
-graph_data %>% filter(!is.na(pollscore) & !is.na(sample_size)) %>% 
-  ggplot(aes(x = pollscore, y = log(sample_size))) +
-  geom_point() + geom_smooth(method = "lm", se = FALSE, color = "red")
+we could see that the R squared value equal to the 0.700563202695799 suggesting that the model captures a substantial portion of the variability in the data and demonstrates that the model performs well in explaining the variability of the pct in the test data, reflecting strong predictive capabilities.
 
-# this model is really here just for explain, this realtionship should not
-# be treated as lienar
-lm_score_sample <- lm(log(sample_size) ~ pollscore, data = graph_data)
-summary(lm_score_sample)
+\newpage
+```{r}
+#| label: fig-bayesian-model-actual-predict
+#| echo: false
+#| warning: false
+plot(test_data$pct, colMeans(predictions), 
+     xlab = "Actual PCT Values", 
+     ylab = "Predicted  PCT Values", 
+     main = "Predicted PCT vs Actual PCT")
+abline(0, 1, col = "red")  # Add a 45-degree reference line
+```
+
+ by visual checking actual pct and predicted pct plot @fig-bayesian-model-actual-predict the points are plotted against a 45-degree line, which represents the ideal scenario where predicted values match the actual values perfectly.
+
+\newpage
+```{r}
+#| label: fig-bayesian-model-residual
+#| echo: false
+#| warning: false
+residuals <- test_data$pct - colMeans(predictions)
+plot(colMeans(predictions), residuals, 
+     xlab = "Predicted Values", 
+     ylab = "Residuals", 
+     main = "Residual Plot")
+abline(h = 0, col = "red")
+```
+
+and the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
+
+In conculsion, the visual checks from the predicted vs.
+actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting.
+It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.
+
+### Model justification
+
+To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
+The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity.
+The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables.
+The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts.
+This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability.
+By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.
+
+# Result
+```{r}
+#| label: fig-bayesian-model-CI
+#| echo: false
+#| warning: false
+# Plot random effects
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+```
+
+@fig-bayesian-model-CI This plot displays the estimated posterior distribution for the intercept parameter in the Bayesian model.
+The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S.
+election data used in the analysis.
+The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.
+
+The range of this credible interval spans from approximately 50 to 54 percent, suggesting that the model estimates the baseline level of support for Kamala Harris to be around this range.
+The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
+
+In conclusion, the estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
+election data analyzed, is centered around 52%, with a credible interval spanning approximately from 50% to 54%.
+This suggests that, according to the Bayesian model, the underlying support for Kamala Harris.
+
+```{r}
+#| echo: false
+#| warning: false
+
+predictions2 <- posterior_predict(bayesian_model_1, newdata = just_harris_data)
+
+predicted_means <- colMeans(predictions2)
+predicted_intervals <- apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+
+# Combine the results with the actual data
+result_summary <- data.frame(
+  Actual_PCT = just_harris_data$pct,
+  Predicted_PCT = predicted_means,
+  Lower_CI = predicted_intervals[1, ],
+  Upper_CI = predicted_intervals[2, ]
+)
+
+ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "U.S. Election Prediction Results",
+    x = "Actual PCT",
+    y = "Predicted PCT"
+  )
+
+print(head(result_summary))
 
@@
 
-## Relationship Between Pollscore and Transparency Score
+# Discussion
 
-```{r}
+## First discussion point {#sec-first-point}
 
-clean_data %>% filter(!is.na(pollscore) & !is.na(transparency_score)) %>% 
-  ggplot(aes(x = pollscore, y = transparency_score)) +
-  geom_point() + geom_smooth(method = "lm", se = FALSE, color = "red")
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
-# this model is really here just for explain, this realtionship should not
-# be treated as lienar
-lm_score_sample <- lm(pollscore ~ transparency_score, data = clean_data)
-summary(lm_score_sample)
+## Second discussion point
 
-```
+Please don't use these as sub-heading labels - change them to be what your point actually is.
 
-## Relationship Between Days Taken From Electon and Support Percentage
+## Third discussion point
 
-```{r}
-head(clean_data)
+## Weaknesses and next steps
 
-clean_data %>% 
-  filter(!is.na(days_taken_from_election) & !is.na(pct)) %>% 
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>% 
-  ggplot(aes(x = days_taken_from_election, y = pct)) +
-  geom_point() + 
-  geom_smooth(method = "lm", se = FALSE, color = "red") +
-  labs(x = "Days Before Election",
-       y = "Poll Percentage") + facet_wrap(~candidate_name)
+Weaknesses and next steps should also be included.
 
-```
+\newpage
 
-## Metholody And Support Percentage
+\appendix
 
-```{r}
+# Appendix {-}
 
-head(clean_data)
+# Additional data details
 
-clean_data %>% 
-  filter(!is.na(methodology) & !is.na(pct)) %>% 
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>% 
-  ggplot(aes(x = methodology, y = pct, fill = candidate_name)) +
-  geom_boxplot(position = position_dodge(width = 0.9)) +
-  labs(x = "Methodology", y = "Percentage", 
-       title = "Poll Percentages by Methodology and Candidate",
-       fill = "Candidate") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "top") +
-  scale_fill_brewer(palette = "Set1")
+# Model details {#sec-model-details}
 
-```
+## Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+## Diagnostics
+
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+
+# FiveThirtyEight Licenses
+[FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).
+
+\newpage
+
+# References

### 2024-11-03 paper/paper.qmd
first verision of data section and made updates to appendix and simulation

--- 2024-11-02 paper/paper.qmd
@@
 library(tidyverse)
-library(palmerpenguins)
 library(here)
@@
 library(knitr)
-library(arrow)
-library(rstanarm)
-library(brms)
-library(dplyr)
-library(lubridate)
+library(maps)
+library(mapproj)
+light_blue <- "#ADD8E6"
+dark_blue <- "#6688CC"
+light_red <- "#FF7777"
+dark_red <- "#B22222"
+purple <- "#9370DB"
 ```
@@
 As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
-However, the traditional poll aggregation does not reflect dynamics of an election, especially with differences in quality of polls.
-A more adaptable model is required to predict the election result based on both polling data and additional variables, such as sample size and pollster credibility.
+However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
+This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
 
-We build a hybrid election forecasting model following the strategies mentioned by @pasek.
+This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
 As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
@@
 The remainder of this paper is structured as follows: [update this …]
-
-The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here] and lubridate [@lubridate].
 
@@
 
-## Explore the data
-
 ## Measurement
 
-The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys.
-According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024].
-The dataset contains a list of questions in polls and surveys and their results.
-When a new poll is conducted, the poll is appended to the dataset.
-FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.
-
-In each of the polls recorded by FiveThirtyEight, all options of each question are recorded in candidate_names, while the proportion of respondents choosing that option is recorded in pct as a percentage.
-The pct is our primary response variable.
-
-There are limitations in these measurements.
-The differences in sampling methods, wording in survey questions and systematic biases are all reflected in the outcome [@radcliffe_538s_2023].
-Thus, in the dataset, FiveThirtyEight includes other variables, such as pollscore and sample_size for their prediction model in addition to the polling results for transparency and accuracy.
-As an example, pollscore indicates the pollster’s reliability rating, a low pollscore indicates a higher reliability rating [@silver_polls_2008].
-
-These variables combined allow researchers to reliably analyze and predict the 2024 US election result over time.
+In this dataset, each row represents a polling question that records the variables of interest.
+Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
+This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
 
@@
 #| warning: false
-cleaned_data = read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
 
@@
   kable(
-    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "candidate_name"),
+    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "Candidate Name"),
     booktabs = TRUE
@@
 
+### States included in anaylsis
+
+After the data cleaning process, 21 states had no polling data. A table showing the number of polls for each state, including those without any polls, is provided in @tbl-state-polls-count.
+
+This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
+The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary
+
 ## Basic Statistics Summary for Data
@@
 ```{r}
-#| label: fig-clean-data-histogram-Harris
-#| fig-cap: the average PCT vs State for Harris
+#| label: fig-avg-pct-state
+#| fig-cap: Historical State Voting Trends Are Maintined in 2024
 #| echo: false
 #| warning: false
-#| height: 40
+#| 
+cleaned_data <- cleaned_data %>% mutate(candidate_name =
+  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
 
-Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
-Harris_data |>
-  group_by(state) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = avg_pct, y = state)) + 
-  geom_col()  +
+cleaned_data %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
+  group_by(state, candidate_name) %>%
+  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
+  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
+  geom_col(position = "stack") +
   theme_minimal() +
   labs(
-    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
-  )
+    x = "Average Polled Support (%)", 
+    y = "State",
+    fill = "Candidate"
+  ) +
+  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
+  geom_vline(aes(xintercept = 50), color = purple, size = 1) +
+  theme_minimal() +
+  theme(legend.position = "bottom") + 
+  theme(axis.text.y = element_text(size = 8))
 
-# cleaned_data %>%
-#   mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
-#                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
-#   ggplot(mapping = aes(x = month, y = total_decedents)) + 
-#   geom_bar(stat = "identity") +
-#   theme_minimal() +
-#   labs(
-#     x = "Month", 
-#     y = "Total Decedents"
-#   )
 ```
+In figure @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump.
+Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.
 
 ```{r}
-#| label: fig-clean-data-histogram-trump
-#| fig-cap: the average PCT vs State for Trump
+#| label: fig-avg-pct-methodology
+#| fig-cap: Most Metholodgies Do Not Bias Towards a Candidate
 #| echo: false
 #| warning: false
-#| height: 40
+#| 
+cleaned_data <- cleaned_data %>% mutate(candidate_name =
+  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
 
-Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
-Trump_data |>
-  group_by(state) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = avg_pct, y = state)) + 
-  geom_col()  +
+cleaned_data %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
+  filter(!is.na(methodology)) %>% 
+  group_by(methodology, candidate_name) %>%
+  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
+  ggplot(mapping = aes(x = avg_pct, y = methodology, fill = candidate_name)) + 
+  geom_col(position = "stack") +
   theme_minimal() +
   labs(
-    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
-  )
+    x = "Average Polled Support (%)", 
+    y = "Methodology",
+    fill = "Candidate"
+  ) +
+  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
+  geom_vline(aes(xintercept = 50), color = purple, size = 1) +
+  theme_minimal() +
+  theme(legend.position = "bottom") + 
+  theme(axis.text.y = element_text(size = 8))
+
+```
+Figure @fig-avg-pct-methodology indicates minimal bias in polling percentages when comparing different polling methodologies, as most polls show an even distribution of support between candidates.
+However, polls that incorporate IVR (Interactive Voice Response) and online panels tend to exhibit greater variability, often resulting in more pronounced gaps in support between Donald Trump and Kamala Harris, depending on the specific poll.
+
+## Relationship Between Variables
+
+```{r}
+#| label: fig-pollscore-samplesize
+#| fig-cap: More Reliable Pollsters Have Larger Sample Sizes in their Polls
+#| echo: false
+#| warning: false
+
+# removing outliers and na's
+graph_data <- cleaned_data %>% 
+  filter(!is.na(pollscore) & !is.na(sample_size))
+
+outlier_top <- (quantile(graph_data$sample_size, 0.75) + 
+                  IQR(graph_data$sample_size)*1.5)
+
+outlier_bottom <- (quantile(graph_data$sample_size, 0.25) - 
+                  IQR(graph_data$sample_size)*1.5)
+
+graph_data <- cleaned_data %>%
+  filter(sample_size < outlier_top & sample_size > outlier_bottom)
+
+graph_data %>% filter(!is.na(pollscore) & !is.na(sample_size)) %>% 
+  ggplot(aes(x = pollscore, y = sample_size)) +
+  geom_point(alpha = 0.33) + geom_smooth(method = "lm", se = FALSE, color = light_red)
+
+# this model is really here just for explain, this realtionship should not
+# be treated as lienar
+cor(graph_data$sample_size, graph_data$pollscore)
+lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
+summary(lm_score_sample)
+
+```
+Figure @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
+The figure also shows that most polls have a sample size around 800 to 1200 participants.
+It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.
+
+```{r}
+#| label: fig-pollsamplesize-canidate-pct
+#| fig-cap: The Sample Size of a Poll does not impact a Candidate's Voting Percentage
+#| echo: false
+#| warning: false
+
+cleaned_data %>% 
+  filter(!is.na(sample_size) & !is.na(pct)) %>% 
+  ggplot(aes(x = sample_size, y = pct)) +
+  geom_point(alpha = 0.4) + 
+  geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Sample Size of the Poll",
+       y = "Candidate Percentage") + facet_wrap(~candidate_name)
+
+```
+Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.
+
+```{r}
+#| label: fig-pollscore-canidate-pct
+#| fig-cap: More Reliable Pollsters Score Higher For Trump
+#| echo: false
+#| warning: false
+
+cleaned_data %>% 
+  filter(!is.na(pollscore) & !is.na(pct)) %>% 
+  ggplot(aes(x = pollscore, y = pct)) +
+  geom_point(alpha = 0.4) + 
+  geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Pollscore",
+       y = "Candidate Percentage") + facet_wrap(~candidate_name)
+
+```
+Figure @fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage.
+For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support.
+Conversely, Kamala Harris shows the opposite trend: as the poll score decreases, her percentage tends to rise.
+This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.
+
+```{r}
+#| label: tbl-methodology-candidate-support
+#| fig-cap: pct vs methodology
+#| echo: false
+#| warning: false
+
+cleaned_data %>% 
+  filter(!is.na(methodology) & !is.na(pct)) %>% 
+  group_by(methodology, candidate_name) %>% 
+  summarise(methodology, candidate_name, n = n(), avg_pct = mean(pct), sd = sd(pct)) %>% 
+  arrange(methodology, candidate_name) %>% unique()
+
+```
+Based on @tbl-methodology-candidate-support, we might need to remove methodology from the model.
+
+```{r}
+#| label: tbl-state-candidate-support
+#| fig-cap: pct vs methodology
+#| echo: false
+#| warning: false
+
+cleaned_data %>% 
+  filter(!is.na(state) & !is.na(pct)) %>% 
+  group_by(state, candidate_name) %>% 
+  summarise(state, candidate_name, n = n(), avg_pct = mean(pct), sd = sd(pct)) %>% 
+  arrange(state, candidate_name) %>% unique()
 
@@
 
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-us
-
-## Basic Model
-```{r}
-just_harris_data = Harris_data |> na.omit()
-model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology, data = just_harris_data)
-summary(model_MLR)
-just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
-
-predictions = predict(model_MLR, just_harris_data)
-
-ggplot(just_harris_data, aes(x = end_date)) +
-  geom_point(aes(y = pct), color = "black") +
-  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
-  facet_wrap(vars(methodology)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
-```
+Based on @tbl-state-candidate-support, state might have to be a removed variable.
 
 ```{r}
-baye_model_data = just_harris_data
-baye_model_data$pct = as.factor(baye_model_data$pct)
-baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
-baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
-baye_model_data$methodology = as.factor(baye_model_data$methodology)
-baye_model_data$pct <- as.numeric(baye_model_data$pct)
-# Define the Bayesian model with brms
+#| label: fig-polldate-canidate-pct
+#| fig-cap: Both Canidate Percentages Are Gaining Support Closer to the Election
+#| echo: false
+#| warning: false
 
-formula <- pct ~ pollscore + days_taken_from_election + (1 | methodology) + (1 | state)
-
-priors = normal(0, 2.5, autoscale = TRUE)
-
-bayesian_model_1 <- stan_glmer(
-  formula = formula,
-  data = just_harris_data,
-  family = gaussian(),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
-pp_check(bayesian_model_1)
-summary(bayesian_model_1)
-
-# Plot random effects
-plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+cleaned_data %>% 
+  filter(!is.na(days_taken_from_election) & !is.na(pct)) %>% 
+  ggplot(aes(x = days_taken_from_election, y = pct)) +
+  geom_point(alpha = 0.4) + 
+  geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Timing of Polls Relative to Election Day (Days Prior to the Election)",
+       y = "Candidate Percentage") + facet_wrap(~candidate_name)
 
 ```
+Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election.
+This is a result of non-major candidate support rapidly decreasing.
+Specifically, Trump support is increasing at a faster rate than Harris.
+However, the polls in general show a slight lead for Harris throughout the last 100 days.
 
+The rapid decline in third party support could be due to Robert F. Kennedy dropping out of the race (this sentences should probably be in results or discussion section).
+
+\newpage
 # Model
@@
 
-# Methodology of YouGov
-YouGov's methodology documentations are seperated in two articles.
-The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
-
-## Population, Frame, and Sample
-As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
-The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
-Voter files indicates whether asomeone voted in a given election, thus YouGov's population covers all voters in previous US elections.
-
-YouGov’s sampling frame consists of its online panel members.
-These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
-YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
-
-YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
-YouGov interviews nearly 100,000 people in the first set of estimates.
-For the second set of estimates, YouGov didn't just start over with a new sample.
-They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
-
-## Sample Recruitment
-Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
-They must provide demographic details upon joining, which helps in selecting representative samples for each survey.
-When respondents complete a survey, they are awarded points that can be exchanged for money.
-
-## Sampling Approach and Trade-offs
-YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
-This method allows quick and cost-effective data collection.
-However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
-YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
-
-## Non-response Handling
-YouGov apply statistical weighting to adjust for the differences between the sample and target population.
-The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
-Additionally, quality control measures exclude unreliable responses to improve data accuracy.
-The respondents are offered a small incentive to decrease the non-response and increase participation.
-
-## Strengths and Weaknesses of the Questionnaire
-YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
-The pollster can recruit a large amount of panelists because of the online format.
-Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
-
-As a non-probability sample, it might miss certain demographic groups not covered by the online population.
-While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
-Additionaly, the categories in the surbey is oversimplified with bias.
-For instance, in the poll result published by YouGov, gender is divided into Male and Female.
-Race is divided into White, Black, Hispanic and Other.
-This indicates a lack of representation.
-
 # FiveThirtyEight Licenses
 [FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).
+
+# Overview of American Electon System
+
+# Brief Description of American Federal Goverment
+
+The American federal or national government is split into three branches, executive., legislative, and judical The executive branch includes the president and the military.
+The legislative branch include two subgroups, the House of Representatives and the Senate.
+These two subgroups create laws.
+Every state has two Senators and one Representative per approximately 750,000 people.
+The judicial branch is court system.
+
+## What is the Electoral College {#sec-electoral-college}
+
+The Electoral College is the system used in the United States to elect the president and vice president.
+Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives).
+When voters cast their ballots, they are actually voting for a slate of electors pledged to a candidate.
+The candidate who receives a majority of electoral votes (270 out of 538) wins the presidency.
+This system means that winning the popular vote in a state generally results in winning all of that state's electoral votes.
+The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.
+
+The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts.
+Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election.
+On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election.
+As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there.
+
+```{r}
+#| label: fig-currentstateofelection
+#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
+#| echo: false
+#| warning: false
+
+# Define your colors for each state (replace these colors with your desired ones)
+state_colors <- c(
+  alabama = dark_red,
+  arizona = purple,
+  arkansas = dark_red,
+  california = dark_blue,
+  colorado = dark_blue,
+  connecticut = dark_blue,
+  delaware = dark_blue,
+  `district of columbia` = dark_blue,
+  florida = light_blue,
+  georgia = purple,
+  idaho = dark_red,
+  illinois = dark_blue,
+  indiana = dark_red,
+  iowa = light_red,
+  kansas = dark_red,
+  kentucky = dark_red,
+  louisiana = dark_red,
+  maine = "gray",
+  maryland = dark_blue,
+  massachusetts = dark_blue,
+  michigan = purple,
+  minnesota = light_blue,
+  mississippi = dark_red,
+  missouri = dark_red,
+  montana = dark_red,
+  nebraska = "gray",
+  nevada = purple,
+  `new hampshire` = light_blue,
+  `new jersey` = dark_blue,
+  `new mexico` = dark_blue,
+  `new york` = dark_blue,
+  `north carolina` = purple,
+  `north dakota` = dark_red,
+  ohio = light_red,
+  oklahoma = dark_red,
+  oregon = dark_blue,
+  pennsylvania = purple,
+  `rhode island` = dark_blue,
+  `south carolina` = dark_red,
+  `south dakota` = dark_red,
+  tennessee = dark_red,
+  texas = light_red,
+  utah = dark_red,
+  vermont = dark_blue,
+  virginia = light_blue,
+  washington = dark_blue,
+  `west virginia` = dark_red,
+  wisconsin = purple,
+  wyoming = dark_red
+)
+
+# Load the states map data
+states_map <- map_data(map = "state")
+
+plotclr <- c(dark_red, light_blue, purple, gray, light_blue, dark_blue)
+legend_description <- c(
+  "Likely Republican",
+  "Lean Republican",
+  "Swing States",
+  "Lean Democrat",
+  "Likely Democrat",
+  "Split by District"
+)
+
+ggplot(states_map, aes(x = long, y = lat, group = group, fill = region)) +
+  geom_polygon(color = "white", size = 0.2) +
+  coord_map("albers", lat0 = 39, lat1 = 45) +
+  scale_fill_manual(values = state_colors, 
+                    breaks = plotclr,
+                    labels = legend_description,
+                    name = "Election Status") +  
+  guides(fill = guide_legend(override.aes = list(size = 5))) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        legend.title = element_text(size = 12, face = "bold"),
+        legend.text = element_text(size = 10),
+        axis.text = element_blank(),
+        axis.title = element_blank(),
+        panel.grid = element_blank())
+
+```
+### Legend
+
+```{r}
+#| label: tbl-electionmaplegend
+#| tbl-cap: Interpretation of Electoral College Map
+#| echo: false
+#| warning: false
+
+legend_tibble <- tibble(
+  Color = c("Dark Red", "Light Red", "Purple", "Light Blue", "Dark Blue", "Gray"),
+  Status = c("Republican", "Lean Republican", "Toss-Up", "Lean Democrat", "Democrat", "Split By District")
+)
+
+legend_tibble %>% kable(col.names = c("color", "status"))
+
+```
+The state statuses presented in this map are based on evaluations from @cnn, @foxnews, and @msnbc, which are generally agreed upon within the American political community.
+These organizations assessed historical voting patterns and recent polling data to derive their conclusions.
+
+Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district.
+In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic.
+In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican.
+Furthermore, Alaska and Hawaii are not in the map.
+Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.
+
+The seven generally agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.
+
+# States Poll Count
+
+```{r}
+#| label: tbl-state-polls-count
+#| tbl-cap: Polls included in Analysis Per State
+#| echo: false
+#| warning: false
+
+states <- c(state.name, "National")
+polls_per_state_tbl <- tibble(state = states)
+
+polls_per_state <- cleaned_data %>% group_by(state) %>% 
+  summarise(polls_count = n_distinct(poll_id))
+
+polls_per_state_tbl <- polls_per_state_tbl %>% 
+  left_join(polls_per_state, by = "state") %>% 
+  mutate(polls_count = if_else(!is.na(polls_count), polls_count, 0)) %>% 
+  select(state, polls_count)
+
+polls_per_state_tbl %>% kable(col.names = c("State", "Number of Polls In Analysis"))
+
+num_states_missing <- count(polls_per_state_tbl %>% 
+                              filter(polls_count == 0) %>% 
+                              select(state))
+
+```
 

### 2024-11-03 scripts/00-simulate_data.R

@@ -21,6 +21,8 @@ candidate_names <- c("Donald Trump", "Kamala Harris", "Jill Stein",
                     "Robert F. Kennedy")
 pollster <- c("Marist", "YouGov", "Ipsos", "Emerson", "SurveyUSA")
 
+methodologies <- c("IVR", "Email", "Online Panel", "Text", "Phone")
+
 # Reliably blue states
 blue_states <- c("California", "Colorado", "Connecticut", "Delaware", "Hawaii",
                  "Illinois", "Maine", "Maryland", "Massachusetts", "Minnesota",
@@ -41,6 +43,7 @@ swing_states <- c("Arizona", "Georgia", "Michigan", "Nevada",
 # Generate Poll IDs and States
 unique_poll_ids <- sample(10000:99999, length(states))
 poll_ids <- rep(unique_poll_ids, each = 4)
+methodology_column <- sample(methodologies, 4*length(states), replace = TRUE)
 state_column <- rep(states, each = 4)
 state_column <- rep(state_column, length.out = 4*length(states))
 sample_size <- sample(100:20000, length(states))
@@ -82,6 +85,7 @@ analysis_data <- tibble(
   pollster = pollster_column,
   poll_id = poll_ids,
   state = state_column,
+  methodology = methodology_column,
   sample_size = sample_size,
   candidate_name = rep(candidate_names, length.out = 204),
 )

### 2024-11-03 scripts/03-clean_data.R

@@ -39,7 +39,7 @@ cleaned_data <- cleaned_data %>%
   select(
     poll_id, pollster_id, pollster, question_id,
     sample_size, pollscore, methodology, days_taken_from_election,
-    end_date, start_date, state, answer, pct
+    end_date, start_date, state, candidate_name, pct
   )
 
 # Delete NA columns

### 2024-11-03 paper/paper.qmd
add a temp electoral college map

--- 2024-11-03 paper/paper.qmd
@@
 library(knitr)
+library(arrow)
+library(rstanarm)
 library(brms)
-library(rstanarm)
+library(dplyr)
+library(lubridate)
 library(gridExtra)
@@
 library(rstan)
-library(brms)
 
-```
-
-```{r}
-#| echo: false
-#| warning: false
-cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
+cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
 ```
@@
 As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
-However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
-This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+However, the traditional poll aggregation does not reflect dynamics of an election, especially with differences in quality of polls.
+A more adaptable model is required to predict the election result based on both polling data and additional variables, such as sample size and pollster credibility.
 
-This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
+We build a hybrid election forecasting model following the strategies mentioned by @pasek.
 As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
@@
 The remainder of this paper is structured as follows: [update this …]
+
+The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here], gridExtra [@gridextra], Matrix [@Matrix], Rstan [@Rstan] and lubridate [@lubridate].
 

### 2024-11-03 paper/paper.qmd
fixed electoral college map legend

--- 2024-11-03 paper/paper.qmd
@@
   scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
-  geom_vline(aes(xintercept = 50), color = purple, size = 1) +
   theme_minimal() +
@@
 Figure @fig-avg-pct-methodology indicates minimal bias in polling percentages when comparing different polling methodologies, as most polls show an even distribution of support between candidates.
-However, polls that incorporate IVR (Interactive Voice Response) and online panels tend to exhibit greater variability, often resulting in more pronounced gaps in support between Donald Trump and Kamala Harris, depending on the specific poll.
+However, polls only incorporate IVR (Interactive Voice Response) tend to skew towards Harris
 
@@
 # Define your colors for each state (replace these colors with your desired ones)
+state_colors_legend <- c(
+  "Republican" = dark_red,      # Dark Red
+  "Lean Republican" = light_red, # Light Red
+  "Toss-Up" = purple,         # Purple
+  "Lean Democrat" = light_blue,   # Light Blue
+  "Democrat" = dark_blue,        # Dark Blue
+  "Split By District" = "gray"
+)
+
 state_colors <- c(
@@
   nevada = purple,
-  `new hampshire` = light_blue,
+  `new hampshire` = dark_blue,
   `new jersey` = dark_blue,
@@
 
+state_colors <- setNames(state_colors, tolower(names(state_colors)))
+
+color_to_status <- setNames(names(state_colors_legend), state_colors_legend)
+
+states_map <- states_map %>%
+  mutate(status = color_to_status[state_colors[region]])
+
 plotclr <- c(dark_red, light_blue, purple, gray, light_blue, dark_blue)
@@
 
-ggplot(states_map, aes(x = long, y = lat, group = group, fill = region)) +
+ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
   geom_polygon(color = "white", size = 0.2) +
   coord_map("albers", lat0 = 39, lat1 = 45) +
-  scale_fill_manual(values = state_colors, 
-                    breaks = plotclr,
-                    labels = legend_description,
-                    name = "Election Status") +  
-  guides(fill = guide_legend(override.aes = list(size = 5))) +
+  scale_fill_manual(
+    values = state_colors_legend,
+    name = "Election Status"
+  ) +
   theme_minimal() +
-  theme(legend.position = "bottom",
+  theme(
+    legend.position = "bottom",
     legend.title = element_text(size = 12, face = "bold"),
@@
     axis.title = element_blank(),
-        panel.grid = element_blank())
-
-```
-### Legend
-
-```{r}
-#| label: tbl-electionmaplegend
-#| tbl-cap: Interpretation of Electoral College Map
-#| echo: false
-#| warning: false
-
-legend_tibble <- tibble(
-  Color = c("Dark Red", "Light Red", "Purple", "Light Blue", "Dark Blue", "Gray"),
-  Status = c("Republican", "Lean Republican", "Toss-Up", "Lean Democrat", "Democrat", "Split By District")
+    panel.grid = element_blank()
   )
-
-legend_tibble %>% kable(col.names = c("color", "status"))
 

### 2024-11-03 paper/paper.qmd
fixed graph axis labels

@@
   ggplot(aes(x = pollscore, y = sample_size)) +
-  geom_point(alpha = 0.33) + geom_smooth(method = "lm", se = FALSE, color = light_red)
+  geom_point(alpha = 0.33) + geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Poll Score", y = "Poll Sample Size")
 

### 2024-11-03 paper/paper.qmd
delete unnecessary data

--- 2024-11-03 paper/paper.qmd
@@
 library(rstan)
+
+light_blue <- "#ADD8E6"
+dark_blue <- "#6688CC"
+light_red <- "#FF7777"
+dark_red <- "#B22222"
+purple <- "#9370DB"
 

### 2024-11-03 scripts/00-simulate_data.R

@@ -11,6 +11,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 set.seed(853)
 
 
@@ -23,6 +24,10 @@ candidate_names <- c(
   "Robert F. Kennedy"
 )
 pollster <- c("Marist", "YouGov", "Ipsos", "Emerson", "SurveyUSA")
+
+# Methodology
+methodologies <- c("IVR", "Email", "Online Panel", "Text", "Phone")
+
 
 # Reliably blue states
 blue_states <- c(
@@ -47,11 +52,14 @@ swing_states <- c(
   "North Carolina", "Pennsylvania", "Wisconsin"
 )
 
+
+
 # Generate Poll IDs and States
 unique_poll_ids <- sample(10000:99999, length(states))
 unique_question_ids <- sample(10000:999999, length(states))
 poll_ids <- rep(unique_poll_ids, each = 4)
 question_ids <- rep(unique_question_ids, each = 4)
+methodology_column <- sample(methodologies, 4*length(states), replace = TRUE)
 state_column <- rep(states, each = 4)
 state_column <- rep(state_column, length.out = 4 * length(states))
 sample_size <- sample(100:20000, length(states))
@@ -93,6 +101,7 @@ analysis_data <- tibble(
   poll_id = poll_ids,
   question_ids = question_ids,
   state = state_column,
+  methodology = methodology_column,
   sample_size = sample_size,
   candidate_name = rep(candidate_names, length.out = 216),
 )

### 2024-11-03 scripts/01-test_simulated_data.R

@@ -35,9 +35,9 @@ test_that("Dataset row count is 216", {
 })
 
 # Check if the dataset has 7 columns
-test_that("Dataset column count is 7", {
-  expect_equal(ncol(analysis_data), 7,
-    info = "The dataset does not have 7 columns."
+test_that("Dataset column count is 8", {
+  expect_equal(ncol(analysis_data), 8,
+    info = "The dataset does not have 8 columns."
   )
 })
 
@@ -78,4 +78,5 @@ test_that("Character columns are of type character", {
   expect_type(analysis_data$pollster, "character")
   expect_type(analysis_data$state, "character")
   expect_type(analysis_data$candidate_name, "character")
+  expect_type(analysis_data$methodology, "character")
 })

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -71,4 +71,5 @@ test_that("Character columns are of type character", {
   expect_type(analysis_data$pollster, "character")
   expect_type(analysis_data$state, "character")
   expect_type(analysis_data$candidate_name, "character")
+  expect_type(analysis_data$methodology, "character")
 })

### 2024-11-03 paper/paper.qmd
changed metholodology chart to boxplot

--- 2024-11-03 paper/paper.qmd
@@
   filter(!is.na(methodology)) %>% 
-  group_by(methodology, candidate_name) %>%
-  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
-  ggplot(mapping = aes(x = avg_pct, y = methodology, fill = candidate_name)) + 
-  geom_col(position = "stack") +
+  ggplot(mapping = aes(x = methodology, y = pct)) + 
+  geom_boxplot(width = 0.6, position = position_dodge(width = 0.8), outlier.shape = NA) +
+  coord_flip() +
   theme_minimal() +
   labs(
-    x = "Average Polled Support (%)", 
-    y = "Methodology",
+    x = "Methodology",
+    y = "Polled Support (%)",
     fill = "Candidate"
@@
   scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
-  theme_minimal() +
   theme(legend.position = "bottom") + 
@@
 ```
-Figure @fig-avg-pct-methodology indicates minimal bias in polling percentages when comparing different polling methodologies, as most polls show an even distribution of support between candidates.
-However, polls only incorporate IVR (Interactive Voice Response) tend to skew towards Harris
+Figure @fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.
 
@@
 # be treated as lienar
-cor(graph_data$sample_size, graph_data$pollscore)
-lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
-summary(lm_score_sample)
-
+# cor(graph_data$sample_size, graph_data$pollscore)
+# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
+# summary(lm_score_sample)
 ```
@@
 This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.
-
-```{r}
-#| label: tbl-methodology-candidate-support
-#| fig-cap: pct vs methodology
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(methodology) & !is.na(pct)) %>% 
-  group_by(methodology, candidate_name) %>% 
-  summarise(methodology, candidate_name, n = n(), avg_pct = mean(pct), sd = sd(pct)) %>% 
-  arrange(methodology, candidate_name) %>% unique()
-
-```
-Based on @tbl-methodology-candidate-support, we might need to remove methodology from the model.
-
-```{r}
-#| label: tbl-state-candidate-support
-#| fig-cap: pct vs methodology
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(state) & !is.na(pct)) %>% 
-  group_by(state, candidate_name) %>% 
-  summarise(state, candidate_name, n = n(), avg_pct = mean(pct), sd = sd(pct)) %>% 
-  arrange(state, candidate_name) %>% unique()
-
-```
 
@@
   nevada = purple,
-  `new hampshire` = dark_blue,
+  `new hampshire` = light_blue,
   `new jersey` = dark_blue,
-  `new mexico` = dark_blue,
+  `new mexico` = light_blue,
   `new york` = dark_blue,

### 2024-11-03 paper/paper.qmd
added the idealized survey

@@
 #| label: fig-avg-pct-methodology
-#| fig-cap: Most Metholodgies Do Not Bias Towards a Candidate
+#| fig-cap: Polls with more Methodologies have Less Variability.
 #| echo: false
@@
 
-# Overview of American Electon System
+## Overview of American Electon System
 
-# Brief Description of American Federal Goverment
+## Brief Description of American Federal Goverment
 
@@
 
-## What is the Electoral College {#sec-electoral-college}
+### What is the Electoral College {#sec-electoral-college}
 
@@
   `district of columbia` = dark_blue,
-  florida = light_blue,
+  florida = light_red,
   georgia = purple,
@@
 
-# States Poll Count
+## States Poll Count
 
@@
 
+## The Ideal Survey
+
+### Objective
+
+The research team has developed a survey and distribution methodology with a hypothetical budget of $100,000.
+This objective is to create a polling system that accurately predicts the 2024 United States Election.
+
+### Sampling Frame
+
+Given the monetary constraint, the team's first decision is to mostly include swing states and districts.
+These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia.
+Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-PLACEHOLDER.
+However, the poll will also include Texas, Florida, and Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.
+Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida and Minnesota.
+
+### Survey Sending Process
+
+The team will then find and use multiple databases, such as @uspostalserviceresidentialaddressfile, to find addresses, telephones, and emails of potential voters in those states.
+According to @pewresearch, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums.
+Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email.
+Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.
+
+### Sampling Methodology
+
+Each state will have its own poll, but the polling methodology and the survey given will remain the same.
+Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes.
+Ideally, the stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for.
+However, these factors are almost impossible to determine while sending the survey.
+Therefore, the research team decided to ask demographic questions directly inside the survey.
+After data collection, the team will weigh data based on demographics.
+For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.
+
+Furthermore, the team will aim to sample approximately 1,000 people from each true swing state.
+This sample size is large enough to provide reliable conclusions but not so large that it resembles sampling with replacement.
+For true swing states, the team aims to survey 1,000 individuals.
+However, if the final sample size falls short of this target, it is not considered a significant issue.
+
+### Survey Implementation & Question Creation
+
+The team's ideal survey will be made using @qualtrics.
+It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants.
+To identify participants who are not engaged, the research team has decided to add a worthless question.
+This question is extremely simple and clearly has one correct answer.
+Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data.
+An example of this is question number 7.
+
+Order bias occurs when the sequence of questions subconsciously influences participants' responses.
+To prevent this, many questions should be randomized upon entry to the survey.
+However, some questions must follow a specific order, such as question 11, while others like questions 1-4 can be randomized.
+The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics paid plan.
+
+Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently.
+To prevent this, answer choices should be randomized upon survey entry.
+Questions 8 & 9 could benefit from this.
+Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics paid plan.
+
+A leading question occurs when a question is written in a way that suggests the user to give a certain answer.
+For example, "given that children are the future of our country, should we invest more money in their education".
+To prevent this, the researchers have ensured questions are written in a style where no unnecessary details or opinions are added.
+
+### Survey Questions
+
+Click this (link)[https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8] for the @qualtrics survey.
+
+List of all of the questions:
+
+1. Select your race(s) (racial options where chosen based on @whitehouseracialoptions-PLACEHOLDER)
+   - White
+   - Black or African American 
+   - American Indian or Alaska Native 
+   - Native Hawaiian or Pacific Islander 
+   - Middle Eastern or North African
+   - Asian 
+   - Other 
+   - Prefer not to answer
+
+2. Please select your gender 
+   - Male 
+   - Female 
+   - Non-binary 
+   - Prefer not to say
+
+3. Please enter your age (in numbers) 
+  - This is a text input field.
+Please note that this field has the auto-validation feature set to numbers in @qualtrics.
+As a result, participants can only input numbers in this field and are alert if they have not.
+
+4. Please select the highest degree of education you have obtained
+   - GED Certificate 
+   - High School Diploma 
+   - Undergraduate Degree 
+   - Graduate Degree (Masters/Phd) 
+   - None 
+   - Other
+   - Prefer not to answer
+
+5. Are you a registered voter for the 2024 United States Presidential Election? 
+   - Yes 
+   - No
+
+6. Place yourself on the political spectrum 
+   - Far Left 
+   - Center Left 
+   - Center 
+   - Center Right 
+   - Far Right
+
+7. Can pigs fly? 
+   - Yes 
+   - Maybe
+   - No 
+   - Prefer Not To Say
+
+8. What political party have you registered with? 
+   - Republicans 
+   - Democrats 
+   - Green Party 
+   - Libertarian 
+   - Other
+   - Independent (unregistered)
+
+9. Who will you vote for in the 2024 presidential election? 
+   - Democrat - Kamala Harris 
+   - Republican - Donald Trump 
+   - Green Party - Gill Stein 
+   - Libertarian - Chase Oliver 
+   - Other
+   - Will not vote
+
+### Potential Problems with the Methodology And Polls
+
+While the team's methodology and survey creates a robust system, there are potential issues.
+The reliance of weighting results based on a candidate's demographics can lead to error propagation.
+For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have a large impact on the polls prediction of what candidate will win the state.
+There is also a selection bias in terms of the monetary reward.
+Potentially, people who like monetary rewards could be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.
+
 \newpage

### 2024-11-03 paper/paper.qmd
fixed headings

@@
 
-## Overview of American Electon System
+# Overview of American Electon System
 
@@
 
-### What is the Electoral College {#sec-electoral-college}
+## What is the Electoral College {#sec-electoral-college}
 
@@
 
-## States Poll Count
+# States Poll Count
 
@@
 
-## The Ideal Survey
+# The Ideal Survey
 
-### Objective
+## Objective
 
@@
 
-### Sampling Frame
+## Sampling Frame
 
@@
 
-### Survey Sending Process
+## Survey Sending Process
 
@@
 
-### Sampling Methodology
+## Sampling Methodology
 
@@
 
-### Survey Implementation & Question Creation
+## Survey Implementation & Question Creation
 
@@
 
-### Survey Questions
+## Survey Questions
 
-Click this (link)[https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8] for the @qualtrics survey.
+Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics survey.
 
@@
 
-### Potential Problems with the Methodology And Polls
+## Potential Problems with the Methodology And Polls
 

### 2024-11-03 paper/paper.qmd
removed text
update paper, literature and renv

@@
 library(knitr)
-library(maps)
-library(mapproj)
+library(arrow)
+library(rstanarm)
+library(brms)
+library(dplyr)
+library(lubridate)
+library(gridExtra)
+library(Matrix)
+library(rstan)
+
 light_blue <- "#ADD8E6"
@@
 purple <- "#9370DB"
+
+cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
 ```
@@
 As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
-However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
-This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
+However, conventional aggregation models often overlook dynamic election factors, including the credibility and quality variations among poll sources.
 
-This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
+We build a hybrid election forecasting model following the strategies mentioned by @pasek.
 As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
-Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics.
-Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
+Our model incorporates this approach, filtering polls from @fivethirtyeight based on a numeric grade that indicates each pollster's reliability.
 
-In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model.
-We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
-presidential election.
+The objective of this study is to predict the support percentages for the primary candidates, Kamala Harris and Donald Trump, using this hybrid model.
+We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability.
+We also quantify other polling attributes, such as sample size, poll reliability scores, and polling methodology, which serve as predictors.
+Our results show that this model provides a more stable and accurate forecast than traditional poll aggregation alone.
+
+The remainder of this paper is structured as follows: @sec-data provides an overview and exploration of the data.
+@sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model.
+We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.
+
+The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here], gridExtra [@gridextra], Matrix [@Matrix], Rstan [@Rstan] and lubridate [@lubridate].
+
+## Estimand
 
@@
 The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.
-
-The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]
-
-The remainder of this paper is structured as follows: [update this …]
 
@@
 For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
-- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
+- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
 - **numeric predictor:**\
- sample size(sample_size: The total number of respondents participating in the poll) \
- timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
- pollscore(A numeric value representing the score or reliability of the pollster in question) \
+ `sample size` (sample_size: The total number of respondents participating in the poll) \
+ `days_taken_from_election` (the time gap between the poll start date and the real election date i.e days_taken_from_election = real US election date - poll start date) \
+ `pollscore` (A numeric value representing the score or reliability of the pollster in question) \
 - **categorical predictor**
-state(The U.S. state where the poll was conducted or focused) \
-methodology(The method used to conduct the poll)\
+`state` (The U.S. state where the poll was conducted or focused) \
+`methodology` (The method used to conduct the poll)\
+
+## Data Exploration
+
+```{r}
+#| label: fig-pairs
+#| fig-cap: "Pairs Plot of Key Numeric Variables Related to Percentage Support (pct)"
+#| echo: false
+#| warning: false
+# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
+numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
+
+# Create the pairs plot
+pairs(numeric_data)
+```
+
+In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
+The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
+The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
+No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
 
 ## Measurement
+
+The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys.
+According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024].
+The dataset contains a list of questions in polls and surveys and their results.
+When a new poll is conducted, the poll is appended to the dataset.
+FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.
+
+In each of the polls recorded by FiveThirtyEight, all options of each question are recorded in candidate_names, while the proportion of respondents choosing that option is recorded in pct as a percentage.
+The pct is our primary response variable.
+
+There are limitations in these measurements.
+The differences in sampling methods, wording in survey questions and systematic biases are all reflected in the outcome [@radcliffe_538s_2023].
+Thus, in the dataset, FiveThirtyEight includes other variables, such as pollscore and sample_size for their prediction model in addition to the polling results for transparency and accuracy.
+As an example, pollscore indicates the pollster’s reliability rating, a low pollscore indicates a higher reliability rating [@silver_polls_2008].
 
@@
 This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
+
+These variables combined allow researchers to reliably analyze and predict the 2024 US election result over time.
 
@@
 
+The first 6 rows of the dataset is displayed in @tbl-cleaned-data.
+
 ```{r}
@@
 #| warning: false
-cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
 
@@
   kable(
-    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "Candidate Name"),
+    col.names = c("pct", "Sample Size", "pollscore", "Days taken from election","state", "methodology", "Candidate name"),
     booktabs = TRUE
@@
 This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
-The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary
+The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.
 
@@
 ```
-In figure @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump.
+
+In @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump.
 Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.
@@
 ```
-Figure @fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.
+
+@fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.
 
@@
 ```
-Figure @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
+@fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
 The figure also shows that most polls have a sample size around 800 to 1200 participants.
@@
 ```{r}
-#| label: fig-pollscore-canidate-pct
+#| label: fig-pollscore-candidate-pct
 #| fig-cap: More Reliable Pollsters Score Higher For Trump
@@
 ```
-Figure @fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage.
+
+@fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage.
 For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support.
@@
 This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.
-
-Based on @tbl-state-candidate-support, state might have to be a removed variable.
 
@@
 ```
+
 Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election.
@@
 
-\newpage
-# Model
+# Model {#sec-model}
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modeling strategy is twofold.
+Firstly, we aim to accurately predict the support percentage (PCT) for Harris and Trump based on relevant poll data and key influencing factors.
+Secondly, we seek to evaluate the efficacy of different modeling approaches—from simple linear regression (SLR) to multiple linear regression (MLR) and Bayesian hierarchical models—to understand their predictive capabilities and assess the underlying relationships between variables.
+By comparing these models, we can determine which approach provides the most robust and reliable predictions, while considering the variability and potential uncertainty in the data.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+## Model set-up {#sec-modelsetup}
 
-## Model set-up
+The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$
+
+- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95
 
@@
 
+## Basic Model
+```{r}
+#| label: fig-model-pct-pollscore
+#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
+#| echo: false
+#| warning: false
+
+Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+just_harris_data = Harris_data |> na.omit()
+lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+# summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+We first set up a SLR model for predicting pct based on pollscore. The model is:
+
+$\hat{\text{pct}} = \beta_0 + \beta_1 \times \text{pollscore}$
+
+In this SLR model, the response variable is pct and the only one predictor is the pollscore.
+@fig-model-pct-pollscore visualizes the relationship between the actual and predicted values of percentage support (pct) for Kamala Harris, based on a simple linear regression model with pollscore as the sole predictor.
+Blue points represent individual comparisons between actual and predicted values.
+The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.
+
+The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line).
+This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct.
+The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
+
+```{r}
+#| label: fig-model-pct-pollscore-etc
+#| echo: false
+#| warning: false
+just_harris_data = Harris_data |> na.omit()
+model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
+# summary(model_MLR)
+just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
+
+predictions = predict(model_MLR, just_harris_data)
+```
+
+```{r}
+#| label: fig-model-MLR
+#| fig-cap: "Actual vs Predicted Percentage Support for Kamala Harris (Multiple Predictors)"
+#| echo: false
+#| warning: false
+ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+Next, we predicts pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:
+
+$\hat{\text{pct}} = \beta_0 + \beta_1 \cdot \text{pollscore} + \beta_2 \cdot \text{days\_taken\_from\_election} + \beta_3 \cdot \text{methodology} + \beta_4 \cdot \text{sample\_size} + \beta_5 \cdot \text{state}$
+
+@fig-model-MLR compares actual pct values with the predicted pct values from the MLR model.
+Blue points represent individual comparisons between actual and predicted values.
+The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.
+
+The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
+The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60).
+This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.
+
+## Bayesian Model
+
+Next, we set up a bayesian model as the following in accordance to @sec-modelsetup:
+
+\begin{align*}
+\text{pct}_i &\sim \mathcal{N}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_1 \cdot \text{pollscore}_i + \beta_2 \cdot \text{days\_taken\_from\_election}_i + \beta_3 \cdot \text{sample\_size}_i \\
++ u_{\text{methodology}[i]} + u_{\text{state}[i]} \\
+\alpha &\sim \text{Normal}(0, 2.5) \\
+\beta_j &\sim \text{Normal}(0, 2.5), \quad j = 1, 2, 3 \\
+u_{\text{methodology}} &\sim \text{Normal}(0, \sigma_{\text{methodology}}) \\
+u_{\text{state}} &\sim \text{Normal}(0, \sigma_{\text{state}})
+\end{align*}
+
+This hierarchical Bayesian model is designed to capture both the fixed effects of predictors and the random effects of grouping factors.
+The response variable is pct (percentage of support), predictors are pollscore, days taken from election, and sample size.
+A normal prior with mean 0 and standard deviation 2.5 is set for all coefficients and the intercept, scaled automatically.
+
+```{r}
+#| label: fig-bayesian-model
+#| fig-cap: "Posterior Predictive Check for the Bayesian Model Predicting pct"
+#| echo: false
+#| warning: false
+baye_model_data = just_harris_data
+baye_model_data$pct = as.factor(baye_model_data$pct)
+baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
+baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
+baye_model_data$methodology = as.factor(baye_model_data$methodology)
+baye_model_data$pct <- as.numeric(baye_model_data$pct)
+# Define the Bayesian model with brms
+# test and train dataset
+
+formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)
+
+priors = normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = formula,
+  data = just_harris_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+pp_check(bayesian_model_1)
+# summary(bayesian_model_1)
+```
+
+### Posterior Predictive Checks 
+@fig-bayesian-model displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.
+
+The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
+The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
+However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
+
+The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
+Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
+Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.
+
+### Train Test Validation
+we implemented a train-test split validation approach to evaluate the performance of a Bayesian hierarchical model predicting percentage support (pct) for Kamala Harris.
+The dataset was divided into training (70%) and test (30%) sets to fit and validate the model, respectively.
+We used the stan_glmer function to fit the model on the training data, with predictors including pollscore, days_taken_from_election, sample_size, and random intercepts for methodology and state.
+After training, we generated predictions on the test set and calculated an R squared value to quantify the model's predictive accuracy.
+```{r}
+#| warning: false
+#| echo: false
+# Split data into training and test sets
+set.seed(123)
+train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
+train_data <- just_harris_data[train_indices, ]
+test_data <- just_harris_data[-train_indices, ]
+
+# Fit the model on the training data
+bayesian_model_train <- stan_glmer(
+  formula = formula,
+  data = train_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
+# Predict on the test data and check performance
+predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
+```
+
+\begin{align}
+SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
+SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
+R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
+\text{R-squared} = 0.700563202695799
+\end{align}
+
+```{r}
+#| warning: false
+#| echo: false
+ss_total <- sum((test_data$pct - mean(test_data$pct))^2)
+ss_residual <- sum((test_data$pct - colMeans(predictions))^2)
+r_squared <- 1 - (ss_residual / ss_total)
+# "R-squared: 0.700563202695799"
+```
+
+Where:\
+\begin{itemize}
+    \item \( y_i \) represents the actual values,
+    \item \( \bar{y} \) is the mean of the actual values,
+    \item \( \hat{y}_i \) are the predicted values,
+    \item \( SS_{\text{total}} \) is the total sum of squares,
+    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
+\end{itemize}
+
+We could see that the R squared value equal to the 0.700563202695799 suggesting that the model captures a substantial portion of the variability in the data and demonstrates that the model performs well in explaining the variability of the pct in the test data, reflecting strong predictive capabilities.
+
+```{r}
+#| label: fig-bayesian-model-actual-predict
+#| fig-cap: "Predicted vs Actual Percentage Support (pct) for Kamala Harris in the Bayesian Model"
+#| echo: false
+#| warning: false
+plot(test_data$pct, colMeans(predictions), 
+     xlab = "Actual PCT Values", 
+     ylab = "Predicted  PCT Values", 
+     main = "Predicted PCT vs Actual PCT")
+abline(0, 1, col = "red")  # Add a 45-degree reference line
+```
+
+ By visual checking actual pct and predicted pct plot in @fig-bayesian-model-actual-predict, the points are plotted against a 45-degree line, which represents the ideal scenario where predicted values match the actual values perfectly.
+
+```{r}
+#| label: fig-bayesian-model-residual
+#| fig-cap: "Residual Plot for the Bayesian Hierarchical Model Predictions"
+#| echo: false
+#| warning: false
+residuals <- test_data$pct - colMeans(predictions)
+plot(colMeans(predictions), residuals, 
+     xlab = "Predicted Values", 
+     ylab = "Residuals", 
+     main = "Residual Plot")
+abline(h = 0, col = "red")
+```
+
+In addition, the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
+
+In conculsion, the visual checks from the predicted vs.
+actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting.
+It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.
+
 ### Model justification
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
+The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity.
+The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables.
+The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts.
+This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability.
+By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+# Result {#sec-result}
+```{r}
+#| label: fig-bayesian-model-CI
+#| echo: false
+#| warning: false
+# Plot random effects
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+```
 
-# Results
+@fig-bayesian-model-CI displays the estimated posterior distribution for the intercept parameter in the Bayesian model.
+The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S.
+election data used in the analysis.
+The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.
 
-Our results are summarized in @tbl-modelresults.
+The range of this credible interval spans from approximately 50 to 54 percent, suggesting that the model estimates the baseline level of support for Kamala Harris to be around this range.
+The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
 
-# Discussion
+In conclusion, the estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
+election data analyzed, is centered around 52%, with a credible interval spanning approximately from 50% to 54%.
+This suggests that, according to the Bayesian model, the underlying support for Kamala Harris.
 
-## First discussion point {#sec-first-point}
+```{r}
+#| echo: false
+#| warning: false
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+predictions2 <- posterior_predict(bayesian_model_1, newdata = just_harris_data)
 
-## Second discussion point
+predicted_means <- colMeans(predictions2)
+predicted_intervals <- apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+# Combine the results with the actual data
+result_summary <- data.frame(
+  Actual_PCT = just_harris_data$pct,
+  Predicted_PCT = predicted_means,
+  Lower_CI = predicted_intervals[1, ],
+  Upper_CI = predicted_intervals[2, ]
+)
 
-## Third discussion point
+ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "U.S. Election Prediction Results",
+    x = "Actual PCT",
+    y = "Predicted PCT"
+  )
 
-## Weaknesses and next steps
+print(head(result_summary))
 
-Weaknesses and next steps should also be included.
+```
+
+# Discussion {#sec-discussion}
+
+## Summary of Findings
+
+In this paper, we developed a hybrid election forecasting model to predict the 2024 U.S.
+presidential election results.
+Our model provides a stable and accurate forecast compared to traditional methods by combining poll-of-polls aggregation with additional factors such as poll score, sample size, days from the election, state, and methodology.
+
+Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability.
+The results shows that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.
+
+## Insights on Polling and Election Dynamics
+
+One of the main takeaways from this study is the impact of poll quality on election forecasts.
+We weight and filtered the polls by their reliability.
+This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.
+
+## Limitations of the Model
+
+While this model presents a more comprehensive approach, there are limitations.
+First, Any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data.
+For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.
+
+Second, the model only reflect current voting dynamics as it is based on historical data.
+This model does not account for short-term variability such as Biden's exit.
+This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.
+
+## Future Directions
+To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles.
+We includes previous year's polling dataset in this R project.
+By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics.
+This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion.
+Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, ultimately refining its accuracy and reliability for future applications in election forecasting.
 
@@
 # Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 ```
-The state statuses presented in this map are based on evaluations from @cnn, @foxnews, and @msnbc, which are generally agreed upon within the American political community.
+
+The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are generally agreed upon within the American political community.
 These organizations assessed historical voting patterns and recent polling data to derive their conclusions.
@@
 
-The team will then find and use multiple databases, such as @uspostalserviceresidentialaddressfile, to find addresses, telephones, and emails of potential voters in those states.
-According to @pewresearch, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums.
+The team will then find and use multiple databases, such as @usps_customer_nodate, to find addresses, telephones, and emails of potential voters in those states.
+According to @pew_research_center_us_nodate, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums.
 Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email.
@@
 
-The team's ideal survey will be made using @qualtrics.
+The team's ideal survey will be made using @qualtrics_qualtrics_nodate.
 It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants.
@@
 However, some questions must follow a specific order, such as question 11, while others like questions 1-4 can be randomized.
-The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics paid plan.
+The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics_qualtrics_nodate paid plan.
 
@@
 Questions 8 & 9 could benefit from this.
-Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics paid plan.
+Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics_qualtrics_nodate paid plan.
 
@@
 
-Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics survey.
+Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics_qualtrics_nodate survey.
 
@@
 
-1. Select your race(s) (racial options where chosen based on @whitehouseracialoptions-PLACEHOLDER)
+1. Select your race(s) (racial options where chosen based on @orvis_omb_2024)
    - White
@@
   - This is a text input field.
-Please note that this field has the auto-validation feature set to numbers in @qualtrics.
+Please note that this field has the auto-validation feature set to numbers in @qualtrics_qualtrics_nodate.
 As a result, participants can only input numbers in this field and are alert if they have not.
@@
 
+# Methodology of YouGov
+YouGov's methodology documentations are seperated in two articles.
+The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
+
+## Population, Frame, and Sample
+As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
+The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
+Voter files indicates whether someone voted in a given election, thus YouGov's population covers all voters in previous US elections.
+
+YouGov’s sampling frame consists of its online panel members.
+These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
+YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
+
+YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
+YouGov interviews nearly 100,000 people in the first set of estimates.
+For the second set of estimates, YouGov didn't just start over with a new sample.
+They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
+
+## Sample Recruitment
+Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
+They must provide demographic details upon joining, which helps in selecting representative samples for each survey.
+When respondents complete a survey, they are awarded points that can be exchanged for money.
+
+## Sampling Approach and Trade-offs
+YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
+This method allows quick and cost-effective data collection.
+However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
+YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
+
+## Non-response Handling
+YouGov apply statistical weighting to adjust for the differences between the sample and target population.
+The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
+Additionally, quality control measures exclude unreliable responses to improve data accuracy.
+The respondents are offered a small incentive to decrease the non-response and increase participation.
+
+## Strengths and Weaknesses of the Questionnaire
+YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
+The pollster can recruit a large amount of panelists because of the online format.
+Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
+
+As a non-probability sample, it might miss certain demographic groups not covered by the online population.
+While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
+Additionaly, the categories in the surbey is oversimplified with bias.
+For instance, in the poll result published by @yougov_say24_nodate, gender is divided into Male and Female.
+Race is divided into White, Black, Hispanic and Other.
+This indicates a lack of representation.
+
 \newpage

### 2024-11-04 paper/paper.qmd
update paper

@@
 ---
-title: "Prediction of 2024 US election ..."
+title: "Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality"
+subtitle: "A Hybrid Model Predicts a Narrow Lead for Kamala Harris in Swing States as Election Day Approaches"
 author: 
@@
 date-format: long
-abstract: "We forecast the winner of the 2024 US presidential election using “poll-of-polls” by building a linear model."
+abstract: "This paper develops a hybrid model to forecast the 2024 U.S.
+presidential election by combining poll aggregation techniques with additional variables, including pollster reliability, sample size, timing relative to the election, geographic region, and polling methodology.
+By integrating these factors, our model accounts for variations in poll quality and regional differences in voter sentiment, providing a more stable and accurate prediction compared to traditional poll aggregation alone.
+Our results show that Kamala Harris holds a slight lead over Donald Trump in most swing states, with support levels stabilizing closer to election day.
+This approach underscores the importance of nuanced forecasting models in election analysis, revealing how poll quality and timing can significantly influence support predictions.
+Ultimately, this model offers a comprehensive tool for understanding the dynamics of voter sentiment and improving the accuracy of election predictions."
 format: pdf
@@
 These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia.
-Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-PLACEHOLDER.
+Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college.
 However, the poll will also include Texas, Florida, and Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.

### 2024-11-04 scripts/00-simulate_data.R

@@ -59,7 +59,7 @@ unique_poll_ids <- sample(10000:99999, length(states))
 unique_question_ids <- sample(10000:999999, length(states))
 poll_ids <- rep(unique_poll_ids, each = 4)
 question_ids <- rep(unique_question_ids, each = 4)
-methodology_column <- sample(methodologies, 4*length(states), replace = TRUE)
+methodology_column <- sample(methodologies, 4 * length(states), replace = TRUE)
 state_column <- rep(states, each = 4)
 state_column <- rep(state_column, length.out = 4 * length(states))
 sample_size <- sample(100:20000, length(states))

### 2024-11-04 paper/paper.qmd
colin code update6

--- 2024-11-03 paper/paper.qmd
@@
 library(tidyverse)
+library(palmerpenguins)
 library(here)
@@
 library(knitr)
-library(maps)
-library(mapproj)
-light_blue <- "#ADD8E6"
-dark_blue <- "#6688CC"
-light_red <- "#FF7777"
-dark_red <- "#B22222"
-purple <- "#9370DB"
+library(brms)
+library(rstanarm)
+library(gridExtra)
+library(Matrix)
+library(rstan)
+library(brms)
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
 ```
@@
 For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
-- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
+- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
 - **numeric predictor:**\
- sample size(sample_size: The total number of respondents participating in the poll) \
- timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
- pollscore(A numeric value representing the score or reliability of the pollster in question) \
+ `sample size`(sample_size: The total number of respondents participating in the poll) \
+ `timegap`(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
+ `pollscore`(A numeric value representing the score or reliability of the pollster in question) \
 - **categorical predictor**
-state(The U.S. state where the poll was conducted or focused) \
-methodology(The method used to conduct the poll)\
+`state`(The U.S. state where the poll was conducted or focused) \
+`methodology`(The method used to conduct the poll)\
+
+## Data Exploration
+```{r}
+#| label: tbl-explore-data
+#| echo: false
+#| warning: false
+a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'sample size', y='pct', 
+       title = 'pct vs sample size')
+
+b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'pollscore', y='pct', 
+       title = 'pct vs pollscore')
+
+c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'methodology', y='pct', 
+       title = 'pct vs methodology')
+
+d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'max gust speed', y='pct', 
+       title = 'pct vs days taken from election')
+
+e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'state', y='pct', 
+       title = 'pct vs state')
+
+grid.arrange(a, b, c, d, e, nrow=2)
+```
+
+- pct vs sample size: This scatter plot shows the pct against the sample size, with a fitted trend line indicating a slight positive relationship.
+The data points are denser for lower sample sizes, suggesting that smaller sample sizes are more common in the dataset.
+
+- pct vs pollscore: This scatter plot illustrates the pct against pollscore.
+The fitted trend line suggests a weak negative relationship between pollscore and pct.
+The points are scattered without a strong linear pattern.
+
+- pct vs methodology: A boxplot comparing pct for different polling methodologies.
+The pct distribution varies across methodologies, with some showing greater spread or median differences.
+This suggests that the polling methodology may influence pct outcomes.
+
+- pct vs days taken from election: A scatter plot displaying pct versus the number of days before the election.
+The trend line indicates a slight negative relationship, suggesting that as the election date approaches, pct may decrease slightly.
+
+- pct vs state: A boxplot depicting pct across different states.
+The pct distribution varies by state, with some states showing wider variability or different median values, implying state-specific effects on pct.
+
+```{r}
+#| label: fig-pairs
+#| echo: false
+#| warning: false
+# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
+numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
+
+# Create the pairs plot
+pairs(numeric_data)
+```
+
+@fig-pairs The pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
+The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
+The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
+No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
 
@@
 ## Clean Data
-The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
+@tbl-cleaned-data The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
 First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable.
@@
   kable(
-    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "Candidate Name"),
+    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "candidate_name"),
     booktabs = TRUE
@@
 
-### States included in anaylsis
-
-After the data cleaning process, 21 states had no polling data. A table showing the number of polls for each state, including those without any polls, is provided in @tbl-state-polls-count.
-
-This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
-The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary
-
 ## Basic Statistics Summary for Data
@@
 ```{r}
-#| label: fig-avg-pct-state
-#| fig-cap: Historical State Voting Trends Are Maintined in 2024
+#| label: fig-clean-data-histogram-Harris-Trump
+#| fig-cap: the average PCT vs State for Harris and Trump
 #| echo: false
 #| warning: false
-#| 
-cleaned_data <- cleaned_data %>% mutate(candidate_name =
-  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
-
-cleaned_data %>%
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
-  group_by(state, candidate_name) %>%
-  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
-  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
-  geom_col(position = "stack") +
+#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
+#| layout-nrow: 2
+#| fig-width: 40
+#| fig-height: 20
+Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+Harris_data |>
+  group_by(state) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  ggplot(mapping = aes(x = state, y = avg_pct)) + 
+  geom_col()  +
   theme_minimal() +
   labs(
-    x = "Average Polled Support (%)", 
-    y = "State",
-    fill = "Candidate"
-  ) +
-  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
-  theme_minimal() +
-  theme(legend.position = "bottom") + 
-  theme(axis.text.y = element_text(size = 8))
+    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
+  )
 
-```
-In figure @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump.
-Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.
-
-```{r}
-#| label: fig-avg-pct-methodology
-#| fig-cap: Polls with more Methodologies have Less Variability.
-#| echo: false
-#| warning: false
-#| 
-cleaned_data <- cleaned_data %>% mutate(candidate_name =
-  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
-
-cleaned_data %>%
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
-  filter(!is.na(methodology)) %>% 
-  ggplot(mapping = aes(x = methodology, y = pct)) + 
-  geom_boxplot(width = 0.6, position = position_dodge(width = 0.8), outlier.shape = NA) +
-  coord_flip() +
+Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
+Trump_data |>
+  group_by(state) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  ggplot(mapping = aes(x = state, y = avg_pct)) + 
+  geom_col()  +
   theme_minimal() +
   labs(
-    x = "Methodology",
-    y = "Polled Support (%)",
-    fill = "Candidate"
-  ) +
-  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
-  theme(legend.position = "bottom") + 
-  theme(axis.text.y = element_text(size = 8))
+    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
+  )
 
+# cleaned_data %>%
+#   mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
+#                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
+#   ggplot(mapping = aes(x = month, y = total_decedents)) + 
+#   geom_bar(stat = "identity") +
+#   theme_minimal() +
+#   labs(
+#     x = "Month", 
+#     y = "Total Decedents"
+#   )
 ```
-Figure @fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.
-
-## Relationship Between Variables
-
-```{r}
-#| label: fig-pollscore-samplesize
-#| fig-cap: More Reliable Pollsters Have Larger Sample Sizes in their Polls
-#| echo: false
-#| warning: false
-
-# removing outliers and na's
-graph_data <- cleaned_data %>% 
-  filter(!is.na(pollscore) & !is.na(sample_size))
-
-outlier_top <- (quantile(graph_data$sample_size, 0.75) + 
-                  IQR(graph_data$sample_size)*1.5)
-
-outlier_bottom <- (quantile(graph_data$sample_size, 0.25) - 
-                  IQR(graph_data$sample_size)*1.5)
-
-graph_data <- cleaned_data %>%
-  filter(sample_size < outlier_top & sample_size > outlier_bottom)
-
-graph_data %>% filter(!is.na(pollscore) & !is.na(sample_size)) %>% 
-  ggplot(aes(x = pollscore, y = sample_size)) +
-  geom_point(alpha = 0.33) + geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Poll Score", y = "Poll Sample Size")
-
-# this model is really here just for explain, this realtionship should not
-# be treated as lienar
-# cor(graph_data$sample_size, graph_data$pollscore)
-# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
-# summary(lm_score_sample)
-```
-Figure @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
-The figure also shows that most polls have a sample size around 800 to 1200 participants.
-It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.
-
-```{r}
-#| label: fig-pollsamplesize-canidate-pct
-#| fig-cap: The Sample Size of a Poll does not impact a Candidate's Voting Percentage
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(sample_size) & !is.na(pct)) %>% 
-  ggplot(aes(x = sample_size, y = pct)) +
-  geom_point(alpha = 0.4) + 
-  geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Sample Size of the Poll",
-       y = "Candidate Percentage") + facet_wrap(~candidate_name)
-
-```
-Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.
-
-```{r}
-#| label: fig-pollscore-canidate-pct
-#| fig-cap: More Reliable Pollsters Score Higher For Trump
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(pollscore) & !is.na(pct)) %>% 
-  ggplot(aes(x = pollscore, y = pct)) +
-  geom_point(alpha = 0.4) + 
-  geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Pollscore",
-       y = "Candidate Percentage") + facet_wrap(~candidate_name)
-
-```
-Figure @fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage.
-For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support.
-Conversely, Kamala Harris shows the opposite trend: as the poll score decreases, her percentage tends to rise.
-This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.
-
-Based on @tbl-state-candidate-support, state might have to be a removed variable.
-
-```{r}
-#| label: fig-polldate-canidate-pct
-#| fig-cap: Both Canidate Percentages Are Gaining Support Closer to the Election
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(days_taken_from_election) & !is.na(pct)) %>% 
-  ggplot(aes(x = days_taken_from_election, y = pct)) +
-  geom_point(alpha = 0.4) + 
-  geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Timing of Polls Relative to Election Day (Days Prior to the Election)",
-       y = "Candidate Percentage") + facet_wrap(~candidate_name)
-
-```
-Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election.
-This is a result of non-major candidate support rapidly decreasing.
-Specifically, Trump support is increasing at a faster rate than Harris.
-However, the polls in general show a slight lead for Harris throughout the last 100 days.
-
-The rapid decline in third party support could be due to Robert F. Kennedy dropping out of the race (this sentences should probably be in results or discussion section).
 
 \newpage
+
+@fig-clean-data-histogram-Harris-Trump-1 The histogram displays the average support percentage for Kamala Harris across different U.S.
+states.
+The data indicates that support for Harris varies significantly across states.
+Notable observations include relatively high average support in states such as California and New York, which are known for being more Democratic-leaning.
+On the other hand, there are states with lower average support percentages, particularly in more traditionally Republican or swing states.
+The distribution suggests regional variations in support, with some states showing consistent backing for Harris while others indicate a weaker performance.
+
+@fig-clean-data-histogram-Harris-Trump-2 The second histogram shows the average support percentage for Donald Trump across various states.
+It highlights substantial support in states such as Florida and Texas, which align with historical trends of strong Republican support.
+Trump's average support appears robust in many midwestern and southern states, which are known for their conservative voter base.
+However, in more liberal-leaning states such as California and New York, the average support is lower, reflecting these states' tendency to lean Democratic.
+
 # Model
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of our modeling strategy is twofold.
+Firstly, we aim to accurately predict the support percentage (PCT) for Harris and Trump based on relevant poll data and key influencing factors.
+Secondly, we seek to evaluate the efficacy of different modeling approaches—from simple linear regression (SLR) to multiple linear regression (MLR) and Bayesian hierarchical models—to understand their predictive capabilities and assess the underlying relationships between variables.
+By comparing these models, we can determine which approach provides the most robust and reliable predictions, while considering the variability and potential uncertainty in the data.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$
+
+- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$
+
+- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95
 
@@
 
+## Basic Model
+```{r}
+#| label: fig-model-pct-pollscore
+#| echo: false
+#| warning: false
+just_harris_data = Harris_data |> na.omit()
+lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+# summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+@fig-model-pct-pollscore in this SLR model, the response variable is pct and the only one predictor is the pollscore.
+The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line).
+This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct.
+The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
+
+```{r}
+#| echo: false
+#| warning: false
+just_harris_data = Harris_data |> na.omit()
+model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
+# summary(model_MLR)
+just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
+
+predictions = predict(model_MLR, just_harris_data)
+
+ggplot(just_harris_data, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
+  facet_wrap(vars(methodology)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
+```
+
+```{r}
+#| label: fig-model-MLR
+#| echo: false
+#| warning: false
+ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+@fig-model-MLR, in this MLR model, we predicts pct using pollscore, days_taken_from_election, methodology, sample_size, and state as predictors.
+The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
+The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60).
+This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.
+
+## Bayesian Model
+```{r}
+#| label: fig-bayesian-model-first
+#| echo: false
+#| warning: false
+bayesian_model_1 = readRDS(here("models/bayesian_model_1.rds"))
+
+pp_check(bayesian_model_1)
+# summary(bayesian_model_1)
+```
+
+### Posterior Predictive Checks 
+@fig-bayesian-model-first This plot displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.
+
+The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
+The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
+However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
+
+The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
+Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
+Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.
+
+### Train Test Validation
+
+```{r}
+#| echo: false
+#| warning: false
+# Split data into training and test sets
+set.seed(123)
+# train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
+# train_data <- just_harris_data[train_indices, ]
+# test_data <- just_harris_data[-train_indices, ]
+train_data = read_csv(here("data/02-analysis_data/train_data.csv"))
+test_data = read_csv(here("data/02-analysis_data/test_data.csv"))
+
+# Fit the model on the training data
+bayesian_model_train = readRDS(here("models/bayesian_model_train.rds"))
+# bayesian_model_train <- stan_glmer(
+#   formula = formula,
+#   data = train_data,
+#   family = gaussian(),
+#   prior = priors,
+#   prior_intercept = priors,
+#   seed = 123,
+#   cores = 4,
+#   adapt_delta = 0.95
+# )
+
+# Predict on the test data and check performance
+predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+ss_total <- sum((test_data$pct - mean(test_data$pct))^2)
+ss_residual <- sum((test_data$pct - colMeans(predictions))^2)
+r_squared <- 1 - (ss_residual / ss_total)
+print(paste("R-squared:", r_squared))
+# "R-squared: 0.700563202695799"
+```
+
+\begin{align}
+SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
+SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
+R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
+\text{R-squared} = 0.700563202695799
+\end{align}
+
+Where:\
+\begin{itemize}
+    \item \( y_i \) represents the actual values,
+    \item \( \bar{y} \) is the mean of the actual values,
+    \item \( \hat{y}_i \) are the predicted values,
+    \item \( SS_{\text{total}} \) is the total sum of squares,
+    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
+\end{itemize}
+
+we could see that the R squared value equal to the 0.700563202695799 suggesting that the model captures a substantial portion of the variability in the data and demonstrates that the model performs well in explaining the variability of the pct in the test data, reflecting strong predictive capabilities.
+
+\newpage
+```{r}
+#| label: fig-bayesian-model-actual-predict
+#| echo: false
+#| warning: false
+plot(test_data$pct, colMeans(predictions), 
+     xlab = "Actual PCT Values", 
+     ylab = "Predicted  PCT Values", 
+     main = "Predicted PCT vs Actual PCT")
+abline(0, 1, col = "red")  # Add a 45-degree reference line
+```
+
+ by visual checking actual pct and predicted pct plot @fig-bayesian-model-actual-predict the points are plotted against a 45-degree line, which represents the ideal scenario where predicted values match the actual values perfectly.
+
+\newpage
+```{r}
+#| label: fig-bayesian-model-residual
+#| echo: false
+#| warning: false
+residuals <- test_data$pct - colMeans(predictions)
+plot(colMeans(predictions), residuals, 
+     xlab = "Predicted Values", 
+     ylab = "Residuals", 
+     main = "Residual Plot")
+abline(h = 0, col = "red")
+```
+
+and the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
+
+In conculsion, the visual checks from the predicted vs.
+actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting.
+It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.
+
 ### Model justification
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
+The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity.
+The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables.
+The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts.
+This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability.
+By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+# Result
+```{r}
+#| label: fig-bayesian-model-CI
+#| echo: false
+#| warning: false
+# Plot random effects
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+```
 
-# Results
+@fig-bayesian-model-CI This plot displays the estimated posterior distribution for the intercept parameter in the Bayesian model.
+The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S.
+election data used in the analysis.
+The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.
 
-Our results are summarized in @tbl-modelresults.
+The range of this credible interval spans from approximately 50 to 54 percent, suggesting that the model estimates the baseline level of support for Kamala Harris to be around this range.
+The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
+
+In conclusion, the estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
+election data analyzed, is centered around 52%, with a credible interval spanning approximately from 50% to 54%.
+This suggests that, according to the Bayesian model, the underlying support for Kamala Harris.
+
+```{r}
+#| label: tbl-bayesian-model-result-summary
+#| echo: false
+#| warning: false
+
+predictions2 = posterior_predict(bayesian_model_1, newdata = just_harris_data)
+
+predicted_means = colMeans(predictions2)
+predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+
+# Combine the results with the actual data
+result_summary = data.frame(
+  Actual_PCT = just_harris_data$pct,
+  Predicted_PCT = predicted_means,
+  Lower_CI = predicted_intervals[1, ],
+  Upper_CI = predicted_intervals[2, ]
+)
+
+kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")
+
+# print(head(result_summary))
+
+```
+
+```{r}
+#| label: fig-bayesian-model-result-CI
+#| echo: false
+#| warning: false
+ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "U.S. Election Prediction Results",
+    x = "Actual PCT",
+    y = "Predicted PCT"
+  )
+```
+
+@fig-bayesian-model-result-CI The model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages.
+The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges.
+This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes.
+@tbl-bayesian-model-result-summary The summary table also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.
 
@@
 
-# Overview of American Electon System
-
-## Brief Description of American Federal Goverment
-
-The American federal or national government is split into three branches, executive., legislative, and judical The executive branch includes the president and the military.
-The legislative branch include two subgroups, the House of Representatives and the Senate.
-These two subgroups create laws.
-Every state has two Senators and one Representative per approximately 750,000 people.
-The judicial branch is court system.
-
-## What is the Electoral College {#sec-electoral-college}
-
-The Electoral College is the system used in the United States to elect the president and vice president.
-Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives).
-When voters cast their ballots, they are actually voting for a slate of electors pledged to a candidate.
-The candidate who receives a majority of electoral votes (270 out of 538) wins the presidency.
-This system means that winning the popular vote in a state generally results in winning all of that state's electoral votes.
-The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.
-
-The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts.
-Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election.
-On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election.
-As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there.
-
-```{r}
-#| label: fig-currentstateofelection
-#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
-#| echo: false
-#| warning: false
-
-# Define your colors for each state (replace these colors with your desired ones)
-state_colors_legend <- c(
-  "Republican" = dark_red,      # Dark Red
-  "Lean Republican" = light_red, # Light Red
-  "Toss-Up" = purple,         # Purple
-  "Lean Democrat" = light_blue,   # Light Blue
-  "Democrat" = dark_blue,        # Dark Blue
-  "Split By District" = "gray"
-)
-
-state_colors <- c(
-  alabama = dark_red,
-  arizona = purple,
-  arkansas = dark_red,
-  california = dark_blue,
-  colorado = dark_blue,
-  connecticut = dark_blue,
-  delaware = dark_blue,
-  `district of columbia` = dark_blue,
-  florida = light_red,
-  georgia = purple,
-  idaho = dark_red,
-  illinois = dark_blue,
-  indiana = dark_red,
-  iowa = light_red,
-  kansas = dark_red,
-  kentucky = dark_red,
-  louisiana = dark_red,
-  maine = "gray",
-  maryland = dark_blue,
-  massachusetts = dark_blue,
-  michigan = purple,
-  minnesota = light_blue,
-  mississippi = dark_red,
-  missouri = dark_red,
-  montana = dark_red,
-  nebraska = "gray",
-  nevada = purple,
-  `new hampshire` = light_blue,
-  `new jersey` = dark_blue,
-  `new mexico` = light_blue,
-  `new york` = dark_blue,
-  `north carolina` = purple,
-  `north dakota` = dark_red,
-  ohio = light_red,
-  oklahoma = dark_red,
-  oregon = dark_blue,
-  pennsylvania = purple,
-  `rhode island` = dark_blue,
-  `south carolina` = dark_red,
-  `south dakota` = dark_red,
-  tennessee = dark_red,
-  texas = light_red,
-  utah = dark_red,
-  vermont = dark_blue,
-  virginia = light_blue,
-  washington = dark_blue,
-  `west virginia` = dark_red,
-  wisconsin = purple,
-  wyoming = dark_red
-)
-
-# Load the states map data
-states_map <- map_data(map = "state")
-
-state_colors <- setNames(state_colors, tolower(names(state_colors)))
-
-color_to_status <- setNames(names(state_colors_legend), state_colors_legend)
-
-states_map <- states_map %>%
-  mutate(status = color_to_status[state_colors[region]])
-
-plotclr <- c(dark_red, light_blue, purple, gray, light_blue, dark_blue)
-legend_description <- c(
-  "Likely Republican",
-  "Lean Republican",
-  "Swing States",
-  "Lean Democrat",
-  "Likely Democrat",
-  "Split by District"
-)
-
-ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
-  geom_polygon(color = "white", size = 0.2) +
-  coord_map("albers", lat0 = 39, lat1 = 45) +
-  scale_fill_manual(
-    values = state_colors_legend,
-    name = "Election Status"
-  ) +
-  theme_minimal() +
-  theme(
-    legend.position = "bottom",
-    legend.title = element_text(size = 12, face = "bold"),
-    legend.text = element_text(size = 10),
-    axis.text = element_blank(),
-    axis.title = element_blank(),
-    panel.grid = element_blank()
-  )
-
-```
-The state statuses presented in this map are based on evaluations from @cnn, @foxnews, and @msnbc, which are generally agreed upon within the American political community.
-These organizations assessed historical voting patterns and recent polling data to derive their conclusions.
-
-Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district.
-In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic.
-In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican.
-Furthermore, Alaska and Hawaii are not in the map.
-Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.
-
-The seven generally agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.
-
-# States Poll Count
-
-```{r}
-#| label: tbl-state-polls-count
-#| tbl-cap: Polls included in Analysis Per State
-#| echo: false
-#| warning: false
-
-states <- c(state.name, "National")
-polls_per_state_tbl <- tibble(state = states)
-
-polls_per_state <- cleaned_data %>% group_by(state) %>% 
-  summarise(polls_count = n_distinct(poll_id))
-
-polls_per_state_tbl <- polls_per_state_tbl %>% 
-  left_join(polls_per_state, by = "state") %>% 
-  mutate(polls_count = if_else(!is.na(polls_count), polls_count, 0)) %>% 
-  select(state, polls_count)
-
-polls_per_state_tbl %>% kable(col.names = c("State", "Number of Polls In Analysis"))
-
-num_states_missing <- count(polls_per_state_tbl %>% 
-                              filter(polls_count == 0) %>% 
-                              select(state))
-
-```
-
-# The Ideal Survey
-
-## Objective
-
-The research team has developed a survey and distribution methodology with a hypothetical budget of $100,000.
-This objective is to create a polling system that accurately predicts the 2024 United States Election.
-
-## Sampling Frame
-
-Given the monetary constraint, the team's first decision is to mostly include swing states and districts.
-These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia.
-Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-PLACEHOLDER.
-However, the poll will also include Texas, Florida, and Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.
-Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida and Minnesota.
-
-## Survey Sending Process
-
-The team will then find and use multiple databases, such as @uspostalserviceresidentialaddressfile, to find addresses, telephones, and emails of potential voters in those states.
-According to @pewresearch, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums.
-Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email.
-Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.
-
-## Sampling Methodology
-
-Each state will have its own poll, but the polling methodology and the survey given will remain the same.
-Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes.
-Ideally, the stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for.
-However, these factors are almost impossible to determine while sending the survey.
-Therefore, the research team decided to ask demographic questions directly inside the survey.
-After data collection, the team will weigh data based on demographics.
-For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.
-
-Furthermore, the team will aim to sample approximately 1,000 people from each true swing state.
-This sample size is large enough to provide reliable conclusions but not so large that it resembles sampling with replacement.
-For true swing states, the team aims to survey 1,000 individuals.
-However, if the final sample size falls short of this target, it is not considered a significant issue.
-
-## Survey Implementation & Question Creation
-
-The team's ideal survey will be made using @qualtrics.
-It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants.
-To identify participants who are not engaged, the research team has decided to add a worthless question.
-This question is extremely simple and clearly has one correct answer.
-Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data.
-An example of this is question number 7.
-
-Order bias occurs when the sequence of questions subconsciously influences participants' responses.
-To prevent this, many questions should be randomized upon entry to the survey.
-However, some questions must follow a specific order, such as question 11, while others like questions 1-4 can be randomized.
-The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics paid plan.
-
-Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently.
-To prevent this, answer choices should be randomized upon survey entry.
-Questions 8 & 9 could benefit from this.
-Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics paid plan.
-
-A leading question occurs when a question is written in a way that suggests the user to give a certain answer.
-For example, "given that children are the future of our country, should we invest more money in their education".
-To prevent this, the researchers have ensured questions are written in a style where no unnecessary details or opinions are added.
-
-## Survey Questions
-
-Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics survey.
-
-List of all of the questions:
-
-1. Select your race(s) (racial options where chosen based on @whitehouseracialoptions-PLACEHOLDER)
-   - White
-   - Black or African American 
-   - American Indian or Alaska Native 
-   - Native Hawaiian or Pacific Islander 
-   - Middle Eastern or North African
-   - Asian 
-   - Other 
-   - Prefer not to answer
-
-2. Please select your gender 
-   - Male 
-   - Female 
-   - Non-binary 
-   - Prefer not to say
-
-3. Please enter your age (in numbers) 
-  - This is a text input field.
-Please note that this field has the auto-validation feature set to numbers in @qualtrics.
-As a result, participants can only input numbers in this field and are alert if they have not.
-
-4. Please select the highest degree of education you have obtained
-   - GED Certificate 
-   - High School Diploma 
-   - Undergraduate Degree 
-   - Graduate Degree (Masters/Phd) 
-   - None 
-   - Other
-   - Prefer not to answer
-
-5. Are you a registered voter for the 2024 United States Presidential Election? 
-   - Yes 
-   - No
-
-6. Place yourself on the political spectrum 
-   - Far Left 
-   - Center Left 
-   - Center 
-   - Center Right 
-   - Far Right
-
-7. Can pigs fly? 
-   - Yes 
-   - Maybe
-   - No 
-   - Prefer Not To Say
-
-8. What political party have you registered with? 
-   - Republicans 
-   - Democrats 
-   - Green Party 
-   - Libertarian 
-   - Other
-   - Independent (unregistered)
-
-9. Who will you vote for in the 2024 presidential election? 
-   - Democrat - Kamala Harris 
-   - Republican - Donald Trump 
-   - Green Party - Gill Stein 
-   - Libertarian - Chase Oliver 
-   - Other
-   - Will not vote
-
-## Potential Problems with the Methodology And Polls
-
-While the team's methodology and survey creates a robust system, there are potential issues.
-The reliance of weighting results based on a candidate's demographics can lead to error propagation.
-For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have a large impact on the polls prediction of what candidate will win the state.
-There is also a selection bias in terms of the monetary reward.
-Potentially, people who like monetary rewards could be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.
-
 \newpage

### 2024-11-04 scripts/06-model_data.R

@@ -11,25 +11,76 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
-
+library(here)
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+# analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# ### Model data ####
+# first_model <-
+#   stan_glm(
+#     formula = flying_time ~ length + width,
+#     data = analysis_data,
+#     family = gaussian(),
+#     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     prior_aux = exponential(rate = 1, autoscale = TRUE),
+#     seed = 853
+#   )
 
 
-#### Save model ####
+# #### Save model ####
+# saveRDS(
+#   first_model,
+#   file = "models/first_model.rds"
+# )
+
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
+baye_model_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+# Define the Bayesian model with brms
+# test and train dataset
+
+formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)
+
+priors = normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = formula,
+  data = baye_model_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  bayesian_model_1,
+  file = "models/bayesian_model_1.rds"
+)
+
+
+just_harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
[…]

### 2024-11-04 paper/paper.qmd
update paper

--- 2024-11-04 paper/paper.qmd
@@
 ```
+
 @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
@@
 ```
+
 Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.

### 2024-11-04 temp/electoral_college_map_not_prediction.rmd

```{r}
#| label: fig-currentstateofelection
#| echo: false
#| warning: false

library(tidyverse)
library(tibble)
library(knitr)
library(maps)
library(mapproj)
light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"

state_colors_legend <- c(
  "Republican" = dark_red,      # Dark Red
  "Lean Republican" = light_red, # Light Red
  "Toss-Up" = purple,         # Purple
  "Lean Democrat" = light_blue,   # Light Blue
  "Democrat" = dark_blue,        # Dark Blue
  "Split By District" = "gray"
)

# Define your colors for each state (replace these colors with your desired ones)
state_colors <- c(
  alabama = dark_red,
  arizona = purple,
  arkansas = dark_red,
  california = dark_blue,
  colorado = dark_blue,
  connecticut = dark_blue,
  delaware = dark_blue,
  `district of columbia` = dark_blue,
  florida = light_blue,
  georgia = purple,
  idaho = dark_red,
  illinois = dark_blue,
  indiana = dark_red,
  iowa = light_red,
  kansas = dark_red,
  kentucky = dark_red,
  louisiana = dark_red,
  maine = "gray",
  maryland = dark_blue,
  massachusetts = dark_blue,
  michigan = purple,
  minnesota = light_blue,
  mississippi = dark_red,
  missouri = dark_red,
  montana = dark_red,
  nebraska = "gray",
  nevada = purple,
  `new hampshire` = dark_blue,
  `new jersey` = dark_blue,
  `new mexico` = dark_blue,
  `new york` = dark_blue,
  `north carolina` = purple,
  `north dakota` = dark_red,
  ohio = light_red,
  oklahoma = dark_red,
  oregon = dark_blue,
  pennsylvania = purple,
  `rhode island` = dark_blue,
  `south carolina` = dark_red,
  `south dakota` = dark_red,
  tennessee = dark_red,
  texas = light_red,
  utah = dark_red,
  vermont = dark_blue,
  virginia = light_blue,
  washington = dark_blue,
  `west virginia` = dark_red,
  wisconsin = purple,
  wyoming = dark_red
)

# Load the states map data
states_map <- map_data(map = "state")

state_colors <- setNames(state_colors, tolower(names(state_colors)))

color_to_status <- setNames(names(state_colors_legend), state_colors_legend)

states_map <- states_map %>%
  mutate(status = color_to_status[state_colors[region]])

plotclr <- c(dark_red, light_blue, purple, gray, light_blue, dark_blue)
legend_description <- c(
  "Likely Republican",
  "Lean Republican",
  "Swing States",
  "Lean Democrat",
  "Likely Democrat",
  "Split by District"
)

ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
  geom_polygon(color = "white", size = 0.2) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  scale_fill_manual(
    values = state_colors_legend,
    name = "Election Status"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

```
### Legend

```{r}
#| label: tbl-state-polls-count
#| tbl-cap: Interpretation of Electoral College Map
#| echo: false
#| warning: false

legend_tibble <- tibble(
  Color = c("Dark Red", "Light Red", "Purple", "Light Blue", "Dark Blue", "Gray"),
  Status = c("Republican", "Lean Republican", "Toss-Up", "Lean Democrat", "Democrat", "Split By District")
)

legend_tibble %>% kable(col.names = c("color", "status"))

```

### 2024-11-04 paper/paper.qmd
update model and paper

@@
 #| warning: false
-baye_model_data = just_harris_data
-baye_model_data$pct = as.factor(baye_model_data$pct)
-baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
-baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
-baye_model_data$methodology = as.factor(baye_model_data$methodology)
-baye_model_data$pct <- as.numeric(baye_model_data$pct)
-# Define the Bayesian model with brms
-# test and train dataset
-
-formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)
-
-priors = normal(0, 2.5, autoscale = TRUE)
-
-bayesian_model_1 <- stan_glmer(
-  formula = formula,
-  data = just_harris_data,
-  family = gaussian(),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
+bayesian_model_1 = readRDS(here("models/bayesian_model_1.rds"))
+
 pp_check(bayesian_model_1)
-# summary(bayesian_model_1)
 ```
@@
 ```{r}
-#| warning: false
-#| echo: false
+#| echo: false
+#| warning: false
 # Split data into training and test sets
 set.seed(123)
-train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
-train_data <- just_harris_data[train_indices, ]
-test_data <- just_harris_data[-train_indices, ]
+# train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
+# train_data <- just_harris_data[train_indices, ]
+# test_data <- just_harris_data[-train_indices, ]
+train_data = read_parquet(here("data/02-analysis_data/train_data.parquet"))
+test_data = read_parquet(here("data/02-analysis_data/test_data.parquet"))
 
 # Fit the model on the training data
-bayesian_model_train <- stan_glmer(
-  formula = formula,
-  data = train_data,
-  family = gaussian(),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
+bayesian_model_train = readRDS(here("models/bayesian_model_train.rds"))
+# bayesian_model_train <- stan_glmer(
+#   formula = formula,
+#   data = train_data,
+#   family = gaussian(),
+#   prior = priors,
+#   prior_intercept = priors,
+#   seed = 123,
+#   cores = 4,
+#   adapt_delta = 0.95
+# )
 
@@
 #| label: fig-bayesian-model-CI
+#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
 #| echo: false
@@
 ```{r}
-#| echo: false
-#| warning: false
-
-predictions2 <- posterior_predict(bayesian_model_1, newdata = just_harris_data)
-
-predicted_means <- colMeans(predictions2)
-predicted_intervals <- apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+#| label: tbl-bayesian-model-result-summary
+#| tbl-cap: "Bayesian Model Result Summary"
+#| echo: false
+#| warning: false
+
+predictions2 = posterior_predict(bayesian_model_1, newdata = just_harris_data)
+
+predicted_means = colMeans(predictions2)
+predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
 
 # Combine the results with the actual data
-result_summary <- data.frame(
+result_summary = data.frame(
   Actual_PCT = just_harris_data$pct,
@@
 
+kable(head(result_summary, 10))
+```
+
+```{r}
+#| label: fig-bayesian-model-result-CI
+#| fig-cap: "U.S. Election Prediction Results"
+#| echo: false
+#| warning: false
 ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
@@
   )
-
-print(head(result_summary))
-
-```
+```
+
+@fig-bayesian-model-result-CI shows that he model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages.
+The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges.
+This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes.
+@tbl-bayesian-model-result-summary also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.
 

### 2024-11-04 scripts/06-model_data.R

@@ -1,35 +1,64 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Bayesian Model for the paper
+# Author: Colin Sihan Yang
+# Date: 3 November 2024
+# Contact: lx.yu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
-
+library(arrow)
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+cleaned_data = read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
+baye_model_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+# Define the Bayesian model with brms
+# test and train dataset
 
+formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)
 
-#### Save model ####
+priors = normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = formula,
+  data = baye_model_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  bayesian_model_1,
+  file = "models/bayesian_model_1.rds"
+)
+
+
+just_harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
[…]

### 2024-11-04 scripts/06-model_data.R
update style

@@ -12,14 +12,14 @@ library(rstanarm)
 library(arrow)
 #### Read data ####
 
-cleaned_data = read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
-baye_model_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
+baye_model_data <- cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
 # Define the Bayesian model with brms
 # test and train dataset
 
 formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)
 
-priors = normal(0, 2.5, autoscale = TRUE)
+priors <- normal(0, 2.5, autoscale = TRUE)
 
 bayesian_model_1 <- stan_glmer(
   formula = formula,
@@ -38,7 +38,7 @@ saveRDS(
 )
 
 
-just_harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+just_harris_data <- cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
 train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
 train_data <- just_harris_data[train_indices, ]
 test_data <- just_harris_data[-train_indices, ]

### 2024-11-04 paper/paper.qmd
colin code update7

--- 2024-11-04 paper/paper.qmd
@@
 
+test_data_arizona = test_data[test_data$state == "Arizona", ]
 # Predict on the test data and check performance
 predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
+predictions_arizona = posterior_predict(bayesian_model_train, newdata = test_data_arizona)
+```
+
+```{r}
+# Calculate the average predicted PCT for Harris
+avg_predicted_pct_harris <- mean(predictions_arizona)
+
+# Print the result
+avg_predicted_pct_harris
+
 ```

### 2024-11-04 paper/paper.qmd
update paper
colin code update8

@@
 
-## Basic Model
+## Basic  SLR Model
 ```{r}
@@
 The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
+
+```{r}
+#| label: fig-model-pct-sample_size
+#| echo: false
+#| warning: false
+lm_model1 = lm(pct ~ sample_size, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+# summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+
+```
+
+@fig-model-pct-sample_size The distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor.
+This indicate that sample_size has limited impact on pct.
+
+```{r}
+# lm_model1 = lm(pct ~ sample_size + pollscore, data = just_harris_data)
+# predictions = predict(lm_model1, just_harris_data)
+# # summary(lm_model1)
+
+# ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+#   geom_point(color = 'blue') +
+#   geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+#   labs(title = "Actual vs Predicted Values",
+#        x = "Actual pct",
+#        y = "Predicted pct") +
+#   theme_minimal()
+```
+
+## MLR Model
 

### 2024-11-04 paper/paper.qmd
updated seed

--- 2024-11-04 paper/paper.qmd
@@
 
-## Overview
-
-For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
-- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
-- **numeric predictor:**\
- `sample size` (sample_size: The total number of respondents participating in the poll) \
- `days_taken_from_election` (the time gap between the poll start date and the real election date i.e days_taken_from_election = real US election date - poll start date) \
- `pollscore` (A numeric value representing the score or reliability of the pollster in question) \
-- **categorical predictor**
-`state` (The U.S. state where the poll was conducted or focused) \
-`methodology` (The method used to conduct the poll)\
-
-## Data Exploration
-
-```{r}
-#| label: fig-pairs
-#| fig-cap: "Pairs Plot of Key Numeric Variables Related to Percentage Support (pct)"
-#| echo: false
-#| warning: false
-# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
-numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
-
-# Create the pairs plot
-pairs(numeric_data)
-```
-
-In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
-The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
-The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
-No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
-
 ## Measurement
@@
 
+## Data Exploration
+
+The raw data from FiveThirtyEight contains 52 columns, all of the column headers are displayed below: 
+```{r}
+#| echo: false
+#| warning: false
+raw_data <- read_csv(here("data/01-raw_data/president_polls.csv"))
+
+# Get the column names of your data
+column_names <- colnames(raw_data)
+
+# Reshape the column names into a matrix with, for example, 4 columns
+num_cols <- 3
+column_matrix <- matrix(column_names, ncol = num_cols, byrow = TRUE)
+
+# Convert the matrix to a data frame for kable
+column_df <- as.data.frame(column_matrix)
+
+# Display in multiple columns with kable
+kable(column_df, format = "latex", booktabs = TRUE, col.names = NULL)
+```
+
+These columns can be categorized into three types, response variable, numeric predictors and categorical predictors.
+
+The response variable is:
+
+- pct: The percentage of support or vote share that each candidate (Kamala Harris or Donald Trump) received in the poll.
+
+Example of numeric predictors are:
+
+- sample_size: The total number of respondents in each poll.
+- pollscore: A numeric score representing the error and bias of the pollster. Negative numbers are better.
+- numeric_grade: A numeric grade assigned to each pollster, reflecting pollster quality or reliability.
+
+Example of categorical predictors are:
+
+- state: The U.S. state in which the poll was conducted or targeted.
+- methodology: The method used to conduct the poll (e.g., online, phone, in-person).
+- party: The political party of the candidate (e.g., DEM for Democrat, REP for Republican).
+- candidate_name: The name of the candidate, either Kamala Harris or Donald Trump.
+- pollster: The name of the polling organization that conducted the poll.
+- stage: The stage of the election (e.g., "general").
+
+<!-- ## Data Exploration
+
+```{r}
+#| label: fig-pairs
+#| fig-cap: "Pairs Plot of Key Numeric Variables Related to Percentage Support (pct)"
+#| echo: false
+#| warning: false
+# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
+numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
+
+# Create the pairs plot
+pairs(numeric_data)
+```
+
+In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
+The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
+The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
+No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
+-->
+
 ## Clean Data
@@
 Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.
+
+### States included in anaylsis
+
+After the data cleaning process, 21 states had no polling data. A table showing the number of polls for each state, including those without any polls, is provided in @tbl-state-polls-count.
+
+This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
+The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.
+
+## Cleaned data
 
@@
 
-### States included in anaylsis
-
-After the data cleaning process, 21 states had no polling data. A table showing the number of polls for each state, including those without any polls, is provided in @tbl-state-polls-count.
-
-This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
-The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.
-
-## Basic Statistics Summary for Data
+### Basic Statistics Summary for Cleaned Data
 
@@
 
-## Relationship Between Variables
+### Relationship Between Variables
 
@@
 
-## Basic Model
+## Basic SLR Model
 ```{r}
@@
 ```{r}
+#| label: fig-model-pct-sample_size
+#| fig-cap: "Actual vs. Predicted Values for pct Using sample_size as Predictor"
+#| echo: false
+#| warning: false
+lm_model1 = lm(pct ~ sample_size, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+# summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+
+```
+
+In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor.
+This indicate that sample_size has limited impact on pct.
+
+## MLR model
+```{r}
 #| label: fig-model-pct-pollscore-etc
@@
 predictions = predict(model_MLR, just_harris_data)
+
 ```

### 2024-11-04 scripts/06-model_data.R

@@ -11,7 +11,7 @@ library(tidyverse)
 library(rstanarm)
 library(arrow)
 #### Read data ####
-
+set.seed(123)
 cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
 baye_model_data <- cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
 # Define the Bayesian model with brms

### 2024-11-04 paper/paper.qmd
changed model script to create model for trump and harris

--- 2024-11-04 paper/paper.qmd
@@
 
-# Data {#sec-data}
-
-## Overview
-
-For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
-- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
-- **numeric predictor:**\
- `sample size`(sample_size: The total number of respondents participating in the poll) \
- `timegap`(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
- `pollscore`(A numeric value representing the score or reliability of the pollster in question) \
-- **categorical predictor**
-`state`(The U.S. state where the poll was conducted or focused) \
-`methodology`(The method used to conduct the poll)\
-
-## Data Exploration
-```{r}
-#| label: tbl-explore-data
-#| echo: false
-#| warning: false
-a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
-  geom_point() + 
-  geom_smooth(method = lm, se = FALSE) + 
-  labs(x = 'sample size', y='pct', 
-       title = 'pct vs sample size')
-
-b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
-  geom_point() + 
-  geom_smooth(method = lm, se = FALSE) + 
-  labs(x = 'pollscore', y='pct', 
-       title = 'pct vs pollscore')
-
-c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
-  geom_boxplot() + 
-  labs(x = 'methodology', y='pct', 
-       title = 'pct vs methodology')
-
-d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
-  geom_point() + 
-  geom_smooth(method = lm, se = FALSE) + 
-  labs(x = 'max gust speed', y='pct', 
-       title = 'pct vs days taken from election')
-
-e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
-  geom_boxplot() + 
-  labs(x = 'state', y='pct', 
-       title = 'pct vs state')
-
-grid.arrange(a, b, c, d, e, nrow=2)
-```
-
-- pct vs sample size: This scatter plot shows the pct against the sample size, with a fitted trend line indicating a slight positive relationship.
-The data points are denser for lower sample sizes, suggesting that smaller sample sizes are more common in the dataset.
-
-- pct vs pollscore: This scatter plot illustrates the pct against pollscore.
-The fitted trend line suggests a weak negative relationship between pollscore and pct.
-The points are scattered without a strong linear pattern.
-
-- pct vs methodology: A boxplot comparing pct for different polling methodologies.
-The pct distribution varies across methodologies, with some showing greater spread or median differences.
-This suggests that the polling methodology may influence pct outcomes.
-
-- pct vs days taken from election: A scatter plot displaying pct versus the number of days before the election.
-The trend line indicates a slight negative relationship, suggesting that as the election date approaches, pct may decrease slightly.
-
-- pct vs state: A boxplot depicting pct across different states.
-The pct distribution varies by state, with some states showing wider variability or different median values, implying state-specific effects on pct.
-
-```{r}
-#| label: fig-pairs
-#| echo: false
-#| warning: false
-# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
-numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
-
-# Create the pairs plot
-pairs(numeric_data)
-```
-
-@fig-pairs The pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
-The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
-The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
-No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
-
-## Measurement
-
-In this dataset, each row represents a polling question that records the variables of interest.
-Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
-This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
-
-## Clean Data
-@tbl-cleaned-data The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
-First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable.
-Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.
-
-We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S.
-election date.
-Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy.
-Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.
-
-```{r}
-#| label: tbl-cleaned-data
-#| tbl-cap: Sample of cleaned US election data 
-#| echo: false
-#| warning: false
-
-cleaned_data |>
-  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
-  head(6) |>
-  kable(
-    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "candidate_name"),
-    booktabs = TRUE
-    ) 
-```
-
-## Basic Statistics Summary for Data
-
-```{r}
-#| label: fig-clean-data-histogram-Harris-Trump
-#| fig-cap: the average PCT vs State for Harris and Trump
-#| echo: false
-#| warning: false
-#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
-#| layout-nrow: 2
-#| fig-width: 40
-#| fig-height: 20
-Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
-Harris_data |>
-  group_by(state) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = state, y = avg_pct)) + 
-  geom_col()  +
-  theme_minimal() +
-  labs(
-    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
-  )
-
-Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
-Trump_data |>
-  group_by(state) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  ggplot(mapping = aes(x = state, y = avg_pct)) + 
-  geom_col()  +
-  theme_minimal() +
-  labs(
-    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
-  )
-
-# cleaned_data %>%
-#   mutate(month = factor(month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
-#                                           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>%
-#   ggplot(mapping = aes(x = month, y = total_decedents)) + 
-#   geom_bar(stat = "identity") +
-#   theme_minimal() +
-#   labs(
-#     x = "Month", 
-#     y = "Total Decedents"
-#   )
-```
-
-\newpage
-
-@fig-clean-data-histogram-Harris-Trump-1 The histogram displays the average support percentage for Kamala Harris across different U.S.
-states.
-The data indicates that support for Harris varies significantly across states.
-Notable observations include relatively high average support in states such as California and New York, which are known for being more Democratic-leaning.
-On the other hand, there are states with lower average support percentages, particularly in more traditionally Republican or swing states.
-The distribution suggests regional variations in support, with some states showing consistent backing for Harris while others indicate a weaker performance.
-
-@fig-clean-data-histogram-Harris-Trump-2 The second histogram shows the average support percentage for Donald Trump across various states.
-It highlights substantial support in states such as Florida and Texas, which align with historical trends of strong Republican support.
-Trump's average support appears robust in many midwestern and southern states, which are known for their conservative voter base.
-However, in more liberal-leaning states such as California and New York, the average support is lower, reflecting these states' tendency to lean Democratic.
-
-# Model
-
-The goal of our modeling strategy is twofold.
-Firstly, we aim to accurately predict the support percentage (PCT) for Harris and Trump based on relevant poll data and key influencing factors.
-Secondly, we seek to evaluate the efficacy of different modeling approaches—from simple linear regression (SLR) to multiple linear regression (MLR) and Bayesian hierarchical models—to understand their predictive capabilities and assess the underlying relationships between variables.
-By comparing these models, we can determine which approach provides the most robust and reliable predictions, while considering the variability and potential uncertainty in the data.
-
-## Model set-up
-
-The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:
-
-- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$
-
-- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$
-
-- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-us
-
-## Basic Model
-```{r}
-#| label: fig-model-pct-pollscore
-#| echo: false
-#| warning: false
-just_harris_data = Harris_data |> na.omit()
-lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
-predictions = predict(lm_model1, just_harris_data)
-# summary(lm_model1)
-
-ggplot(just_harris_data, aes(x = pct, y = predictions)) +
-  geom_point(color = 'blue') +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
-  theme_minimal()
-```
-
-@fig-model-pct-pollscore in this SLR model, the response variable is pct and the only one predictor is the pollscore.
-The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line).
-This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct.
-The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
-
-```{r}
-#| echo: false
-#| warning: false
-just_harris_data = Harris_data |> na.omit()
-model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
-# summary(model_MLR)
-just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
-
-predictions = predict(model_MLR, just_harris_data)
-
-ggplot(just_harris_data, aes(x = end_date)) +
-  geom_point(aes(y = pct), color = "black") +
-  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
-  facet_wrap(vars(methodology)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
-```
-
-```{r}
-#| label: fig-model-MLR
-#| echo: false
-#| warning: false
-ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
-  geom_point(color = 'blue') +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
-  theme_minimal()
-```
-
-@fig-model-MLR, in this MLR model, we predicts pct using pollscore, days_taken_from_election, methodology, sample_size, and state as predictors.
-The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
-The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60).
-This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.
-
-## Bayesian Model
-```{r}
-#| label: fig-bayesian-model-first
-#| echo: false
-#| warning: false
-bayesian_model_1 = readRDS(here("models/bayesian_model_1.rds"))
-
-pp_check(bayesian_model_1)
-# summary(bayesian_model_1)
-```
-
-### Posterior Predictive Checks 
-@fig-bayesian-model-first This plot displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.
-
-The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
-The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
-However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
-
-The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
-Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
-Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.
-
-### Train Test Validation
-
-```{r}
-#| echo: false
-#| warning: false
-# Split data into training and test sets
-set.seed(123)
-# train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
-# train_data <- just_harris_data[train_indices, ]
-# test_data <- just_harris_data[-train_indices, ]
-train_data = read_csv(here("data/02-analysis_data/train_data.csv"))
-test_data = read_csv(here("data/02-analysis_data/test_data.csv"))
-
-# Fit the model on the training data
-bayesian_model_train = readRDS(here("models/bayesian_model_train.rds"))
-# bayesian_model_train <- stan_glmer(
-#   formula = formula,
-#   data = train_data,
-#   family = gaussian(),
-#   prior = priors,
-#   prior_intercept = priors,
-#   seed = 123,
-#   cores = 4,
-#   adapt_delta = 0.95
-# )
-
-test_data_arizona = test_data[test_data$state == "Arizona", ]
-# Predict on the test data and check performance
-predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
-predictions_arizona = posterior_predict(bayesian_model_train, newdata = test_data_arizona)
-```
-
-```{r}
-# Calculate the average predicted PCT for Harris
-avg_predicted_pct_harris <- mean(predictions_arizona)
-
-# Print the result
-avg_predicted_pct_harris
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-ss_total <- sum((test_data$pct - mean(test_data$pct))^2)
-ss_residual <- sum((test_data$pct - colMeans(predictions))^2)
-r_squared <- 1 - (ss_residual / ss_total)
-print(paste("R-squared:", r_squared))
-# "R-squared: 0.700563202695799"
-```
-
-\begin{align}
-SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
-SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
-R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
-\text{R-squared} = 0.700563202695799
-\end{align}
-
-Where:\
-\begin{itemize}
-    \item \( y_i \) represents the actual values,
-    \item \( \bar{y} \) is the mean of the actual values,
-    \item \( \hat{y}_i \) are the predicted values,
-    \item \( SS_{\text{total}} \) is the total sum of squares,
-    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
-\end{itemize}
-
-we could see that the R squared value equal to the 0.700563202695799 suggesting that the model captures a substantial portion of the variability in the data and demonstrates that the model performs well in explaining the variability of the pct in the test data, reflecting strong predictive capabilities.
-
-\newpage
-```{r}
-#| label: fig-bayesian-model-actual-predict
-#| echo: false
-#| warning: false
-plot(test_data$pct, colMeans(predictions), 
-     xlab = "Actual PCT Values", 
-     ylab = "Predicted  PCT Values", 
-     main = "Predicted PCT vs Actual PCT")
-abline(0, 1, col = "red")  # Add a 45-degree reference line
-```
-
- by visual checking actual pct and predicted pct plot @fig-bayesian-model-actual-predict the points are plotted against a 45-degree line, which represents the ideal scenario where predicted values match the actual values perfectly.
-
-\newpage
-```{r}
-#| label: fig-bayesian-model-residual
-#| echo: false
-#| warning: false
-residuals <- test_data$pct - colMeans(predictions)
-plot(colMeans(predictions), residuals, 
-     xlab = "Predicted Values", 
-     ylab = "Residuals", 
-     main = "Residual Plot")
-abline(h = 0, col = "red")
-```
-
-and the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
-
-In conculsion, the visual checks from the predicted vs.
-actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting.
-It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.
-
-### Model justification
-
-To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
-The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity.
-The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables.
-The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts.
-This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability.
-By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.
-
-# Result
-```{r}
-#| label: fig-bayesian-model-CI
-#| echo: false
-#| warning: false
-# Plot random effects
-plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
-```
-
-@fig-bayesian-model-CI This plot displays the estimated posterior distribution for the intercept parameter in the Bayesian model.
-The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S.
-election data used in the analysis.
-The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.
-
-The range of this credible interval spans from approximately 50 to 54 percent, suggesting that the model estimates the baseline level of support for Kamala Harris to be around this range.
-The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
-
-In conclusion, the estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
-election data analyzed, is centered around 52%, with a credible interval spanning approximately from 50% to 54%.
-This suggests that, according to the Bayesian model, the underlying support for Kamala Harris.
-
-```{r}
-#| label: tbl-bayesian-model-result-summary
-#| echo: false
-#| warning: false
-
-predictions2 = posterior_predict(bayesian_model_1, newdata = just_harris_data)
-
-predicted_means = colMeans(predictions2)
-predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
-
-# Combine the results with the actual data
-result_summary = data.frame(
-  Actual_PCT = just_harris_data$pct,
-  Predicted_PCT = predicted_means,
-  Lower_CI = predicted_intervals[1, ],
-  Upper_CI = predicted_intervals[2, ]
-)
-
-kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")
-
-# print(head(result_summary))
-
-```
-
-```{r}
-#| label: fig-bayesian-model-result-CI
-#| echo: false
-#| warning: false
-ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
-  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
-  labs(
-    title = "U.S. Election Prediction Results",
-    x = "Actual PCT",
-    y = "Predicted PCT"
-  )
-```
-
-@fig-bayesian-model-result-CI The model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages.
-The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges.
-This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes.
-@tbl-bayesian-model-result-summary The summary table also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
 # Appendix {-}
@@
 
+# Our Prediction
+
+```{r}
+#| echo: false
+#| warning: false
+# Split data into training and test sets
+
+set.seed(123)
+
+train_data_harris <- read_csv(here("data/02-analysis_data/train_data_harris.csv"))
+test_data_harris <- read_csv(here("data/02-analysis_data/test_data_harris.csv"))
+bayesian_model_train_harris <- readRDS(here("models/bayesian_model_train_harris.rds"))
+
+train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
+test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
+bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
+
+relevant_states <- c("Arizona", "Nevada", "Georgia", "Pennsylvania", "Michigan", "Minnesota",
+                     "Wisconsin", "Florida", "Texas", "Maine CD-2", "Nebraska CD-2", 
+                     "New Hampshire", "Ohio", "Virginia", "North Carolina", "Iowa")
+
+# Initialize an empty tibble to store results
+swing_state_predictions <- tibble(
+  state = character(),
+  harris_predicted_pct = numeric(),
+  trump_predicted_pct = numeric(),
+  winner = character()
+)
+
+for(state in relevant_states) {
+  #print(state, state %in% unique(test_data_trump), state %in% unique(test_data_harris))
+  test_data_state_harris <- test_data_harris[test_data_harris$state == state, ]
+  test_data_state_trump <- test_data_trump[test_data_trump$state == state, ]
+
+  predictions_state_harris <- posterior_predict(bayesian_model_train_harris, newdata = test_data_state_harris)
+  predictions_state_trump <- posterior_predict(bayesian_model_train_trump, newdata = test_data_state_trump)
+
+  avg_predicted_pct_state_harris <- mean(predictions_state_harris)
+  avg_predicted_pct_state_trump <- mean(predictions_state_trump)
+
+  # Add a new row to the results tibble
+  new_row <- tibble(
+    state = state,
+    harris_predicted_pct = avg_predicted_pct_state_harris,
+    trump_predicted_pct = avg_predicted_pct_state_trump,
+    winner = ifelse(trump_predicted_pct > harris_predicted_pct, "Trump", "Harris")
+  )
+  # print(state)
+  # print(avg_predicted_pct_state_harris)
+  # print(avg_predicted_pct_state_trump)
+  # 
+  swing_state_predictions <- bind_rows(swing_state_predictions, new_row)
+}
+
+swing_state_predictions <- swing_state_predictions %>% arrange(state) %>% unique()
+
+# View the results
+swing_state_predictions
+
+```
+
+```{r}
+#| label: fig-currentstateofelection
+#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
+#| echo: false
+#| warning: false
+
+# Define your colors for each state (replace these colors with your desired ones)
+state_colors_legend <- c(
+  "Republican" = "red",      # Dark Red
+  "Democrat" = "blue",        # Dark Blue
+  "Split By District" = "gray"
+)
+
+purple = "purple"
+light_blue = "green"
+light_red = "orange"
+
+state_colors <- c(
+  alabama = "red",
+  arizona = ifelse((swing_state_predictions %>% 
+                        filter(state == "Arizona") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  arkansas = "red",
+  california = "blue",
+  colorado = "blue",
+  connecticut = "blue",
+  delaware = "blue",
+  `district of columbia` = "blue",
+  florida = ifelse((swing_state_predictions %>% 
+                        filter(state == "Florida") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  georgia = ifelse((swing_state_predictions %>% 
+                        filter(state == "Georgia") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  idaho = "red",
+  illinois = "blue",
+  indiana = "red",
+  iowa = ifelse((swing_state_predictions %>% 
+                        filter(state == "Iowa") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  kansas = "red",
+  kentucky = "red",
+  louisiana = "red",
+  maine = "gray",
+  maryland = "blue",
+  massachusetts = "blue",
+  michigan = ifelse((swing_state_predictions %>% 
+                        filter(state == "Michigan") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  minnesota = ifelse((swing_state_predictions %>% 
+                        filter(state == "Minnesota") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  mississippi = "red",
+  missouri = "red",
+  montana = "red",
+  nebraska = "gray",
+  nevada = ifelse((swing_state_predictions %>% 
+                        filter(state == "Nevada") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  `new hampshire` = ifelse((swing_state_predictions %>% 
+                        filter(state == "New Hampshire") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  `new jersey` = "blue",
+  `new mexico` = "blue",
+  `new york` = "blue",
+  `north carolina` = ifelse((swing_state_predictions %>% 
+                        filter(state == "North Carolina") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  `north dakota` = "red",
+  ohio = ifelse((swing_state_predictions %>% 
+                        filter(state == "Ohio") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  oklahoma = "red",
+  oregon = "blue",
+  pennsylvania = ifelse((swing_state_predictions %>% 
+                        filter(state == "Pennsylvania") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  `rhode island` = "blue",
+  `south carolina` = "red",
+  `south dakota` = "red",
+  tennessee = "red",
+  texas = ifelse((swing_state_predictions %>% 
+                        filter(state == "Texas") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  utah = "red",
+  vermont = "blue",
+  virginia = ifelse((swing_state_predictions %>% 
+                        filter(state == "Virginia") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"),
+  washington = "blue",
+  `west virginia` = "red",
+  wisconsin = ifelse((swing_state_predictions %>% 
+                        filter(state == "Wisconsin") %>% 
+                        select(winner)) == "TRUMP", "red", "blue"), 
+  wyoming = "red")
+
+# Load the states map data
+states_map <- map_data(map = "state")
+
+state_colors <- setNames(state_colors, tolower(names(state_colors)))
+
+color_to_status <- setNames(names(state_colors_legend), state_colors_legend)
+
+states_map <- states_map %>%
+  mutate(status = color_to_status[state_colors[region]])
+
+plotclr <- c("red", gray, "blue")
+legend_description <- c(
+  "Republican",
+  "Democrat",
+  "Split by District"
+)
+
+ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
+  geom_polygon(color = "white", size = 0.2) +
+  coord_map("albers", lat0 = 39, lat1 = 45) +
+  scale_fill_manual(
+    values = state_colors_legend,
+    name = "Election Status"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    legend.title = element_text(size = 12, face = "bold"),
+    legend.text = element_text(size = 10),
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    panel.grid = element_blank()
+  )
+
+```
+
 \newpage

### 2024-11-04 scripts/06-model_data.R

@@ -34,6 +34,8 @@ library(here)
 #   file = "models/first_model.rds"
 # )
 
+set.seed(123)
+
 cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
 baye_model_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
 # Define the Bayesian model with brms
@@ -56,7 +58,7 @@ bayesian_model_1 <- stan_glmer(
 
 saveRDS(
   bayesian_model_1,
-  file = "models/bayesian_model_1.rds"
+  file = "models/bayesian_model_1_harris.rds"
 )
 
 
@@ -65,8 +67,8 @@ train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_
 train_data <- just_harris_data[train_indices, ]
 test_data <- just_harris_data[-train_indices, ]
 
-write.csv(train_data, "./data/02-analysis_data/train_data.csv")
-write.csv(test_data, "./data/02-analysis_data/test_data.csv")
+write.csv(train_data, "./data/02-analysis_data/train_data_harris.csv")
+write.csv(test_data, "./data/02-analysis_data/test_data_harris.csv")
 
 # Fit the model on the training data
 bayesian_model_train <- stan_glmer(
@@ -82,5 +84,60 @@ bayesian_model_train <- stan_glmer(
 
 saveRDS(
   bayesian_model_train,
-  file = "models/bayesian_model_train.rds"
+  file = "models/bayesian_model_train_harris.rds"
 )
+
+
+##### Trump Model Save
+
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
+baye_model_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
+# Define the Bayesian model with brms
+# test and train dataset
+
+formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)
+
+priors = normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = formula,
+  data = baye_model_data,
+  family = gaussian(),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
+saveRDS(
+  bayesian_model_1,
+  file = "models/bayesian_model_1_trump.rds"
+)
+
+
[…]

### 2024-11-04 paper/paper.qmd
fixed error in string

@@
                         filter(state == "Arizona") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   arkansas = "red",
@@
                         filter(state == "Florida") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   georgia = ifelse((swing_state_predictions %>% 
                         filter(state == "Georgia") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   idaho = "red",
@@
                         filter(state == "Iowa") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   kansas = "red",
@@
                         filter(state == "Michigan") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   minnesota = ifelse((swing_state_predictions %>% 
                         filter(state == "Minnesota") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   mississippi = "red",
@@
                         filter(state == "Nevada") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   `new hampshire` = ifelse((swing_state_predictions %>% 
                         filter(state == "New Hampshire") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   `new jersey` = "blue",
@@
                         filter(state == "North Carolina") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   `north dakota` = "red",
@@
                         filter(state == "Ohio") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   oklahoma = "red",
@@
                         filter(state == "Pennsylvania") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   `rhode island` = "blue",
@@
                         filter(state == "Texas") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   utah = "red",
@@
                         filter(state == "Virginia") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"),
+                        select(winner)) == "Trump", "red", "blue"),
   washington = "blue",
@@
                         filter(state == "Wisconsin") %>% 
-                        select(winner)) == "TRUMP", "red", "blue"), 
+                        select(winner)) == "Trump", "red", "blue"), 
   wyoming = "red")
@@
 ```
+Maine CD-2	47.40416	49.31394	Trump
+Nebraska CD-2	49.90928	42.17853	Harris
 

### 2024-11-04 paper/paper.qmd
added CI
added text description for prediction

@@
 
+Our prediction process consists of two primary components.
+First, we develop models for both Trump and Harris based on the variables outlined in Section @sec-PLACEHOLDER.
+This involves partitioning the dataset into training and testing subsets.
+Next, we further divide the testing dataset into swing states and other competitive races.
+We then input this test data into the respective models to generate predictions.
+By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state.
+The candidate with the higher percentage is deemed the winner for that state.
+
+We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa.
+Winners for other states were determined based on historical trends and predictions from sources like @cnn.
+Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
 ```{r}
@@
 # View the results
-swing_state_predictions
+swing_state_predictions %>% 
+  kable(col.names = 
+          c("State", "Harris Predicted Percentage", 
+            "Trump Predicted Percentage", "State Winner"))
 
@@
 #| label: fig-currentstateofelection
-#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
+#| fig-cap: Kamala Harris is Predicited to be the 47th President of the United States
 #| echo: false
@@
 ```
-Maine CD-2	47.40416	49.31394	Trump
-Nebraska CD-2	49.90928	42.17853	Harris
+According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States.
+There are also a few states with predictions not visible in the map, we will describe those predicts them below
+
+In Maine, Harris is projected to win the state’s overall delegates and District 1, while District 2 is expected to go to Trump.
+In Nebraska, Trump is expected to win the state’s delegates along with Districts 1 and 3, while Harris is predicted to win District 2.
+Additionally, Trump is projected to win Alaska, and Harris is expected to win Hawaii.
+
+Overall, the predictions indicate that Harris will receive 298 delegates, while Trump will receive 240 delegates.
+
+```{r}
+
+plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
+
+```
 

### 2024-11-04 paper/paper.qmd
added description of the Trump Model

@@
 
-# Our Prediction
+# Trump Voter Prediction Model
+
+The multiple linear regression model (MLR) for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris.
+Likewise, the Trump dataset is also split into training and testing data.
+The model outputs are below.
+
+```{r}
+#| echo: false
+#| warning: false
+set.seed(123)
+train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
+test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
+bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
+```
+
+```{r}
+#| label: fig-model-trumpmlr
+#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
+#| echo: false
+#| warning: false
+
+just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
+
+predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
+
+predicted_means = colMeans(predictions2)
+predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+
+result_summary = data.frame(
+  Actual_PCT = just_trump$pct,
+  Predicted_PCT = predicted_means,
+  Lower_CI = predicted_intervals[1, ],
+  Upper_CI = predicted_intervals[2, ]
+)
+
+# kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")
+
+ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
+  geom_point(color = "blue", alpha = 0.7) +
+  # geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "Actual vs Predicted Percentage For Trump MLR Model (All Trump Data)",
+    x = "Actual PCT",
+    y = "Predicted PCT"
+  )
+```
+From @fig-model-trumpmlr it is clear that the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line).
+Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
+```{r}
+#| label: fig-model-trumpmlr-res
+#| fig-cap: "MLR Trump Model does not Appear to Overfit"
+#| echo: false
+#| warning: false
+
+predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
+
+plot(test_data_trump$pct, colMeans(predictions2), 
+     xlab = "Actual PCT Values", 
+     ylab = "Predicted  PCT Values", 
+     main = "Predicted PCT vs Actual PCT (Test Data)")
+abline(0, 1, col = "red")  # Add a 45-degree reference line
+
+```
+Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
+This suggests that model can generalization to outside data.
+Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
+```{r}
+#| label: fig-trumpmodelci
+#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
+#| echo: false
+#| warning: false
+
+plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
+```
+Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% confidence interval ranges from around 42.5% to 47.2%.
+This confidence interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.
+
+# Election Prediction
 
@@
 #| warning: false
+#| label: tbl-electionprediction
+#| tbl-cap: Kamala Harris Wins Most of the Swing States
 # Split data into training and test sets
-
-set.seed(123)
 
@@
 bayesian_model_train_harris <- readRDS(here("models/bayesian_model_train_harris.rds"))
-
-train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
-test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
-bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
 
@@
 ```{r}
-#| label: fig-currentstateofelection
+#| label: fig-electionpredictionmap
 #| fig-cap: Kamala Harris is Predicited to be the 47th President of the United States
@@
 
-```{r}
-
-plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
-
-```
-
 \newpage

### 2024-11-04 paper/paper.qmd
update the paper

--- 2024-11-04 paper/paper.qmd
@@
 library(rstan)
-library(brms)
 
-```
+light_blue <- "#ADD8E6"
+dark_blue <- "#6688CC"
+light_red <- "#FF7777"
+dark_red <- "#B22222"
+purple <- "#9370DB"
 
-```{r}
-#| echo: false
-#| warning: false
-cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
+cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
 ```
@@
 
-The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]
+# Data {#sec-data}
 
-The remainder of this paper is structured as follows: [update this …]
+## Measurement
+
+The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys.
+According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024].
+The dataset contains a list of questions in polls and surveys and their results.
+When a new poll is conducted, the poll is appended to the dataset.
+FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.
+
+In each of the polls recorded by FiveThirtyEight, all options of each question are recorded in candidate_names, while the proportion of respondents choosing that option is recorded in pct as a percentage.
+The pct is our primary response variable.
+
+There are limitations in these measurements.
+The differences in sampling methods, wording in survey questions and systematic biases are all reflected in the outcome [@radcliffe_538s_2023].
+Thus, in the dataset, FiveThirtyEight includes other variables, such as pollscore and sample_size for their prediction model in addition to the polling results for transparency and accuracy.
+As an example, pollscore indicates the pollster’s reliability rating, a low pollscore indicates a higher reliability rating [@silver_polls_2008].
+
+In this dataset, each row represents a polling question that records the variables of interest.
+Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
+This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
+
+These variables combined allow researchers to reliably analyze and predict the 2024 US election result over time.
+
+## Data Exploration
+
+The raw data from FiveThirtyEight contains 52 columns, all of the column headers are displayed below: 
+```{r}
+#| echo: false
+#| warning: false
+raw_data <- read_csv(here("data/01-raw_data/president_polls.csv"))
+
+# Get the column names of your data
+column_names <- colnames(raw_data)
+
+# Reshape the column names into a matrix with, for example, 4 columns
+num_cols <- 3
+column_matrix <- matrix(column_names, ncol = num_cols, byrow = TRUE)
+
+# Convert the matrix to a data frame for kable
+column_df <- as.data.frame(column_matrix)
+
+# Display in multiple columns with kable
+kable(column_df, format = "latex", booktabs = TRUE, col.names = NULL)
+```
+
+These columns can be categorized into three types, response variable, numeric predictors and categorical predictors.
+
+The response variable is:
+
+- pct: The percentage of support or vote share that each candidate (Kamala Harris or Donald Trump) received in the poll.
+
+Example of numeric predictors are:
+
+- sample_size: The total number of respondents in each poll.
+- pollscore: A numeric score representing the error and bias of the pollster. Negative numbers are better.
+- numeric_grade: A numeric grade assigned to each pollster, reflecting pollster quality or reliability.
+
+Example of categorical predictors are:
+
+- state: The U.S. state in which the poll was conducted or targeted.
+- methodology: The method used to conduct the poll (e.g., online, phone, in-person).
+- party: The political party of the candidate (e.g., DEM for Democrat, REP for Republican).
+- candidate_name: The name of the candidate, either Kamala Harris or Donald Trump.
+- pollster: The name of the polling organization that conducted the poll.
+- stage: The stage of the election (e.g., "general").
+
+<!-- ## Data Exploration
+
+```{r}
+#| label: fig-pairs
+#| fig-cap: "Pairs Plot of Key Numeric Variables Related to Percentage Support (pct)"
+#| echo: false
+#| warning: false
+# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
+numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
+
+# Create the pairs plot
+pairs(numeric_data)
+```
+
+In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
+The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
+The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
+No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
+-->
+
+## Clean Data
+The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
+First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable.
+Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.
+
+We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S.
+election date.
+Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy.
+Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.
+
+### States included in anaylsis
+
+After the data cleaning process, 21 states had no polling data. A table showing the number of polls for each state, including those without any polls, is provided in @tbl-state-polls-count.
+
+This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
+The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.
+
+## Cleaned data
+
+The first 6 rows of the dataset is displayed in @tbl-cleaned-data.
+
+```{r}
+#| label: tbl-cleaned-data
+#| tbl-cap: Sample of cleaned US election data 
+#| echo: false
+#| warning: false
+
+cleaned_data |>
+  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
+  head(6) |>
+  kable(
+    col.names = c("pct", "Sample Size", "pollscore", "Days taken from election","state", "methodology", "Candidate name"),
+    booktabs = TRUE
+    ) 
+```
+
+### Basic Statistics Summary for Cleaned Data
+
+```{r}
+#| label: fig-avg-pct-state
+#| fig-cap: Historical State Voting Trends Are Maintined in 2024
+#| echo: false
+#| warning: false
+#| 
+cleaned_data <- cleaned_data %>% mutate(candidate_name =
+  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
+
+cleaned_data %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
+  group_by(state, candidate_name) %>%
+  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
+  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
+  geom_col(position = "stack") +
+  theme_minimal() +
+  labs(
+    x = "Average Polled Support (%)", 
+    y = "State",
+    fill = "Candidate"
+  ) +
+  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
+  theme_minimal() +
+  theme(legend.position = "bottom") + 
+  theme(axis.text.y = element_text(size = 8))
+
+```
+
+In @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump.
+Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.
+
+```{r}
+#| label: fig-avg-pct-methodology
+#| fig-cap: Polls with more Methodologies have Less Variability.
+#| echo: false
+#| warning: false
+#| 
+cleaned_data <- cleaned_data %>% mutate(candidate_name =
+  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
+
+cleaned_data %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
+  filter(!is.na(methodology)) %>% 
+  ggplot(mapping = aes(x = methodology, y = pct)) + 
+  geom_boxplot(width = 0.6, position = position_dodge(width = 0.8), outlier.shape = NA) +
+  coord_flip() +
+  theme_minimal() +
+  labs(
+    x = "Methodology",
+    y = "Polled Support (%)",
+    fill = "Candidate"
+  ) +
+  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
+  theme(legend.position = "bottom") + 
+  theme(axis.text.y = element_text(size = 8))
+
+```
+
+@fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.
+
+### Relationship Between Variables
+
+```{r}
+#| label: fig-pollscore-samplesize
+#| fig-cap: More Reliable Pollsters Have Larger Sample Sizes in their Polls
+#| echo: false
+#| warning: false
+
+# removing outliers and na's
+graph_data <- cleaned_data %>% 
+  filter(!is.na(pollscore) & !is.na(sample_size))
+
+outlier_top <- (quantile(graph_data$sample_size, 0.75) + 
+                  IQR(graph_data$sample_size)*1.5)
+
+outlier_bottom <- (quantile(graph_data$sample_size, 0.25) - 
+                  IQR(graph_data$sample_size)*1.5)
+
+graph_data <- cleaned_data %>%
+  filter(sample_size < outlier_top & sample_size > outlier_bottom)
+
+graph_data %>% filter(!is.na(pollscore) & !is.na(sample_size)) %>% 
+  ggplot(aes(x = pollscore, y = sample_size)) +
+  geom_point(alpha = 0.33) + geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Poll Score", y = "Poll Sample Size")
+
+# this model is really here just for explain, this realtionship should not
+# be treated as lienar
+# cor(graph_data$sample_size, graph_data$pollscore)
+# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
+# summary(lm_score_sample)
+```
+
+@fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
+The figure also shows that most polls have a sample size around 800 to 1200 participants.
+It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.
+
+```{r}
+#| label: fig-pollsamplesize-canidate-pct
+#| fig-cap: The Sample Size of a Poll does not impact a Candidate's Voting Percentage
+#| echo: false
+#| warning: false
+
+cleaned_data %>% 
+  filter(!is.na(sample_size) & !is.na(pct)) %>% 
+  ggplot(aes(x = sample_size, y = pct)) +
+  geom_point(alpha = 0.4) + 
+  geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Sample Size of the Poll",
+       y = "Candidate Percentage") + facet_wrap(~candidate_name)
+
+```
+
+Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.
+
+```{r}
+#| label: fig-pollscore-candidate-pct
+#| fig-cap: More Reliable Pollsters Score Higher For Trump
+#| echo: false
+#| warning: false
+
+cleaned_data %>% 
+  filter(!is.na(pollscore) & !is.na(pct)) %>% 
+  ggplot(aes(x = pollscore, y = pct)) +
+  geom_point(alpha = 0.4) + 
+  geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Pollscore",
+       y = "Candidate Percentage") + facet_wrap(~candidate_name)
+
+```
+
+@fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage.
+For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support.
+Conversely, Kamala Harris shows the opposite trend: as the poll score decreases, her percentage tends to rise.
+This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.
+
+```{r}
+#| label: fig-polldate-canidate-pct
+#| fig-cap: Both Canidate Percentages Are Gaining Support Closer to the Election
+#| echo: false
+#| warning: false
+
+cleaned_data %>% 
+  filter(!is.na(days_taken_from_election) & !is.na(pct)) %>% 
+  ggplot(aes(x = days_taken_from_election, y = pct)) +
+  geom_point(alpha = 0.4) + 
+  geom_smooth(method = "lm", se = FALSE, color = light_red) +
+  labs(x = "Timing of Polls Relative to Election Day (Days Prior to the Election)",
+       y = "Candidate Percentage") + facet_wrap(~candidate_name)
+
+```
+
+Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election.
+This is a result of non-major candidate support rapidly decreasing.
+Specifically, Trump support is increasing at a faster rate than Harris.
+However, the polls in general show a slight lead for Harris throughout the last 100 days.
+
+The rapid decline in third party support could be due to Robert F. Kennedy dropping out of the race (this sentences should probably be in results or discussion section).
+
+# Model {#sec-model}
+
+The goal of our modeling strategy is twofold.
+Firstly, we aim to accurately predict the support percentage (PCT) for Harris and Trump based on relevant poll data and key influencing factors.
+Secondly, we seek to evaluate the efficacy of different modeling approaches—from simple linear regression (SLR) to multiple linear regression (MLR) and Bayesian hierarchical models—to understand their predictive capabilities and assess the underlying relationships between variables.
+By comparing these models, we can determine which approach provides the most robust and reliable predictions, while considering the variability and potential uncertainty in the data.
+
+## Model set-up {#sec-modelsetup}
+
+The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:
+
+- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$
+
+- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$
+
+- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+## Basic SLR Model
+```{r}
+#| label: fig-model-pct-pollscore
+#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
+#| echo: false
+#| warning: false
+
+Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
+just_harris_data = Harris_data |> na.omit()
+lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+# summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+We first set up a SLR model for predicting pct based on pollscore. The model is:
+
+$\hat{\text{pct}} = \beta_0 + \beta_1 \times \text{pollscore}$
+
+In this SLR model, the response variable is pct and the only one predictor is the pollscore.
+@fig-model-pct-pollscore visualizes the relationship between the actual and predicted values of percentage support (pct) for Kamala Harris, based on a simple linear regression model with pollscore as the sole predictor.
+Blue points represent individual comparisons between actual and predicted values.
+The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.
+
+The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line).
+This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct.
+The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
+
+```{r}
+#| label: fig-model-pct-sample_size
+#| fig-cap: "Actual vs. Predicted Values for pct Using sample_size as Predictor"
+#| echo: false
+#| warning: false
+lm_model1 = lm(pct ~ sample_size, data = just_harris_data)
+predictions = predict(lm_model1, just_harris_data)
+# summary(lm_model1)
+
+ggplot(just_harris_data, aes(x = pct, y = predictions)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+
+```
+
+In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor.
+This indicate that sample_size has limited impact on pct.
+
+## MLR model
+```{r}
+#| label: fig-model-pct-pollscore-etc
+#| echo: false
+#| warning: false
+just_harris_data = Harris_data |> na.omit()
+model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
+# summary(model_MLR)
+just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
+
+predictions = predict(model_MLR, just_harris_data)
+
+```
+
+```{r}
+#| label: fig-model-MLR
+#| fig-cap: "Actual vs Predicted Percentage Support for Kamala Harris (Multiple Predictors)"
+#| echo: false
+#| warning: false
+ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
+  geom_point(color = 'blue') +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual pct",
+       y = "Predicted pct") +
+  theme_minimal()
+```
+
+Next, we predicts pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:
+
+$\hat{\text{pct}} = \beta_0 + \beta_1 \cdot \text{pollscore} + \beta_2 \cdot \text{days\_taken\_from\_election} + \beta_3 \cdot \text{methodology} + \beta_4 \cdot \text{sample\_size} + \beta_5 \cdot \text{state}$
+
+@fig-model-MLR compares actual pct values with the predicted pct values from the MLR model.
+Blue points represent individual comparisons between actual and predicted values.
+The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.
+
+The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
+The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60).
+This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.
+
+## Bayesian Model
+
+Next, we set up a bayesian model as the following in accordance to @sec-modelsetup:
+
+\begin{align*}
+\text{pct}_i &\sim \mathcal{N}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_1 \cdot \text{pollscore}_i + \beta_2 \cdot \text{days\_taken\_from\_election}_i + \beta_3 \cdot \text{sample\_size}_i \\
++ u_{\text{methodology}[i]} + u_{\text{state}[i]} \\
+\alpha &\sim \text{Normal}(0, 2.5) \\
+\beta_j &\sim \text{Normal}(0, 2.5), \quad j = 1, 2, 3 \\
+u_{\text{methodology}} &\sim \text{Normal}(0, \sigma_{\text{methodology}}) \\
+u_{\text{state}} &\sim \text{Normal}(0, \sigma_{\text{state}})
+\end{align*}
+
+This hierarchical Bayesian model is designed to capture both the fixed effects of predictors and the random effects of grouping factors.
+The response variable is pct (percentage of support), predictors are pollscore, days taken from election, and sample size.
+A normal prior with mean 0 and standard deviation 2.5 is set for all coefficients and the intercept, scaled automatically.
+
+```{r}
+#| label: fig-bayesian-model
+#| fig-cap: "Posterior Predictive Check for the Bayesian Model Predicting pct"
+#| echo: false
+#| warning: false
+bayesian_model_1 = readRDS(here("models/bayesian_model_1.rds"))
+
+pp_check(bayesian_model_1)
+```
+
+### Posterior Predictive Checks 
+@fig-bayesian-model displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.
+
+The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
+The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
+However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
+
+The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
+Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
+Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.
+
+### Train Test Validation
+we implemented a train-test split validation approach to evaluate the performance of a Bayesian hierarchical model predicting percentage support (pct) for Kamala Harris.
+The dataset was divided into training (70%) and test (30%) sets to fit and validate the model, respectively.
+We used the stan_glmer function to fit the model on the training data, with predictors including pollscore, days_taken_from_election, sample_size, and random intercepts for methodology and state.
+After training, we generated predictions on the test set and calculated an R squared value to quantify the model's predictive accuracy.
+```{r}
+#| echo: false
+#| warning: false
+# Split data into training and test sets
+set.seed(123)
+# train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
+# train_data <- just_harris_data[train_indices, ]
+# test_data <- just_harris_data[-train_indices, ]
+train_data = read_parquet(here("data/02-analysis_data/train_data.parquet"))
+test_data = read_parquet(here("data/02-analysis_data/test_data.parquet"))
+
+# Fit the model on the training data
+bayesian_model_train = readRDS(here("models/bayesian_model_train.rds"))
+# bayesian_model_train <- stan_glmer(
+#   formula = formula,
+#   data = train_data,
+#   family = gaussian(),
+#   prior = priors,
+#   prior_intercept = priors,
+#   seed = 123,
+#   cores = 4,
+#   adapt_delta = 0.95
+# )
+
+# Predict on the test data and check performance
+predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
+```
+
+\begin{align}
+SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
+SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
+R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
+\text{R-squared} = 0.700563202695799
+\end{align}
+
+```{r}
+#| warning: false
+#| echo: false
+ss_total <- sum((test_data$pct - mean(test_data$pct))^2)
+ss_residual <- sum((test_data$pct - colMeans(predictions))^2)
+r_squared <- 1 - (ss_residual / ss_total)
+# "R-squared: 0.700563202695799"
+```
+
+Where:\
+\begin{itemize}
+    \item \( y_i \) represents the actual values,
+    \item \( \bar{y} \) is the mean of the actual values,
+    \item \( \hat{y}_i \) are the predicted values,
+    \item \( SS_{\text{total}} \) is the total sum of squares,
+    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
+\end{itemize}
+
+We could see that the R squared value equal to the 0.700563202695799 suggesting that the model captures a substantial portion of the variability in the data and demonstrates that the model performs well in explaining the variability of the pct in the test data, reflecting strong predictive capabilities.
+
+```{r}
+#| label: fig-bayesian-model-actual-predict
+#| fig-cap: "Predicted vs Actual Percentage Support (pct) for Kamala Harris in the Bayesian Model"
+#| echo: false
+#| warning: false
+plot(test_data$pct, colMeans(predictions), 
+     xlab = "Actual PCT Values", 
+     ylab = "Predicted  PCT Values", 
+     main = "Predicted PCT vs Actual PCT")
+abline(0, 1, col = "red")  # Add a 45-degree reference line
+```
+
+ By visual checking actual pct and predicted pct plot in @fig-bayesian-model-actual-predict, the points are plotted against a 45-degree line, which represents the ideal scenario where predicted values match the actual values perfectly.
+
+```{r}
+#| label: fig-bayesian-model-residual
+#| fig-cap: "Residual Plot for the Bayesian Hierarchical Model Predictions"
+#| echo: false
+#| warning: false
+residuals <- test_data$pct - colMeans(predictions)
+plot(colMeans(predictions), residuals, 
+     xlab = "Predicted Values", 
+     ylab = "Residuals", 
+     main = "Residual Plot")
+abline(h = 0, col = "red")
+```
+
+In addition, the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
+
+In conculsion, the visual checks from the predicted vs.
+actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting.
+It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.
+
+### Model justification
+
+To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
+The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity.
+The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables.
+The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts.
+This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability.
+By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.
+
+# Result {#sec-result}
+```{r}
+#| label: fig-bayesian-model-CI
+#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
+#| echo: false
+#| warning: false
+# Plot random effects
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+```
+
+@fig-bayesian-model-CI displays the estimated posterior distribution for the intercept parameter in the Bayesian model.
+The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S.
+election data used in the analysis.
+The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.
+
+The estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
+election data analyzed, is centered around 52%, with a credible interval spanning approximately from 49.5% to 53.5%.
+The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
+
+```{r}
+#| label: tbl-bayesian-model-result-summary
+#| tbl-cap: "Bayesian Model Result Summary"
+#| echo: false
+#| warning: false
+
+predictions2 = posterior_predict(bayesian_model_1, newdata = just_harris_data)
+
+predicted_means = colMeans(predictions2)
+predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+
+# Combine the results with the actual data
+result_summary = data.frame(
+  Actual_PCT = just_harris_data$pct,
+  Predicted_PCT = predicted_means,
+  Lower_CI = predicted_intervals[1, ],
+  Upper_CI = predicted_intervals[2, ]
+)
+
+kable(head(result_summary, 10))
+```
+
+```{r}
+#| label: fig-bayesian-model-result-CI
+#| fig-cap: "U.S. Election Prediction Results"
+#| echo: false
+#| warning: false
+ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "U.S. Election Prediction Results",
+    x = "Actual PCT",
+    y = "Predicted PCT"
+  )
+```
+
+@fig-bayesian-model-result-CI shows that he model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages.
+The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges.
+This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes.
+@tbl-bayesian-model-result-summary also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.
+
+# Discussion {#sec-discussion}
+
+## Summary of Findings
+
+In this paper, we developed a hybrid election forecasting model to predict the 2024 U.S.
+presidential election results.
+Our model provides a stable and accurate forecast compared to traditional methods by combining poll-of-polls aggregation with additional factors such as poll score, sample size, days from the election, state, and methodology.
+
+Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability.
+The results shows that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.
+
+## Insights on Polling and Election Dynamics
+
+One of the main takeaways from this study is the impact of poll quality on election forecasts.
+We weight and filtered the polls by their reliability.
+This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.
+
+## Limitations of the Model
+
+While this model presents a more comprehensive approach, there are limitations.
+First, Any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data.
+For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.
+
+Second, the model only reflect current voting dynamics as it is based on historical data.
+This model does not account for short-term variability such as Biden's exit.
+This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.
+
+On top of that, the results can be sensitive to the choice of priors.
+We choose normal(0, 2.5) as we believe it could stabilize the estimates by being informative enough to guide the posterior distribution without overpowering the data.
+However, if the chosen prior does not align well with the true underlying distribution, it may lead to biased results or overly wide credible intervals.
+
+Also, as it is a model based on a dataset, it is challenging to capture all the real-life features and dynamics.
+The model might not include all relevant predictors or interactions due to limitations in data availability or complexity, which can lead to incomplete representations of the factors influencing voting behavior.
+Additionally, without strong regularization techniques, the model may become prone to overfitting, particularly when using complex hierarchical structures or including numerous predictors.
+This overfitting can reduce the model’s generalizability to new or unseen data.
+
+Moreover, errors and offsets inherent in polling data, such as response bias, nonresponse adjustments, and sampling variability, can propagate into the model’s results.
+These aspects introduce an additional layer of uncertainty that can affect the model's reliability and predictive performance.
+While Bayesian methods provide a robust framework to incorporate uncertainty, the final outputs must be interpreted cautiously, acknowledging these underlying limitations.
+
+## Future Directions
+To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles.
+We includes previous year's polling dataset in this R project.
+By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics.
+This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion.
+Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, ultimately refining its accuracy and reliability for future applications in election forecasting.
+
+\newpage
+
+\appendix
 
 # Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 
-# Our Prediction
+# Trump Voter Prediction Model
+
+The multiple linear regression model (MLR) for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris.
+Likewise, the Trump dataset is also split into training and testing data.
+The model outputs are below.
+
+```{r}
+#| echo: false
+#| warning: false
+set.seed(123)
+train_data_trump <- read_parquet(here("data/02-analysis_data/train_data_trump.parquet"))
+test_data_trump <- read_parquet(here("data/02-analysis_data/test_data_trump.parquet"))
+bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
+```
+
+```{r}
+#| label: fig-model-trumpmlr
+#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
+#| echo: false
+#| warning: false
+just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
+predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
+predicted_means = colMeans(predictions2)
+predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+result_summary = data.frame(
+  Actual_PCT = just_trump$pct,
+  Predicted_PCT = predicted_means,
+  Lower_CI = predicted_intervals[1, ],
+  Upper_CI = predicted_intervals[2, ]
+)
+# kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")
+ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
+  geom_point(color = "blue", alpha = 0.7) +
+  # geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "Actual vs Predicted Percentage For Trump MLR Model (All Trump Data)",
+    x = "Actual PCT",
+    y = "Predicted PCT"
+  )
+```
+
+From @fig-model-trumpmlr it is clear that the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line).
+Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+```{r}
+#| label: fig-model-trumpmlr-res
+#| fig-cap: "MLR Trump Model does not Appear to Overfit"
+#| echo: false
+#| warning: false
+predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
+plot(test_data_trump$pct, colMeans(predictions2), 
+     xlab = "Actual PCT Values", 
+     ylab = "Predicted  PCT Values", 
+     main = "Predicted PCT vs Actual PCT (Test Data)")
+abline(0, 1, col = "red")  # Add a 45-degree reference line
+```
+
+Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
+This suggests that model can generalization to outside data.
+Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+```{r}
+#| label: fig-trumpmodelci
+#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
+#| echo: false
+#| warning: false
+plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
+```
+
+Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% confidence interval ranges from around 42.5% to 47.2%.
+This confidence interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.
+
+# Election Prediction
 
 Our prediction process consists of two primary components.
-First, we develop models for both Trump and Harris based on the variables outlined in Section @sec-PLACEHOLDER.
+First, we develop models for both Trump and Harris based on the variables outlined in @sec-model.
 This involves partitioning the dataset into training and testing subsets.
@@
 We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa.
-Winners for other states were determined based on historical trends and predictions from sources like @cnn.
+Winners for other states were determined based on historical trends and predictions from sources like @cnn_electoral_nodate.
 Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
+
 ```{r}
@@
 #| warning: false
+#| label: tbl-electionprediction
+#| tbl-cap: "Kamala Harris Wins Most of the Swing States"
 # Split data into training and test sets
 
-set.seed(123)
-
-train_data_harris <- read_csv(here("data/02-analysis_data/train_data_harris.csv"))
-test_data_harris <- read_csv(here("data/02-analysis_data/test_data_harris.csv"))
+train_data_harris <- read_parquet(here("data/02-analysis_data/train_data_harris.parquet"))
+test_data_harris <- read_parquet(here("data/02-analysis_data/test_data_harris.parquet"))
 bayesian_model_train_harris <- readRDS(here("models/bayesian_model_train_harris.rds"))
-
-train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
-test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
-bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
 
@@
             "Trump Predicted Percentage", "State Winner"))
-
 ```
@@
 ```{r}
-#| label: fig-currentstateofelection
-#| fig-cap: Kamala Harris is Predicited to be the 47th President of the United States
+#| label: fig-electionpredictionmap
+#| fig-cap: "Kamala Harris is Predicted to be the 47th President of the United States"
 #| echo: false
@@
 ```
+
 According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States.
-There are also a few states with predictions not visible in the map, we will describe those predicts them below
+There are also a few states with predictions not visible in the map, we will describe those predicts them below.
 
@@
 
+# Overview of American Electon System
+
+## Brief Description of American Federal Goverment
+
+The American federal or national government is split into three branches, executive., legislative, and judical The executive branch includes the president and the military.
+The legislative branch include two subgroups, the House of Representatives and the Senate.
+These two subgroups create laws.
+Every state has two Senators and one Representative per approximately 750,000 people.
+The judicial branch is court system.
+
+## What is the Electoral College {#sec-electoral-college}
+
+The Electoral College is the system used in the United States to elect the president and vice president.
+Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives).
+When voters cast their ballots, they are actually voting for a slate of electors pledged to a candidate.
+The candidate who receives a majority of electoral votes (270 out of 538) wins the presidency.
+This system means that winning the popular vote in a state generally results in winning all of that state's electoral votes.
+The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.
+
+The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts.
+Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election.
+On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election.
+As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there.
+
 ```{r}
+#| label: fig-currentstateofelection
+#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
+#| echo: false
+#| warning: false
+light_blue <- "#ADD8E6"
+dark_blue <- "#6688CC"
+light_red <- "#FF7777"
+dark_red <- "#B22222"
+purple <- "#9370DB"
+# Define your colors for each state (replace these colors with your desired ones)
+state_colors_legend <- c(
+  "Republican" = dark_red,      # Dark Red
+  "Lean Republican" = light_red, # Light Red
+  "Toss-Up" = purple,         # Purple
+  "Lean Democrat" = light_blue,   # Light Blue
+  "Democrat" = dark_blue,        # Dark Blue
+  "Split By District" = "gray"
+)
 
-plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
+state_colors <- c(
+  alabama = dark_red,
+  arizona = purple,
+  arkansas = dark_red,
+  california = dark_blue,
+  colorado = dark_blue,
+  connecticut = dark_blue,
+  delaware = dark_blue,
+  `district of columbia` = dark_blue,
+  florida = light_red,
+  georgia = purple,
+  idaho = dark_red,
+  illinois = dark_blue,
+  indiana = dark_red,
+  iowa = light_red,
+  kansas = dark_red,
+  kentucky = dark_red,
+  louisiana = dark_red,
+  maine = "gray",
+  maryland = dark_blue,
+  massachusetts = dark_blue,
+  michigan = purple,
+  minnesota = light_blue,
+  mississippi = dark_red,
+  missouri = dark_red,
+  montana = dark_red,
+  nebraska = "gray",
+  nevada = purple,
+  `new hampshire` = light_blue,
+  `new jersey` = dark_blue,
+  `new mexico` = light_blue,
+  `new york` = dark_blue,
+  `north carolina` = purple,
+  `north dakota` = dark_red,
+  ohio = light_red,
+  oklahoma = dark_red,
+  oregon = dark_blue,
+  pennsylvania = purple,
+  `rhode island` = dark_blue,
+  `south carolina` = dark_red,
+  `south dakota` = dark_red,
+  tennessee = dark_red,
+  texas = light_red,
+  utah = dark_red,
+  vermont = dark_blue,
+  virginia = light_blue,
+  washington = dark_blue,
+  `west virginia` = dark_red,
+  wisconsin = purple,
+  wyoming = dark_red
+)
+
+# Load the states map data
+states_map <- map_data(map = "state")
+
+state_colors <- setNames(state_colors, tolower(names(state_colors)))
+
+color_to_status <- setNames(names(state_colors_legend), state_colors_legend)
+
+states_map <- states_map %>%
+  mutate(status = color_to_status[state_colors[region]])
+
+plotclr <- c(dark_red, light_blue, purple, gray, light_blue, dark_blue)
+legend_description <- c(
+  "Likely Republican",
+  "Lean Republican",
+  "Swing States",
+  "Lean Democrat",
+  "Likely Democrat",
+  "Split by District"
+)
+
+ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
+  geom_polygon(color = "white", size = 0.2) +
+  coord_map("albers", lat0 = 39, lat1 = 45) +
+  scale_fill_manual(
+    values = state_colors_legend,
+    name = "Election Status"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    legend.title = element_text(size = 12, face = "bold"),
+    legend.text = element_text(size = 10),
+    axis.text = element_blank(),
+    axis.title = element_blank(),
+    panel.grid = element_blank()
+  )
 
 ```
+
+The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are generally agreed upon within the American political community.
+These organizations assessed historical voting patterns and recent polling data to derive their conclusions.
+
+Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district.
+In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic.
+In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican.
+Furthermore, Alaska and Hawaii are not in the map.
+Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.
+
+The seven generally agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.
+
+# States Poll Count
+
+```{r}
+#| label: tbl-state-polls-count
+#| tbl-cap: Polls included in Analysis Per State
+#| echo: false
+#| warning: false
+
+states <- c(state.name, "National")
+polls_per_state_tbl <- tibble(state = states)
+
+polls_per_state <- cleaned_data %>% group_by(state) %>% 
+  summarise(polls_count = n_distinct(poll_id))
+
+polls_per_state_tbl <- polls_per_state_tbl %>% 
+  left_join(polls_per_state, by = "state") %>% 
+  mutate(polls_count = if_else(!is.na(polls_count), polls_count, 0)) %>% 
+  select(state, polls_count)
+
+polls_per_state_tbl %>% kable(col.names = c("State", "Number of Polls In Analysis"))
+
+num_states_missing <- count(polls_per_state_tbl %>% 
+                              filter(polls_count == 0) %>% 
+                              select(state))
+
+```
+
+# The Ideal Survey
+
+## Objective
+
+The research team has developed a survey and distribution methodology with a hypothetical budget of $100,000.
+This objective is to create a polling system that accurately predicts the 2024 United States Election.
+
+## Sampling Frame
+
+Given the monetary constraint, the team's first decision is to mostly include swing states and districts.
+These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia.
+Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college.
+However, the poll will also include Texas, Florida, and Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.
+Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida and Minnesota.
+
+## Survey Sending Process
+
+The team will then find and use multiple databases, such as @usps_customer_nodate, to find addresses, telephones, and emails of potential voters in those states.
+According to @pew_research_center_us_nodate, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums.
+Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email.
+Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.
+
+## Sampling Methodology
+
+Each state will have its own poll, but the polling methodology and the survey given will remain the same.
+Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes.
+Ideally, the stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for.
+However, these factors are almost impossible to determine while sending the survey.
+Therefore, the research team decided to ask demographic questions directly inside the survey.
+After data collection, the team will weigh data based on demographics.
+For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.
+
+Furthermore, the team will aim to sample approximately 1,000 people from each true swing state.
+This sample size is large enough to provide reliable conclusions but not so large that it resembles sampling with replacement.
+For true swing states, the team aims to survey 1,000 individuals.
+However, if the final sample size falls short of this target, it is not considered a significant issue.
+
+## Survey Implementation & Question Creation
+
+The team's ideal survey will be made using @qualtrics_qualtrics_nodate.
+It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants.
+To identify participants who are not engaged, the research team has decided to add a worthless question.
+This question is extremely simple and clearly has one correct answer.
+Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data.
+An example of this is question number 7.
+
+Order bias occurs when the sequence of questions subconsciously influences participants' responses.
+To prevent this, many questions should be randomized upon entry to the survey.
+However, some questions must follow a specific order, such as question 11, while others like questions 1-4 can be randomized.
+The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics_qualtrics_nodate paid plan.
+
+Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently.
+To prevent this, answer choices should be randomized upon survey entry.
+Questions 8 & 9 could benefit from this.
+Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics_qualtrics_nodate paid plan.
+
+A leading question occurs when a question is written in a way that suggests the user to give a certain answer.
+For example, "given that children are the future of our country, should we invest more money in their education".
+To prevent this, the researchers have ensured questions are written in a style where no unnecessary details or opinions are added.
+
+## Survey Questions
+
+Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics_qualtrics_nodate survey.
+
+List of all of the questions:
+
+1. Select your race(s) (racial options where chosen based on @orvis_omb_2024)
+   - White
+   - Black or African American 
+   - American Indian or Alaska Native 
+   - Native Hawaiian or Pacific Islander 
+   - Middle Eastern or North African
+   - Asian 
+   - Other 
+   - Prefer not to answer
+
+2. Please select your gender 
+   - Male 
+   - Female 
+   - Non-binary 
+   - Prefer not to say
+
+3. Please enter your age (in numbers) 
+  - This is a text input field.
+Please note that this field has the auto-validation feature set to numbers in @qualtrics_qualtrics_nodate.
+As a result, participants can only input numbers in this field and are alert if they have not.
+
+4. Please select the highest degree of education you have obtained
+   - GED Certificate 
+   - High School Diploma 
+   - Undergraduate Degree 
+   - Graduate Degree (Masters/Phd) 
+   - None 
+   - Other
+   - Prefer not to answer
+
+5. Are you a registered voter for the 2024 United States Presidential Election? 
+   - Yes 
+   - No
+
+6. Place yourself on the political spectrum 
+   - Far Left 
+   - Center Left 
+   - Center 
+   - Center Right 
+   - Far Right
+
+7. Can pigs fly? 
+   - Yes 
+   - Maybe
+   - No 
+   - Prefer Not To Say
+
+8. What political party have you registered with? 
+   - Republicans 
+   - Democrats 
+   - Green Party 
+   - Libertarian 
+   - Other
+   - Independent (unregistered)
+
+9. Who will you vote for in the 2024 presidential election? 
+   - Democrat - Kamala Harris 
+   - Republican - Donald Trump 
+   - Green Party - Gill Stein 
+   - Libertarian - Chase Oliver 
+   - Other
+   - Will not vote
+
+## Potential Problems with the Methodology And Polls
+
+While the team's methodology and survey creates a robust system, there are potential issues.
+The reliance of weighting results based on a candidate's demographics can lead to error propagation.
+For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have a large impact on the polls prediction of what candidate will win the state.
+There is also a selection bias in terms of the monetary reward.
+Potentially, people who like monetary rewards could be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.
+
+# Methodology of YouGov
+YouGov's methodology documentations are seperated in two articles.
+The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
+
+## Population, Frame, and Sample
+As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
+The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
+Voter files indicates whether someone voted in a given election, thus YouGov's population covers all voters in previous US elections.
+
+YouGov’s sampling frame consists of its online panel members.
+These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
+YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
+
+YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
+YouGov interviews nearly 100,000 people in the first set of estimates.
+For the second set of estimates, YouGov didn't just start over with a new sample.
+They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
+
+## Sample Recruitment
+Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
+They must provide demographic details upon joining, which helps in selecting representative samples for each survey.
+When respondents complete a survey, they are awarded points that can be exchanged for money.
+
+## Sampling Approach and Trade-offs
+YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
+This method allows quick and cost-effective data collection.
+However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
+YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
+
+## Non-response Handling
+YouGov apply statistical weighting to adjust for the differences between the sample and target population.
+The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
+Additionally, quality control measures exclude unreliable responses to improve data accuracy.
+The respondents are offered a small incentive to decrease the non-response and increase participation.
+
+## Strengths and Weaknesses of the Questionnaire
+YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
+The pollster can recruit a large amount of panelists because of the online format.
+Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
+
+As a non-probability sample, it might miss certain demographic groups not covered by the online population.
+While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
+Additionaly, the categories in the surbey is oversimplified with bias.
+For instance, in the poll result published by @yougov_say24_nodate, gender is divided into Male and Female.
+Race is divided into White, Black, Hispanic and Other.
+This indicates a lack of representation.
 

### 2024-11-04 paper/paper.qmd
add limit on prediction

@@
 
-## Insights on Polling and Election Dynamics
+## Polling and Election Dynamics
 
@@
 
-## Limitations of the Model
+## Limitations of the Model {#sec-model-limitations}
 
@@
 While Bayesian methods provide a robust framework to incorporate uncertainty, the final outputs must be interpreted cautiously, acknowledging these underlying limitations.
+
+## Limitation of the Prediction
+
+In @sec-prediction, the team predicted that Kamala Harris would secure 298 delegates while Donald Trump would obtain 240 delegates.
+However, this prediction has several limitations.
+Firstly, the model used to generate these predictions inherits the constraints outlined in  @sec-model-limitations.
+Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.
+
+Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27.
+Moreover, as noted in Section @sec-prediction, the analysis focused solely on closely contested races, specifitatiocally swing states.
+Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions.
+Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.
 
@@
 
-# Election Prediction
+# Election Prediction {#sec-prediction}
 

### 2024-11-04 paper/paper.qmd

@@
 
-## Data Exploration
+## Data Exploration {#sec-dataexplore}
 
@@
 
-<!-- ## Data Exploration
-
-```{r}
-#| label: fig-pairs
-#| fig-cap: "Pairs Plot of Key Numeric Variables Related to Percentage Support (pct)"
-#| echo: false
-#| warning: false
-# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
-numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
-
-# Create the pairs plot
-pairs(numeric_data)
-```
-
-In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
-The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
-The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
-No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
--->
+```{r}
+#| label: fig-explore-data
+#| fig-cap: "Exploratory Analysis of Predictors and Candidate Support (pct)"
+#| echo: false
+#| warning: false
+a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'sample size', y='pct', 
+       title = 'pct vs sample size')
+
+b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'pollscore', y='pct', 
+       title = 'pct vs pollscore')
+
+c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'methodology', y='pct', 
+       title = 'pct vs methodology')
+
+d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
+  geom_point() + 
+  geom_smooth(method = lm, se = FALSE) + 
+  labs(x = 'max gust speed', y='pct', 
+       title = 'pct vs days taken from election')
+
+e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
+  geom_boxplot() + 
+  labs(x = 'state', y='pct', 
+       title = 'pct vs state')
+
+grid.arrange(a, b, c, d, e, nrow=2)
+```
+
+@fig-explore-data explores the relationship between candidate support percentage (pct) and various predictors in the dataset: (a) sample size of polls, (b) pollscore (an indicator of poll reliability), (c) polling methodology, (d) days taken from the election, and (e) state.
 
@@
 ### Model justification
+
+We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots.
+As shown in @fig-explore-data in @sec-dataexplore, The relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy.
+The plot of pct versus poll score reveals a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct.
+Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value.
+These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.
 

### 2024-11-04 paper/paper.qmd
added sources of bias to discussion

--- 2024-11-04 paper/paper.qmd
@@
 ---
-title: "Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality"
-subtitle: "A Hybrid Model Predicts a Narrow Lead for Kamala Harris in Swing States as Election Day Approaches"
+title: "Prediction of 2024 US election ..."
 author: 
@@
 date-format: long
-abstract: "This paper develops a hybrid model to forecast the 2024 U.S.
-presidential election by combining poll aggregation techniques with additional variables, including pollster reliability, sample size, timing relative to the election, geographic region, and polling methodology.
-By integrating these factors, our model accounts for variations in poll quality and regional differences in voter sentiment, providing a more stable and accurate prediction compared to traditional poll aggregation alone.
-Our results show that Kamala Harris holds a slight lead over Donald Trump in most swing states, with support levels stabilizing closer to election day.
-This approach underscores the importance of nuanced forecasting models in election analysis, revealing how poll quality and timing can significantly influence support predictions.
-Ultimately, this model offers a comprehensive tool for understanding the dynamics of voter sentiment and improving the accuracy of election predictions."
+abstract: "We forecast the winner of the 2024 US presidential election using “poll-of-polls” by building a linear model."
 format: pdf
@@
 library(knitr)
-library(arrow)
+library(brms)
 library(rstanarm)
-library(brms)
-library(dplyr)
-library(lubridate)
 library(gridExtra)
@@
 library(rstan)
+library(brms)
 
-light_blue <- "#ADD8E6"
-dark_blue <- "#6688CC"
-light_red <- "#FF7777"
-dark_red <- "#B22222"
-purple <- "#9370DB"
+```
 
-cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
+```{r}
+#| echo: false
+#| warning: false
+cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
 ```
@@
 As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts.
-However, conventional aggregation models often overlook dynamic election factors, including the credibility and quality variations among poll sources.
+However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data.
+This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.
 
-We build a hybrid election forecasting model following the strategies mentioned by @pasek.
+This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek.
 As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors.
-Our model incorporates this approach, filtering polls from @fivethirtyeight based on a numeric grade that indicates each pollster's reliability.
+Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics.
+Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.
 
-The objective of this study is to predict the support percentages for the primary candidates, Kamala Harris and Donald Trump, using this hybrid model.
-We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability.
-We also quantify other polling attributes, such as sample size, poll reliability scores, and polling methodology, which serve as predictors.
-Our results show that this model provides a more stable and accurate forecast than traditional poll aggregation alone.
-
-The remainder of this paper is structured as follows: @sec-data provides an overview and exploration of the data.
-@sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model.
-We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.
-
-The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here], gridExtra [@gridextra], Matrix [@Matrix], Rstan [@Rstan] and lubridate [@lubridate].
-
-## Estimand
+In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model.
+We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S.
+presidential election.
 
@@
 
-# Data {#sec-data}
+The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]
 
-## Measurement
-
-The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys.
-According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024].
-The dataset contains a list of questions in polls and surveys and their results.
-When a new poll is conducted, the poll is appended to the dataset.
-FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.
-
-In each of the polls recorded by FiveThirtyEight, all options of each question are recorded in candidate_names, while the proportion of respondents choosing that option is recorded in pct as a percentage.
-The pct is our primary response variable.
-
-There are limitations in these measurements.
-The differences in sampling methods, wording in survey questions and systematic biases are all reflected in the outcome [@radcliffe_538s_2023].
-Thus, in the dataset, FiveThirtyEight includes other variables, such as pollscore and sample_size for their prediction model in addition to the polling results for transparency and accuracy.
-As an example, pollscore indicates the pollster’s reliability rating, a low pollscore indicates a higher reliability rating [@silver_polls_2008].
-
-In this dataset, each row represents a polling question that records the variables of interest.
-Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
-This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
-
-These variables combined allow researchers to reliably analyze and predict the 2024 US election result over time.
-
-## Data Exploration
-
-The raw data from FiveThirtyEight contains 52 columns, all of the column headers are displayed below: 
-```{r}
-#| echo: false
-#| warning: false
-raw_data <- read_csv(here("data/01-raw_data/president_polls.csv"))
-
-# Get the column names of your data
-column_names <- colnames(raw_data)
-
-# Reshape the column names into a matrix with, for example, 4 columns
-num_cols <- 3
-column_matrix <- matrix(column_names, ncol = num_cols, byrow = TRUE)
-
-# Convert the matrix to a data frame for kable
-column_df <- as.data.frame(column_matrix)
-
-# Display in multiple columns with kable
-kable(column_df, format = "latex", booktabs = TRUE, col.names = NULL)
-```
-
-These columns can be categorized into three types, response variable, numeric predictors and categorical predictors.
-
-The response variable is:
-
-- pct: The percentage of support or vote share that each candidate (Kamala Harris or Donald Trump) received in the poll.
-
-Example of numeric predictors are:
-
-- sample_size: The total number of respondents in each poll.
-- pollscore: A numeric score representing the error and bias of the pollster. Negative numbers are better.
-- numeric_grade: A numeric grade assigned to each pollster, reflecting pollster quality or reliability.
-
-Example of categorical predictors are:
-
-- state: The U.S. state in which the poll was conducted or targeted.
-- methodology: The method used to conduct the poll (e.g., online, phone, in-person).
-- party: The political party of the candidate (e.g., DEM for Democrat, REP for Republican).
-- candidate_name: The name of the candidate, either Kamala Harris or Donald Trump.
-- pollster: The name of the polling organization that conducted the poll.
-- stage: The stage of the election (e.g., "general").
-
-<!-- ## Data Exploration
-
-```{r}
-#| label: fig-pairs
-#| fig-cap: "Pairs Plot of Key Numeric Variables Related to Percentage Support (pct)"
-#| echo: false
-#| warning: false
-# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
-numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)
-
-# Create the pairs plot
-pairs(numeric_data)
-```
-
-In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships.
-The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity.
-The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering.
-No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.
--->
-
-## Clean Data
-The data cleaning process involves several steps to ensure the quality and relevance of the polling data.
-First, we filter the dataset to retain only poll results with a numeric grade of 2.7 or higher, indicating that the polls are considered reliable.
-Next, we address missing values in the state attribute: polls with NA in the state column are considered national polls.
-
-We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S.
-election date.
-Additionally, we filter the dataset to include only polls conducted after July 21, 2024, the date when Kamala Harris declared her candidacy.
-Finally, we remove any remaining rows that contain missing values to ensure a clean dataset.
-
-### States included in anaylsis
-
-After the data cleaning process, 21 states had no polling data. A table showing the number of polls for each state, including those without any polls, is provided in @tbl-state-polls-count.
-
-This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
-The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.
-
-## Cleaned data
-
-The first 6 rows of the dataset is displayed in @tbl-cleaned-data.
-
-```{r}
-#| label: tbl-cleaned-data
-#| tbl-cap: Sample of cleaned US election data 
-#| echo: false
-#| warning: false
-
-cleaned_data |>
-  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
-  head(6) |>
-  kable(
-    col.names = c("pct", "Sample Size", "pollscore", "Days taken from election","state", "methodology", "Candidate name"),
-    booktabs = TRUE
-    ) 
-```
-
-### Basic Statistics Summary for Cleaned Data
-
-```{r}
-#| label: fig-avg-pct-state
-#| fig-cap: Historical State Voting Trends Are Maintined in 2024
-#| echo: false
-#| warning: false
-#| 
-cleaned_data <- cleaned_data %>% mutate(candidate_name =
-  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
-
-cleaned_data %>%
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
-  group_by(state, candidate_name) %>%
-  summarise(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop") %>%
-  ggplot(mapping = aes(x = avg_pct, y = state, fill = candidate_name)) + 
-  geom_col(position = "stack") +
-  theme_minimal() +
-  labs(
-    x = "Average Polled Support (%)", 
-    y = "State",
-    fill = "Candidate"
-  ) +
-  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
-  theme_minimal() +
-  theme(legend.position = "bottom") + 
-  theme(axis.text.y = element_text(size = 8))
-
-```
-
-In @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump.
-Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.
-
-```{r}
-#| label: fig-avg-pct-methodology
-#| fig-cap: Polls with more Methodologies have Less Variability.
-#| echo: false
-#| warning: false
-#| 
-cleaned_data <- cleaned_data %>% mutate(candidate_name =
-  if_else(!(candidate_name %in% c("Kamala Harris", "Donald Trump")), "Other", candidate_name))
-
-cleaned_data %>%
-  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
-  filter(!is.na(methodology)) %>% 
-  ggplot(mapping = aes(x = methodology, y = pct)) + 
-  geom_boxplot(width = 0.6, position = position_dodge(width = 0.8), outlier.shape = NA) +
-  coord_flip() +
-  theme_minimal() +
-  labs(
-    x = "Methodology",
-    y = "Polled Support (%)",
-    fill = "Candidate"
-  ) +
-  scale_fill_manual(values = c("Kamala Harris" = light_blue, "Donald Trump" = light_red)) +
-  theme(legend.position = "bottom") + 
-  theme(axis.text.y = element_text(size = 8))
-
-```
-
-@fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.
-
-### Relationship Between Variables
-
-```{r}
-#| label: fig-pollscore-samplesize
-#| fig-cap: More Reliable Pollsters Have Larger Sample Sizes in their Polls
-#| echo: false
-#| warning: false
-
-# removing outliers and na's
-graph_data <- cleaned_data %>% 
-  filter(!is.na(pollscore) & !is.na(sample_size))
-
-outlier_top <- (quantile(graph_data$sample_size, 0.75) + 
-                  IQR(graph_data$sample_size)*1.5)
-
-outlier_bottom <- (quantile(graph_data$sample_size, 0.25) - 
-                  IQR(graph_data$sample_size)*1.5)
-
-graph_data <- cleaned_data %>%
-  filter(sample_size < outlier_top & sample_size > outlier_bottom)
-
-graph_data %>% filter(!is.na(pollscore) & !is.na(sample_size)) %>% 
-  ggplot(aes(x = pollscore, y = sample_size)) +
-  geom_point(alpha = 0.33) + geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Poll Score", y = "Poll Sample Size")
-
-# this model is really here just for explain, this realtionship should not
-# be treated as lienar
-# cor(graph_data$sample_size, graph_data$pollscore)
-# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
-# summary(lm_score_sample)
-```
-
-@fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
-The figure also shows that most polls have a sample size around 800 to 1200 participants.
-It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.
-
-```{r}
-#| label: fig-pollsamplesize-canidate-pct
-#| fig-cap: The Sample Size of a Poll does not impact a Candidate's Voting Percentage
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(sample_size) & !is.na(pct)) %>% 
-  ggplot(aes(x = sample_size, y = pct)) +
-  geom_point(alpha = 0.4) + 
-  geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Sample Size of the Poll",
-       y = "Candidate Percentage") + facet_wrap(~candidate_name)
-
-```
-
-Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.
-
-```{r}
-#| label: fig-pollscore-candidate-pct
-#| fig-cap: More Reliable Pollsters Score Higher For Trump
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(pollscore) & !is.na(pct)) %>% 
-  ggplot(aes(x = pollscore, y = pct)) +
-  geom_point(alpha = 0.4) + 
-  geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Pollscore",
-       y = "Candidate Percentage") + facet_wrap(~candidate_name)
-
-```
-
-@fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage.
-For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support.
-Conversely, Kamala Harris shows the opposite trend: as the poll score decreases, her percentage tends to rise.
-This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.
-
-```{r}
-#| label: fig-polldate-canidate-pct
-#| fig-cap: Both Canidate Percentages Are Gaining Support Closer to the Election
-#| echo: false
-#| warning: false
-
-cleaned_data %>% 
-  filter(!is.na(days_taken_from_election) & !is.na(pct)) %>% 
-  ggplot(aes(x = days_taken_from_election, y = pct)) +
-  geom_point(alpha = 0.4) + 
-  geom_smooth(method = "lm", se = FALSE, color = light_red) +
-  labs(x = "Timing of Polls Relative to Election Day (Days Prior to the Election)",
-       y = "Candidate Percentage") + facet_wrap(~candidate_name)
-
-```
-
-Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election.
-This is a result of non-major candidate support rapidly decreasing.
-Specifically, Trump support is increasing at a faster rate than Harris.
-However, the polls in general show a slight lead for Harris throughout the last 100 days.
-
-The rapid decline in third party support could be due to Robert F. Kennedy dropping out of the race (this sentences should probably be in results or discussion section).
-
-# Model {#sec-model}
-
-The goal of our modeling strategy is twofold.
-Firstly, we aim to accurately predict the support percentage (PCT) for Harris and Trump based on relevant poll data and key influencing factors.
-Secondly, we seek to evaluate the efficacy of different modeling approaches—from simple linear regression (SLR) to multiple linear regression (MLR) and Bayesian hierarchical models—to understand their predictive capabilities and assess the underlying relationships between variables.
-By comparing these models, we can determine which approach provides the most robust and reliable predictions, while considering the variability and potential uncertainty in the data.
-
-## Model set-up {#sec-modelsetup}
-
-The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:
-
-- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$
-
-- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$
-
-- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-## Basic SLR Model
-```{r}
-#| label: fig-model-pct-pollscore
-#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
-#| echo: false
-#| warning: false
-
-Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
-just_harris_data = Harris_data |> na.omit()
-lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
-predictions = predict(lm_model1, just_harris_data)
-# summary(lm_model1)
-
-ggplot(just_harris_data, aes(x = pct, y = predictions)) +
-  geom_point(color = 'blue') +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
-  theme_minimal()
-```
-
-We first set up a SLR model for predicting pct based on pollscore. The model is:
-
-$\hat{\text{pct}} = \beta_0 + \beta_1 \times \text{pollscore}$
-
-In this SLR model, the response variable is pct and the only one predictor is the pollscore.
-@fig-model-pct-pollscore visualizes the relationship between the actual and predicted values of percentage support (pct) for Kamala Harris, based on a simple linear regression model with pollscore as the sole predictor.
-Blue points represent individual comparisons between actual and predicted values.
-The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.
-
-The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line).
-This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct.
-The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
-
-```{r}
-#| label: fig-model-pct-sample_size
-#| fig-cap: "Actual vs. Predicted Values for pct Using sample_size as Predictor"
-#| echo: false
-#| warning: false
-lm_model1 = lm(pct ~ sample_size, data = just_harris_data)
-predictions = predict(lm_model1, just_harris_data)
-# summary(lm_model1)
-
-ggplot(just_harris_data, aes(x = pct, y = predictions)) +
-  geom_point(color = 'blue') +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
-  theme_minimal()
-
-```
-
-In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor.
-This indicate that sample_size has limited impact on pct.
-
-## MLR model
-```{r}
-#| label: fig-model-pct-pollscore-etc
-#| echo: false
-#| warning: false
-just_harris_data = Harris_data |> na.omit()
-model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
-# summary(model_MLR)
-just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))
-
-predictions = predict(model_MLR, just_harris_data)
-
-```
-
-```{r}
-#| label: fig-model-MLR
-#| fig-cap: "Actual vs Predicted Percentage Support for Kamala Harris (Multiple Predictors)"
-#| echo: false
-#| warning: false
-ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
-  geom_point(color = 'blue') +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
-  theme_minimal()
-```
-
-Next, we predicts pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:
-
-$\hat{\text{pct}} = \beta_0 + \beta_1 \cdot \text{pollscore} + \beta_2 \cdot \text{days\_taken\_from\_election} + \beta_3 \cdot \text{methodology} + \beta_4 \cdot \text{sample\_size} + \beta_5 \cdot \text{state}$
-
-@fig-model-MLR compares actual pct values with the predicted pct values from the MLR model.
-Blue points represent individual comparisons between actual and predicted values.
-The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.
-
-The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
-The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60).
-This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.
-
-## Bayesian Model
-
-Next, we set up a bayesian model as the following in accordance to @sec-modelsetup:
-
-\begin{align*}
-\text{pct}_i &\sim \mathcal{N}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_1 \cdot \text{pollscore}_i + \beta_2 \cdot \text{days\_taken\_from\_election}_i + \beta_3 \cdot \text{sample\_size}_i \\
-+ u_{\text{methodology}[i]} + u_{\text{state}[i]} \\
-\alpha &\sim \text{Normal}(0, 2.5) \\
-\beta_j &\sim \text{Normal}(0, 2.5), \quad j = 1, 2, 3 \\
-u_{\text{methodology}} &\sim \text{Normal}(0, \sigma_{\text{methodology}}) \\
-u_{\text{state}} &\sim \text{Normal}(0, \sigma_{\text{state}})
-\end{align*}
-
-This hierarchical Bayesian model is designed to capture both the fixed effects of predictors and the random effects of grouping factors.
-The response variable is pct (percentage of support), predictors are pollscore, days taken from election, and sample size.
-A normal prior with mean 0 and standard deviation 2.5 is set for all coefficients and the intercept, scaled automatically.
-
-```{r}
-#| label: fig-bayesian-model
-#| fig-cap: "Posterior Predictive Check for the Bayesian Model Predicting pct"
-#| echo: false
-#| warning: false
-bayesian_model_1 = readRDS(here("models/bayesian_model_1.rds"))
-
-pp_check(bayesian_model_1)
-```
-
-### Posterior Predictive Checks 
-@fig-bayesian-model displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.
-
-The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
-The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
-However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
-
-The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
-Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
-Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.
-
-### Train Test Validation
-we implemented a train-test split validation approach to evaluate the performance of a Bayesian hierarchical model predicting percentage support (pct) for Kamala Harris.
-The dataset was divided into training (70%) and test (30%) sets to fit and validate the model, respectively.
-We used the stan_glmer function to fit the model on the training data, with predictors including pollscore, days_taken_from_election, sample_size, and random intercepts for methodology and state.
-After training, we generated predictions on the test set and calculated an R squared value to quantify the model's predictive accuracy.
-```{r}
-#| echo: false
-#| warning: false
-# Split data into training and test sets
-set.seed(123)
-# train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
-# train_data <- just_harris_data[train_indices, ]
-# test_data <- just_harris_data[-train_indices, ]
-train_data = read_parquet(here("data/02-analysis_data/train_data.parquet"))
-test_data = read_parquet(here("data/02-analysis_data/test_data.parquet"))
-
-# Fit the model on the training data
-bayesian_model_train = readRDS(here("models/bayesian_model_train.rds"))
-# bayesian_model_train <- stan_glmer(
-#   formula = formula,
-#   data = train_data,
-#   family = gaussian(),
-#   prior = priors,
-#   prior_intercept = priors,
-#   seed = 123,
-#   cores = 4,
-#   adapt_delta = 0.95
-# )
-
-# Predict on the test data and check performance
-predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
-```
-
-\begin{align}
-SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
-SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
-R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
-\text{R-squared} = 0.700563202695799
-\end{align}
-
-```{r}
-#| warning: false
-#| echo: false
-ss_total <- sum((test_data$pct - mean(test_data$pct))^2)
-ss_residual <- sum((test_data$pct - colMeans(predictions))^2)
-r_squared <- 1 - (ss_residual / ss_total)
-# "R-squared: 0.700563202695799"
-```
-
-Where:\
-\begin{itemize}
-    \item \( y_i \) represents the actual values,
-    \item \( \bar{y} \) is the mean of the actual values,
-    \item \( \hat{y}_i \) are the predicted values,
-    \item \( SS_{\text{total}} \) is the total sum of squares,
-    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
-\end{itemize}
-
-We could see that the R squared value equal to the 0.700563202695799 suggesting that the model captures a substantial portion of the variability in the data and demonstrates that the model performs well in explaining the variability of the pct in the test data, reflecting strong predictive capabilities.
-
-```{r}
-#| label: fig-bayesian-model-actual-predict
-#| fig-cap: "Predicted vs Actual Percentage Support (pct) for Kamala Harris in the Bayesian Model"
-#| echo: false
-#| warning: false
-plot(test_data$pct, colMeans(predictions), 
-     xlab = "Actual PCT Values", 
-     ylab = "Predicted  PCT Values", 
-     main = "Predicted PCT vs Actual PCT")
-abline(0, 1, col = "red")  # Add a 45-degree reference line
-```
-
- By visual checking actual pct and predicted pct plot in @fig-bayesian-model-actual-predict, the points are plotted against a 45-degree line, which represents the ideal scenario where predicted values match the actual values perfectly.
-
-```{r}
-#| label: fig-bayesian-model-residual
-#| fig-cap: "Residual Plot for the Bayesian Hierarchical Model Predictions"
-#| echo: false
-#| warning: false
-residuals <- test_data$pct - colMeans(predictions)
-plot(colMeans(predictions), residuals, 
-     xlab = "Predicted Values", 
-     ylab = "Residuals", 
-     main = "Residual Plot")
-abline(h = 0, col = "red")
-```
-
-In addition, the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
-
-In conculsion, the visual checks from the predicted vs.
-actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting.
-It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.
-
-### Model justification
-
-To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
-The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity.
-The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables.
-The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts.
-This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability.
-By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.
-
-# Result {#sec-result}
-```{r}
-#| label: fig-bayesian-model-CI
-#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
-#| echo: false
-#| warning: false
-# Plot random effects
-plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
-```
-
-@fig-bayesian-model-CI displays the estimated posterior distribution for the intercept parameter in the Bayesian model.
-The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S.
-election data used in the analysis.
-The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.
-
-The estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
-election data analyzed, is centered around 52%, with a credible interval spanning approximately from 49.5% to 53.5%.
-The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
-
-```{r}
-#| label: tbl-bayesian-model-result-summary
-#| tbl-cap: "Bayesian Model Result Summary"
-#| echo: false
-#| warning: false
-
-predictions2 = posterior_predict(bayesian_model_1, newdata = just_harris_data)
-
-predicted_means = colMeans(predictions2)
-predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
-
-# Combine the results with the actual data
-result_summary = data.frame(
-  Actual_PCT = just_harris_data$pct,
-  Predicted_PCT = predicted_means,
-  Lower_CI = predicted_intervals[1, ],
-  Upper_CI = predicted_intervals[2, ]
-)
-
-kable(head(result_summary, 10))
-```
-
-```{r}
-#| label: fig-bayesian-model-result-CI
-#| fig-cap: "U.S. Election Prediction Results"
-#| echo: false
-#| warning: false
-ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
-  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
-  labs(
-    title = "U.S. Election Prediction Results",
-    x = "Actual PCT",
-    y = "Predicted PCT"
-  )
-```
-
-@fig-bayesian-model-result-CI shows that he model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages.
-The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges.
-This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes.
-@tbl-bayesian-model-result-summary also highlights that most predicted values align closely with actual values, supporting the model’s robustness in capturing the overall trend of election outcomes.
-
-# Discussion {#sec-discussion}
-
-## Summary of Findings
-
-In this paper, we developed a hybrid election forecasting model to predict the 2024 U.S.
-presidential election results.
-Our model provides a stable and accurate forecast compared to traditional methods by combining poll-of-polls aggregation with additional factors such as poll score, sample size, days from the election, state, and methodology.
-
-Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability.
-The results shows that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.
-
-## Polling and Election Dynamics
-
-One of the main takeaways from this study is the impact of poll quality on election forecasts.
-We weight and filtered the polls by their reliability.
-This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.
-
-## Limitations of the Model {#sec-model-limitations}
-
-While this model presents a more comprehensive approach, there are limitations.
-First, Any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data.
-For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.
-
-Second, the model only reflect current voting dynamics as it is based on historical data.
-This model does not account for short-term variability such as Biden's exit.
-This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.
-
-On top of that, the results can be sensitive to the choice of priors.
-We choose normal(0, 2.5) as we believe it could stabilize the estimates by being informative enough to guide the posterior distribution without overpowering the data.
-However, if the chosen prior does not align well with the true underlying distribution, it may lead to biased results or overly wide credible intervals.
-
-Also, as it is a model based on a dataset, it is challenging to capture all the real-life features and dynamics.
-The model might not include all relevant predictors or interactions due to limitations in data availability or complexity, which can lead to incomplete representations of the factors influencing voting behavior.
-Additionally, without strong regularization techniques, the model may become prone to overfitting, particularly when using complex hierarchical structures or including numerous predictors.
-This overfitting can reduce the model’s generalizability to new or unseen data.
-
-Moreover, errors and offsets inherent in polling data, such as response bias, nonresponse adjustments, and sampling variability, can propagate into the model’s results.
-These aspects introduce an additional layer of uncertainty that can affect the model's reliability and predictive performance.
-While Bayesian methods provide a robust framework to incorporate uncertainty, the final outputs must be interpreted cautiously, acknowledging these underlying limitations.
-
-## Limitation of the Prediction
-
-In @sec-prediction, the team predicted that Kamala Harris would secure 298 delegates while Donald Trump would obtain 240 delegates.
-However, this prediction has several limitations.
-Firstly, the model used to generate these predictions inherits the constraints outlined in  @sec-model-limitations.
-Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.
-
-Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27.
-Moreover, as noted in Section @sec-prediction, the analysis focused solely on closely contested races, specifitatiocally swing states.
-Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions.
-Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.
-
-## Future Directions
-To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles.
-We includes previous year's polling dataset in this R project.
-By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics.
-This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion.
-Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, ultimately refining its accuracy and reliability for future applications in election forecasting.
-
-\newpage
-
-\appendix
+The remainder of this paper is structured as follows: [update this …]
 
 # Appendix {-}
+
+# Additional data details
+
+# Model details {#sec-model-details}
+
+## Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+## Diagnostics
+
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 set.seed(123)
-train_data_trump <- read_parquet(here("data/02-analysis_data/train_data_trump.parquet"))
-test_data_trump <- read_parquet(here("data/02-analysis_data/test_data_trump.parquet"))
+train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
+test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
 bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
@@
 #| warning: false
+
 just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
+
 predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
+
 predicted_means = colMeans(predictions2)
 predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+
 result_summary = data.frame(
@@
 )
+
 # kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")
+
 ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
-  geom_point(color = "blue", alpha = 0.7) +
-  # geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_point(alpha = 0.7) +
+  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
   geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
@@
 ```
-
 From @fig-model-trumpmlr it is clear that the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line).
 Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
 ```{r}
@@
 #| warning: false
+
 predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
+
 plot(test_data_trump$pct, colMeans(predictions2), 
@@
 abline(0, 1, col = "red")  # Add a 45-degree reference line
+
 ```
-
 Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
@@
 Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
 ```{r}
@@
 #| warning: false
+
 plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
 ```
-
 Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% confidence interval ranges from around 42.5% to 47.2%.
@@
 
-# Election Prediction {#sec-prediction}
+# Election Prediction
 
 Our prediction process consists of two primary components.
-First, we develop models for both Trump and Harris based on the variables outlined in @sec-model.
+First, we develop models for both Trump and Harris based on the variables outlined in Section @sec-PLACEHOLDER.
 This involves partitioning the dataset into training and testing subsets.
@@
 We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa.
-Winners for other states were determined based on historical trends and predictions from sources like @cnn_electoral_nodate.
+Winners for other states were determined based on historical trends and predictions from sources like @cnn.
 Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
-
 ```{r}
@@
 #| label: tbl-electionprediction
-#| tbl-cap: "Kamala Harris Wins Most of the Swing States"
+#| tbl-cap: Kamala Harris Wins Most of the Swing States
 # Split data into training and test sets
 
-train_data_harris <- read_parquet(here("data/02-analysis_data/train_data_harris.parquet"))
-test_data_harris <- read_parquet(here("data/02-analysis_data/test_data_harris.parquet"))
+train_data_harris <- read_csv(here("data/02-analysis_data/train_data_harris.csv"))
+test_data_harris <- read_csv(here("data/02-analysis_data/test_data_harris.csv"))
 bayesian_model_train_harris <- readRDS(here("models/bayesian_model_train_harris.rds"))
@@
             "Trump Predicted Percentage", "State Winner"))
+
 ```
@@
 #| label: fig-electionpredictionmap
-#| fig-cap: "Kamala Harris is Predicted to be the 47th President of the United States"
+#| fig-cap: Kamala Harris is Predicited to be the 47th President of the United States
 #| echo: false
@@
 ```
-
 According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States.
-There are also a few states with predictions not visible in the map, we will describe those predicts them below.
+There are also a few states with predictions not visible in the map, we will describe those predicts them below
 
@@
 
-# Overview of American Electon System
-
-## Brief Description of American Federal Goverment
-
-The American federal or national government is split into three branches, executive., legislative, and judical The executive branch includes the president and the military.
-The legislative branch include two subgroups, the House of Representatives and the Senate.
-These two subgroups create laws.
-Every state has two Senators and one Representative per approximately 750,000 people.
-The judicial branch is court system.
-
-## What is the Electoral College {#sec-electoral-college}
-
-The Electoral College is the system used in the United States to elect the president and vice president.
-Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives).
-When voters cast their ballots, they are actually voting for a slate of electors pledged to a candidate.
-The candidate who receives a majority of electoral votes (270 out of 538) wins the presidency.
-This system means that winning the popular vote in a state generally results in winning all of that state's electoral votes.
-The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.
-
-The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts.
-Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election.
-On the other hand, states that can vote either Democratic or Republican (swing states) are immensely important when predicting an election.
-As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there.
-
-```{r}
-#| label: fig-currentstateofelection
-#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
-#| echo: false
-#| warning: false
-light_blue <- "#ADD8E6"
-dark_blue <- "#6688CC"
-light_red <- "#FF7777"
-dark_red <- "#B22222"
-purple <- "#9370DB"
-# Define your colors for each state (replace these colors with your desired ones)
-state_colors_legend <- c(
-  "Republican" = dark_red,      # Dark Red
-  "Lean Republican" = light_red, # Light Red
-  "Toss-Up" = purple,         # Purple
-  "Lean Democrat" = light_blue,   # Light Blue
-  "Democrat" = dark_blue,        # Dark Blue
-  "Split By District" = "gray"
-)
-
-state_colors <- c(
-  alabama = dark_red,
-  arizona = purple,
-  arkansas = dark_red,
-  california = dark_blue,
-  colorado = dark_blue,
-  connecticut = dark_blue,
-  delaware = dark_blue,
-  `district of columbia` = dark_blue,
-  florida = light_red,
-  georgia = purple,
-  idaho = dark_red,
-  illinois = dark_blue,
-  indiana = dark_red,
-  iowa = light_red,
-  kansas = dark_red,
-  kentucky = dark_red,
-  louisiana = dark_red,
-  maine = "gray",
-  maryland = dark_blue,
-  massachusetts = dark_blue,
-  michigan = purple,
-  minnesota = light_blue,
-  mississippi = dark_red,
-  missouri = dark_red,
-  montana = dark_red,
-  nebraska = "gray",
-  nevada = purple,
-  `new hampshire` = light_blue,
-  `new jersey` = dark_blue,
-  `new mexico` = light_blue,
-  `new york` = dark_blue,
-  `north carolina` = purple,
-  `north dakota` = dark_red,
-  ohio = light_red,
-  oklahoma = dark_red,
-  oregon = dark_blue,
-  pennsylvania = purple,
-  `rhode island` = dark_blue,
-  `south carolina` = dark_red,
-  `south dakota` = dark_red,
-  tennessee = dark_red,
-  texas = light_red,
-  utah = dark_red,
-  vermont = dark_blue,
-  virginia = light_blue,
-  washington = dark_blue,
-  `west virginia` = dark_red,
-  wisconsin = purple,
-  wyoming = dark_red
-)
-
-# Load the states map data
-states_map <- map_data(map = "state")
-
-state_colors <- setNames(state_colors, tolower(names(state_colors)))
-
-color_to_status <- setNames(names(state_colors_legend), state_colors_legend)
-
-states_map <- states_map %>%
-  mutate(status = color_to_status[state_colors[region]])
-
-plotclr <- c(dark_red, light_blue, purple, gray, light_blue, dark_blue)
-legend_description <- c(
-  "Likely Republican",
-  "Lean Republican",
-  "Swing States",
-  "Lean Democrat",
-  "Likely Democrat",
-  "Split by District"
-)
-
-ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
-  geom_polygon(color = "white", size = 0.2) +
-  coord_map("albers", lat0 = 39, lat1 = 45) +
-  scale_fill_manual(
-    values = state_colors_legend,
-    name = "Election Status"
-  ) +
-  theme_minimal() +
-  theme(
-    legend.position = "bottom",
-    legend.title = element_text(size = 12, face = "bold"),
-    legend.text = element_text(size = 10),
-    axis.text = element_blank(),
-    axis.title = element_blank(),
-    panel.grid = element_blank()
-  )
-
-```
-
-The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are generally agreed upon within the American political community.
-These organizations assessed historical voting patterns and recent polling data to derive their conclusions.
-
-Notably, Nebraska and Maine are indicated in gray due to their delegates being split by district.
-In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic.
-In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican.
-Furthermore, Alaska and Hawaii are not in the map.
-Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.
-
-The seven generally agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.
-
-# States Poll Count
-
-```{r}
-#| label: tbl-state-polls-count
-#| tbl-cap: Polls included in Analysis Per State
-#| echo: false
-#| warning: false
-
-states <- c(state.name, "National")
-polls_per_state_tbl <- tibble(state = states)
-
-polls_per_state <- cleaned_data %>% group_by(state) %>% 
-  summarise(polls_count = n_distinct(poll_id))
-
-polls_per_state_tbl <- polls_per_state_tbl %>% 
-  left_join(polls_per_state, by = "state") %>% 
-  mutate(polls_count = if_else(!is.na(polls_count), polls_count, 0)) %>% 
-  select(state, polls_count)
-
-polls_per_state_tbl %>% kable(col.names = c("State", "Number of Polls In Analysis"))
-
-num_states_missing <- count(polls_per_state_tbl %>% 
-                              filter(polls_count == 0) %>% 
-                              select(state))
-
-```
-
-# The Ideal Survey
-
-## Objective
-
-The research team has developed a survey and distribution methodology with a hypothetical budget of $100,000.
-This objective is to create a polling system that accurately predicts the 2024 United States Election.
-
-## Sampling Frame
-
-Given the monetary constraint, the team's first decision is to mostly include swing states and districts.
-These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia.
-Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college.
-However, the poll will also include Texas, Florida, and Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.
-Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida and Minnesota.
-
-## Survey Sending Process
-
-The team will then find and use multiple databases, such as @usps_customer_nodate, to find addresses, telephones, and emails of potential voters in those states.
-According to @pew_research_center_us_nodate, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums.
-Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email.
-Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.
-
-## Sampling Methodology
-
-Each state will have its own poll, but the polling methodology and the survey given will remain the same.
-Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes.
-Ideally, the stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for.
-However, these factors are almost impossible to determine while sending the survey.
-Therefore, the research team decided to ask demographic questions directly inside the survey.
-After data collection, the team will weigh data based on demographics.
-For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.
-
-Furthermore, the team will aim to sample approximately 1,000 people from each true swing state.
-This sample size is large enough to provide reliable conclusions but not so large that it resembles sampling with replacement.
-For true swing states, the team aims to survey 1,000 individuals.
-However, if the final sample size falls short of this target, it is not considered a significant issue.
-
-## Survey Implementation & Question Creation
-
-The team's ideal survey will be made using @qualtrics_qualtrics_nodate.
-It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants.
-To identify participants who are not engaged, the research team has decided to add a worthless question.
-This question is extremely simple and clearly has one correct answer.
-Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data.
-An example of this is question number 7.
-
-Order bias occurs when the sequence of questions subconsciously influences participants' responses.
-To prevent this, many questions should be randomized upon entry to the survey.
-However, some questions must follow a specific order, such as question 11, while others like questions 1-4 can be randomized.
-The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics_qualtrics_nodate paid plan.
-
-Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently.
-To prevent this, answer choices should be randomized upon survey entry.
-Questions 8 & 9 could benefit from this.
-Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics_qualtrics_nodate paid plan.
-
-A leading question occurs when a question is written in a way that suggests the user to give a certain answer.
-For example, "given that children are the future of our country, should we invest more money in their education".
-To prevent this, the researchers have ensured questions are written in a style where no unnecessary details or opinions are added.
-
-## Survey Questions
-
-Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics_qualtrics_nodate survey.
-
-List of all of the questions:
-
-1. Select your race(s) (racial options where chosen based on @orvis_omb_2024)
-   - White
-   - Black or African American 
-   - American Indian or Alaska Native 
-   - Native Hawaiian or Pacific Islander 
-   - Middle Eastern or North African
-   - Asian 
-   - Other 
-   - Prefer not to answer
-
-2. Please select your gender 
-   - Male 
-   - Female 
-   - Non-binary 
-   - Prefer not to say
-
-3. Please enter your age (in numbers) 
-  - This is a text input field.
-Please note that this field has the auto-validation feature set to numbers in @qualtrics_qualtrics_nodate.
-As a result, participants can only input numbers in this field and are alert if they have not.
-
-4. Please select the highest degree of education you have obtained
-   - GED Certificate 
-   - High School Diploma 
-   - Undergraduate Degree 
-   - Graduate Degree (Masters/Phd) 
-   - None 
-   - Other
-   - Prefer not to answer
-
-5. Are you a registered voter for the 2024 United States Presidential Election? 
-   - Yes 
-   - No
-
-6. Place yourself on the political spectrum 
-   - Far Left 
-   - Center Left 
-   - Center 
-   - Center Right 
-   - Far Right
-
-7. Can pigs fly? 
-   - Yes 
-   - Maybe
-   - No 
-   - Prefer Not To Say
-
-8. What political party have you registered with? 
-   - Republicans 
-   - Democrats 
-   - Green Party 
-   - Libertarian 
-   - Other
-   - Independent (unregistered)
-
-9. Who will you vote for in the 2024 presidential election? 
-   - Democrat - Kamala Harris 
-   - Republican - Donald Trump 
-   - Green Party - Gill Stein 
-   - Libertarian - Chase Oliver 
-   - Other
-   - Will not vote
-
-## Potential Problems with the Methodology And Polls
-
-While the team's methodology and survey creates a robust system, there are potential issues.
-The reliance of weighting results based on a candidate's demographics can lead to error propagation.
-For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have a large impact on the polls prediction of what candidate will win the state.
-There is also a selection bias in terms of the monetary reward.
-Potentially, people who like monetary rewards could be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.
-
-# Methodology of YouGov
-YouGov's methodology documentations are seperated in two articles.
-The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
-
-## Population, Frame, and Sample
-As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
-The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
-Voter files indicates whether someone voted in a given election, thus YouGov's population covers all voters in previous US elections.
-
-YouGov’s sampling frame consists of its online panel members.
-These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
-YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
-
-YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
-YouGov interviews nearly 100,000 people in the first set of estimates.
-For the second set of estimates, YouGov didn't just start over with a new sample.
-They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
-
-## Sample Recruitment
-Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
-They must provide demographic details upon joining, which helps in selecting representative samples for each survey.
-When respondents complete a survey, they are awarded points that can be exchanged for money.
-
-## Sampling Approach and Trade-offs
-YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
-This method allows quick and cost-effective data collection.
-However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
-YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
-
-## Non-response Handling
-YouGov apply statistical weighting to adjust for the differences between the sample and target population.
-The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
-Additionally, quality control measures exclude unreliable responses to improve data accuracy.
-The respondents are offered a small incentive to decrease the non-response and increase participation.
-
-## Strengths and Weaknesses of the Questionnaire
-YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
-The pollster can recruit a large amount of panelists because of the online format.
-Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
-
-As a non-probability sample, it might miss certain demographic groups not covered by the online population.
-While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
-Additionaly, the categories in the surbey is oversimplified with bias.
-For instance, in the poll result published by @yougov_say24_nodate, gender is divided into Male and Female.
-Race is divided into White, Black, Hispanic and Other.
-This indicates a lack of representation.
-
 \newpage
 
+# Sources of Bias In the Polls
+
+Based on the figures in section @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores.
+From the analysis in Figure @fig-methodologyboxplot, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability.
+This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys.
+Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias.
+Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.
+
+From Figure @fig-daysfromelection-pct, the team observed a decline in third-party candidate support as the election approached.
+Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes.
+This trend may be worsened by decreased participation in polls as the election date approaches.
+As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates.
+Additionally, the withdrawal of a prominent third-party candidate, Robert F.
+Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic.
+However, we find the earlier reasons more compelling, as the data in Figure @fig-daysfromelection-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.
+
+As illustrated in Figure @fig-pollscore-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls.
+It is important to note that the 538 poll score system is partly based on historical accuracy.
+Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies.
+This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.
+
+## Limitations of the Model
+
+While this model presents a more comprehensive approach, there are limitations.
+First, Any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data.
+For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.
+
+Second, the model only reflect current voting dynamics as it is based on historical data.
+This model does not account for short-term variability such as Biden's exit.
+This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.
+
+On top of that, the results can be sensitive to the choice of priors.
+We choose normal(0, 2.5) as we believe it could stabilize the estimates by being informative enough to guide the posterior distribution without overpowering the data.
+However, if the chosen prior does not align well with the true underlying distribution, it may lead to biased results or overly wide credible intervals.
+
+Also, as it is a model based on a dataset, it is challenging to capture all the real-life features and dynamics.
+The model might not include all relevant predictors or interactions due to limitations in data availability or complexity, which can lead to incomplete representations of the factors influencing voting behavior.
+Additionally, without strong regularization techniques, the model may become prone to overfitting, particularly when using complex hierarchical structures or including numerous predictors.
+This overfitting can reduce the model’s generalizability to new or unseen data.
+
+Moreover, errors and offsets inherent in polling data, such as response bias, nonresponse adjustments, and sampling variability, can propagate into the model’s results.
+These aspects introduce an additional layer of uncertainty that can affect the model's reliability and predictive performance.
+While Bayesian methods provide a robust framework to incorporate uncertainty, the final outputs must be interpreted cautiously, acknowledging these underlying limitations.
+
+# Limitations of the Prediction
+
+In section @sec-prediction, the team predicted that Kamala Harris would secure 298 delegates while Donald Trump would obtain 240 delegates.
+However, this prediction has several limitations.
+Firstly, the model used to generate these predictions inherits the constraints outlined in section @sec-model-limitations.
+Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.
+
+Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27.
+Moreover, as noted in Section @sec-prediction, the analysis focused solely on closely contested races, specifically swing states.
+Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions.
+Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.
+
 # References

### 2024-11-04 paper/paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
 
-```{r}
-#| label: fig-explore-data
-#| fig-cap: "Exploratory Analysis of Predictors and Candidate Support (pct)"
-#| echo: false
-#| warning: false
-a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
-  geom_point() + 
-  geom_smooth(method = lm, se = FALSE) + 
-  labs(x = 'sample size', y='pct', 
-       title = 'pct vs sample size')
-
-b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
-  geom_point() + 
-  geom_smooth(method = lm, se = FALSE) + 
-  labs(x = 'pollscore', y='pct', 
-       title = 'pct vs pollscore')
-
-c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
-  geom_boxplot() + 
-  labs(x = 'methodology', y='pct', 
-       title = 'pct vs methodology')
-
-d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
-  geom_point() + 
-  geom_smooth(method = lm, se = FALSE) + 
-  labs(x = 'max gust speed', y='pct', 
-       title = 'pct vs days taken from election')
-
-e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
-  geom_boxplot() + 
-  labs(x = 'state', y='pct', 
-       title = 'pct vs state')
-
-grid.arrange(a, b, c, d, e, nrow=2)
-```
-
-@fig-explore-data explores the relationship between candidate support percentage (pct) and various predictors in the dataset: (a) sample size of polls, (b) pollscore (an indicator of poll reliability), (c) polling methodology, (d) days taken from the election, and (e) state.
-
 ## Clean Data
@@
 We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots.
-As shown in @fig-explore-data in @sec-dataexplore, The relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy.
+As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, The relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy.
 The plot of pct versus poll score reveals a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct.
@@
 This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.
+
+## Sources of Bias In the Polls
+
+Based on the figures in @sec-data, the team identified three sources of bias beyond sample size: methodology, the timing of polls relative to the election date, and poll scores.
+From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability.
+This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys.
+Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias.
+Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.
+
+From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached.
+Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes.
+This trend may be worsened by decreased participation in polls as the election date approaches.
+As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates.
+Additionally, the withdrawal of a prominent third-party candidate, Robert F.
+Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic.
+However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.
+
+Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls.
+It is important to note that the 538 poll score system is partly based on historical accuracy.
+Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies.
+This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.
 

### 2024-11-04 paper/paper.qmd
fix grammar

@@
 Our results show that Kamala Harris holds a slight lead over Donald Trump in most swing states, with support levels stabilizing closer to election day.
-This approach underscores the importance of nuanced forecasting models in election analysis, revealing how poll quality and timing can significantly influence support predictions.
-Ultimately, this model offers a comprehensive tool for understanding the dynamics of voter sentiment and improving the accuracy of election predictions."
+This approach reveals how poll quality and timing can significantly influence support predictions.
+This model offers a comprehensive tool for understanding the dynamics of voter sentiment and improving the accuracy of election predictions.
+"
 format: pdf
@@
 
-The remainder of this paper is structured as follows: @sec-data provides an overview and exploration of the data.
+The remainder of this paper is structured as follows: @sec-data provides an overview of the data.
 @sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model.
@@
 According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024].
-The dataset contains a list of questions in polls and surveys and their results.
+The dataset has a list of questions in polls and surveys and their results.
 When a new poll is conducted, the poll is appended to the dataset.
@@
 In this dataset, each row represents a polling question that records the variables of interest.
-Each entry allows us to explore the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
-This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused.
+Each entry allows us to examine the real-world relationships between polling factors and the support percentage (`pct`) for the candidates Kamala Harris and Donald Trump.
+This dataset enables an analysis of how various polling characteristics influence the reported support levels for the candidates we are focused on.
 
@@
 
-## Data Exploration {#sec-dataexplore}
-
-The raw data from FiveThirtyEight contains 52 columns, all of the column headers are displayed below: 
+## Data Examination {#sec-dataexplore}
+
+The raw data from FiveThirtyEight contains 52 columns, all the column headers are displayed below: 
 ```{r}
@@
 
-These columns can be categorized into three types, response variable, numeric predictors and categorical predictors.
+These columns can be categorized into three types, response variable, numeric predictors, and categorical predictors.
 
@@
 
-The first 6 rows of the dataset is displayed in @tbl-cleaned-data.
+The first 6 rows of the dataset are displayed in @tbl-cleaned-data.
 
@@
 @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll.
-The figure also shows that most polls have a sample size around 800 to 1200 participants.
+The figure also shows that most polls have a sample size of around 800 to 1200 participants.
 It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.
@@
 Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election.
-This is a result of non-major candidate support rapidly decreasing.
+This is a result of non-major candidate support decreasing.
 Specifically, Trump support is increasing at a faster rate than Harris.
 However, the polls in general show a slight lead for Harris throughout the last 100 days.
-
-The rapid decline in third party support could be due to Robert F. Kennedy dropping out of the race (this sentences should probably be in results or discussion section).
 
@@
 
-The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:
+The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. We run the model with the following specifications:
 
@@
 
-We first set up a SLR model for predicting pct based on pollscore. The model is:
+We first set up an SLR model for predicting pct based on pollscore. The model is:
 
@@
 This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct.
-The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.
+The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is not sufficiently captured by a linear model with just one predictor.
 
@@
 In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor.
-This indicate that sample_size has limited impact on pct.
+This indicates that sample_size has limited impact on pct.
 
@@
 
-Next, we predicts pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:
+Next, we predict pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:
 
@@
 The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
-The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60).
-This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.
+Most of the data points align well with the red dashed line, particularly within the middle range of pct values (between 40 and 60).
+This alignment indicates that the model performs well in this range, with predicted values correlating strongly with the actual observed outcomes.
 
@@
 The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides.
-The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
+The replicated distributions (y_rep), shown as multiple thin lines, align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data.
 However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.
 
-The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data.
+The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs well in replicating the observed data.
 Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy.
@@
 
-In addition, the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
-
-In conculsion, the visual checks from the predicted vs.
-actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting.
+In addition, the residual plot @fig-bayesian-model-residual is fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.
+
+In conclusion, from the visual checks from the predicted vs.
+actual plot and the residual plot, there is no robust evidence that the Bayesian model is overfitting.
 It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.
@@
 We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots.
-As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, The relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy.
-The plot of pct versus poll score reveals a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct.
+As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, the relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy.
+The plot of pct versus poll score shows a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct.
 Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value.
@@
 
-To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
+To predict the support percentage (PCT) for Harris and Trump, we employed a modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model.
 The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity.
@@
 # Result {#sec-result}
+
+## Harris Voter Prediction Model 
 ```{r}
@@
 The estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S.
-election data analyzed, is centered around 52%, with a credible interval spanning approximately from 49.5% to 53.5%.
-The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
+election data analyzed, is centered around 52%, with a credible interval spanning from 49.5% to 53.5%.
+The narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.
 
@@
 
-@fig-bayesian-model-result-CI shows that he model's predictive performance for Harris's election data appears generally robust, demonstrating a strong alignment between actual and predicted vote percentages.
-The majority of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges.
+@fig-bayesian-model-result-CI shows that the model's predictive performance for Harris's election data appears robust, demonstrating a strong alignment between actual and predicted vote percentages.
+Most of the data points cluster around the line, indicating reliable predictions, some variance is evident, particularly in the lower and higher actual percentage ranges.
 This suggests that while the model performs well on average, it may have larger uncertainty in its predictions at the extremes.
@@
 
-# Discussion {#sec-discussion}
-
-## Summary of Findings
-
-In this paper, we developed a hybrid election forecasting model to predict the 2024 U.S.
-presidential election results.
-Our model provides a stable and accurate forecast compared to traditional methods by combining poll-of-polls aggregation with additional factors such as poll score, sample size, days from the election, state, and methodology.
-
-Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability.
-The results shows that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.
-
-## Polling and Election Dynamics
-
-One of the main takeaways from this study is the impact of poll quality on election forecasts.
-We weight and filtered the polls by their reliability.
-This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.
-
-## Sources of Bias In the Polls
-
-Based on the figures in @sec-data, the team identified three sources of bias beyond sample size: methodology, the timing of polls relative to the election date, and poll scores.
-From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability.
-This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys.
-Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias.
-Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.
-
-From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached.
-Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes.
-This trend may be worsened by decreased participation in polls as the election date approaches.
-As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates.
-Additionally, the withdrawal of a prominent third-party candidate, Robert F.
-Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic.
-However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.
-
-Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls.
-It is important to note that the 538 poll score system is partly based on historical accuracy.
-Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies.
-This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.
-
-## Limitations of the Model {#sec-model-limitations}
-
-While this model presents a more comprehensive approach, there are limitations.
-First, Any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data.
-For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.
-
-Second, the model only reflect current voting dynamics as it is based on historical data.
-This model does not account for short-term variability such as Biden's exit.
-This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.
-
-On top of that, the results can be sensitive to the choice of priors.
-We choose normal(0, 2.5) as we believe it could stabilize the estimates by being informative enough to guide the posterior distribution without overpowering the data.
-However, if the chosen prior does not align well with the true underlying distribution, it may lead to biased results or overly wide credible intervals.
-
-Also, as it is a model based on a dataset, it is challenging to capture all the real-life features and dynamics.
-The model might not include all relevant predictors or interactions due to limitations in data availability or complexity, which can lead to incomplete representations of the factors influencing voting behavior.
-Additionally, without strong regularization techniques, the model may become prone to overfitting, particularly when using complex hierarchical structures or including numerous predictors.
-This overfitting can reduce the model’s generalizability to new or unseen data.
-
-Moreover, errors and offsets inherent in polling data, such as response bias, nonresponse adjustments, and sampling variability, can propagate into the model’s results.
-These aspects introduce an additional layer of uncertainty that can affect the model's reliability and predictive performance.
-While Bayesian methods provide a robust framework to incorporate uncertainty, the final outputs must be interpreted cautiously, acknowledging these underlying limitations.
-
-## Limitation of the Prediction
-
-In @sec-prediction, the team predicted that Kamala Harris would secure 298 delegates while Donald Trump would obtain 240 delegates.
-However, this prediction has several limitations.
-Firstly, the model used to generate these predictions inherits the constraints outlined in  @sec-model-limitations.
-Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.
-
-Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27.
-Moreover, as noted in Section @sec-prediction, the analysis focused solely on closely contested races, specifitatiocally swing states.
-Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions.
-Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.
-
-## Future Directions
-To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles.
-We includes previous year's polling dataset in this R project.
-By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics.
-This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion.
-Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, ultimately refining its accuracy and reliability for future applications in election forecasting.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# FiveThirtyEight Licenses
-[FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).
-
-# Trump Voter Prediction Model
+## Trump Voter Prediction Model
 
@@
 )
-# kable(head(result_summary, 10), caption = "Head of Bayesian Model Result Summary")
 ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
-  geom_point(color = "blue", alpha = 0.7) +
-  # geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_point(alpha = 0.7) +
+  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
   geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
@@
 
-From @fig-model-trumpmlr it is clear that the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line).
+```{r}
+#| label: tbl-model-trumpmlr
+#| tbl-cap: "Head of Bayesian Model Result Summary"
+#| echo: false
+#| warning: false
+
+kable(head(result_summary, 10))
+
+```
+
+From @fig-model-trumpmlr and @tbl-model-trumpmlr the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line).
 Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
 ```{r}
@@
 Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
-This suggests that model can generalization to outside data.
+This suggests that model can generalize outside data.
 Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
@@
 
-# Election Prediction {#sec-prediction}
+## Election Prediction {#sec-prediction}
 
@@
 By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state.
-The candidate with the higher percentage is deemed the winner for that state.
+The candidate with a higher percentage is deemed the winner for that state.
 
@@
 According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States.
-There are also a few states with predictions not visible in the map, we will describe those predicts them below.
+There are also a few states with predictions not visible on the map, we will describe those predictions below.
 
@@
 
+# Discussion {#sec-discussion}
+
+## Summary of Findings
+
+In this paper, we developed a hybrid election forecasting model to predict the 2024 U.S.
+presidential election results.
+Our model provides a stable and accurate forecast compared to traditional methods by combining poll-of-polls aggregation with additional factors such as poll score, sample size, days from the election, state, and methodology.
+
+Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability.
+The results show that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.
+
+## Polling and Election Dynamics
+
+One of the main takeaways from this study is the impact of poll quality on election forecasts.
+We weight and filter the polls by their reliability.
+This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.
+
+## Sources of Bias in the Polls
+
+Based on the figures in @sec-data, the team identified three sources of bias beyond sample size: methodology, the timing of polls relative to the election date, and poll scores.
+From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability.
+This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys.
+Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias.
+Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.
+
+From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached.
+Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes.
+This trend may be worsened by decreased participation in polls as the election date approaches.
+As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates.
+Additionally, the withdrawal of a prominent third-party candidate, Robert F.
+Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic.
+However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.
+
+Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls.
+It is important to note that the FiveThirtyEight poll score system is partly based on historical accuracy.
+Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies.
+This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.
+
+## Limitations of the Model {#sec-model-limitations}
+
+First, any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data.
+For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.
+
+Second, the model only reflects current voting dynamics as it is based on historical data.
+This model does not account for short-term variability such as Biden's exit.
+This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.
+
+On top of that, the results can be sensitive to the choice of priors.
+We choose Normal(0, 2.5) as we believe it could stabilize the estimates by being informative enough to guide the posterior distribution without overpowering the data.
+However, if the chosen prior does not align well with the true underlying distribution, it may lead to biased results or overly wide credible intervals.
+
+Also, as it is a model based on a dataset, it is challenging to capture all the real-life features and dynamics.
+The model might not include all relevant predictors or interactions due to limitations in data availability or complexity, which can lead to incomplete representations of the factors influencing voting behavior.
+Additionally, without strong regularization techniques, the model may become prone to overfitting, particularly when using complex hierarchical structures or including numerous predictors.
+This overfitting can reduce the model’s generalizability to new or unseen data.
+
+Moreover, errors and offsets inherent in polling data, such as response bias, nonresponse adjustments, and sampling variability, can propagate into the model’s results.
+These aspects introduce an additional layer of uncertainty that can affect the model's reliability and predictive performance.
+While Bayesian methods provide a robust framework to incorporate uncertainty, the final outputs must be interpreted cautiously, acknowledging these underlying limitations.
+
+## Limitation of the Prediction
+
+In @sec-prediction, the team predicted that Kamala Harris would secure 298 delegates while Donald Trump would obtain 240 delegates.
+However, this prediction has several limitations.
+Firstly, the model used to generate these predictions inherits the constraints outlined in @sec-model-limitations.
+Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.
+
+Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27.
+Moreover, as noted in Section @sec-prediction, the analysis focused solely on contested races, specifically swing states.
+Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions.
+Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.
+
+## Future Directions
+
+To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles.
+We include previous year's polling dataset in this R project.
+By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics.
+This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion.
+Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, refining its accuracy and reliability for future applications in election forecasting.
+
+\newpage
+
+\appendix
+
+# Appendix {-}
+
+# FiveThirtyEight Licenses
+[FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).
+
 # Overview of American Electon System
 
-## Brief Description of American Federal Goverment
-
-The American federal or national government is split into three branches, executive., legislative, and judical The executive branch includes the president and the military.
-The legislative branch include two subgroups, the House of Representatives and the Senate.
+## Brief Description of American Federal Government
+
+The American federal or national government is split into three branches, executive., legislative, and judicial the executive branch includes the president and the military.
+The legislative branch includes two subgroups, the House of Representatives, and the Senate.
 These two subgroups create laws.
 Every state has two Senators and one Representative per approximately 750,000 people.
-The judicial branch is court system.
+The judicial branch is the court system.
 
@@
 Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives).
-When voters cast their ballots, they are actually voting for a slate of electors pledged to a candidate.
-The candidate who receives a majority of electoral votes (270 out of 538) wins the presidency.
-This system means that winning the popular vote in a state generally results in winning all of that state's electoral votes.
+When voters vote, they are voting for a slate of electors pledged to a candidate.
+The candidate who receives the most electoral votes (270 out of 538) wins the presidency.
+This system means that winning the popular vote in a state result in winning all that state's electoral votes.
 The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.
@@
 As a result, campaigns spend hundreds of millions of dollars campaigning and understanding voters there.
-
 ```{r}
@@
 
-The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are generally agreed upon within the American political community.
+The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are agreed upon within the American political community.
 These organizations assessed historical voting patterns and recent polling data to derive their conclusions.
@@
 In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican.
-Furthermore, Alaska and Hawaii are not in the map.
+Furthermore, Alaska and Hawaii are not on the map.
 Alaska is strongly favoring Republicans while Hawaii strongly favors democrats.
 
-The seven generally agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.
+The seven agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.
 
@@
 
-# The Ideal Survey
+#The Ideal Survey
 
@@
 Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college.
-However, the poll will also include Texas, Florida, and Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.
-Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida and Minnesota.
+However, the poll will also include Texas, Florida, Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.
+Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida, and Minnesota.
 
@@
 Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes.
-Ideally, the stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for.
+Ideally, stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for.
 However, these factors are almost impossible to determine while sending the survey.
@@
 This question is extremely simple and clearly has one correct answer.
-Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data.
+Therefore, if a participant selects the wrong answer, they are not engaged with the survey and their responses should be removed from the final data.
 An example of this is question number 7.
@@
 
-## Potential Problems with the Methodology And Polls
-
-While the team's methodology and survey creates a robust system, there are potential issues.
+## Potential Problems with the Methodology and Polls
+
+While the team's methodology and survey create a robust system, there are potential issues.
 The reliance of weighting results based on a candidate's demographics can lead to error propagation.
-For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have a large impact on the polls prediction of what candidate will win the state.
+For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have impact on the polls prediction of what candidate will win the state.
 There is also a selection bias in terms of the monetary reward.
-Potentially, people who like monetary rewards could be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.
+Potentially, people who like monetary rewards would be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.
 
 # Methodology of YouGov
-YouGov's methodology documentations are seperated in two articles.
+
+YouGov's methodology documentation is separated into two articles.
 The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
@@
 ## Population, Frame, and Sample
-As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel.
-The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
-Voter files indicates whether someone voted in a given election, thus YouGov's population covers all voters in previous US elections.
+
+As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether they belong to YouGov's panel.
+The national voter files are digital databases built by commercial organizations with public government records of voters, as explained by @desilver_q_2018.
+Voter files indicate whether someone voted in each election, thus YouGov's population covers all voters in previous US elections.
 
@@
 These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024.
-YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.
+YouGov collects information on respondents when they join their panel before they are invited to participate in the survey.
 
 YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest.
-YouGov interviews nearly 100,000 people in the first set of estimates.
-For the second set of estimates, YouGov didn't just start over with a new sample.
+YouGov interviews 100,000 people in the first set of estimates.
+For the second set of estimates, YouGov did not just start over with a new sample.
 They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.
@@
 ## Sample Recruitment
+
 Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate].
@@
 ## Sampling Approach and Trade-offs
+
 YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate].
@@
 However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate.
-YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.
+YouGov states that 95% of the US population has internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with slow internet access or without internet access.
 
 ## Non-response Handling
-YouGov apply statistical weighting to adjust for the differences between the sample and target population.
-The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate].
+
+YouGov applied statistical weighting to adjust the differences between the sample and target population.
+The weight is based on demographic characteristics such as age, gender, race, and presidential vote [@yougov_methodology_nodate].
 Additionally, quality control measures exclude unreliable responses to improve data accuracy.
@@
 ## Strengths and Weaknesses of the Questionnaire
+
 YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness.
-The pollster can recruit a large amount of panelists because of the online format.
-Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.
+The pollster can recruit many panelists because of the online format.
+Combining with online tracking technologies, the metadata provided by their panelists can be verified easily.
 
 As a non-probability sample, it might miss certain demographic groups not covered by the online population.
-While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​.
-Additionaly, the categories in the surbey is oversimplified with bias.
+While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling.
+Additionally, the categories in the survey are oversimplified with bias.
 For instance, in the poll result published by @yougov_say24_nodate, gender is divided into Male and Female.
-Race is divided into White, Black, Hispanic and Other.
+Race is divided into White, Black, Hispanic, and Other.
 This indicates a lack of representation.

### 2024-11-04 paper/paper.qmd
fix table and figure headings

@@
 #| label: fig-model-pct-pollscore
-#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
+#| fig-cap: "Pollscore has limited predictive accuracy for Kamala's voter percentage."
 #| echo: false
@@
   geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
+  labs(title = "Single Linear Model (Only Pollscore Predictor) For Harris",
+       x = "Actual Vote Percentage",
+       y = "Predicted Vote Percentage") +
   theme_minimal()
@@
 #| label: fig-model-pct-sample_size
-#| fig-cap: "Actual vs. Predicted Values for pct Using sample_size as Predictor"
+#| fig-cap: "Sample size has Limited Impact on Kamala's Voter Percentage"
 #| echo: false
@@
   geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
+  labs(title = "Sample Size as Predictor Model for Harris",
+       x = "Actual Vote Percentage",
+       y = "Predicted Vote Percentage") +
   theme_minimal()
@@
 
-In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor.
-This indicates that sample_size has limited impact on pct.
+In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample size is the only predictor.
+This indicates that sample size has limited impact on pct.
 
@@
 #| label: fig-model-MLR
-#| fig-cap: "Actual vs Predicted Percentage Support for Kamala Harris (Multiple Predictors)"
+#| fig-cap: "The MLR Model captured significant variance in Kamala Harris' voter percentage"
 #| echo: false
@@
   geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual pct",
-       y = "Predicted pct") +
+  labs(title = "Multiple Predictors Model For Kamala Harris",
+       x = "Actual Voter Percentage",
+       y = "Predicted Voter Percentage") +
   theme_minimal()
@@
 
-@fig-model-MLR compares actual pct values with the predicted pct values from the MLR model.
+@fig-model-MLR compares actual percentage values with the predicted pct values from the MLR model.
 Blue points represent individual comparisons between actual and predicted values.
@@
 
-The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct.
+The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in Kamala Harris' voter percentage.
 Most of the data points align well with the red dashed line, particularly within the middle range of pct values (between 40 and 60).
@@
 #| label: fig-bayesian-model-actual-predict
-#| fig-cap: "Predicted vs Actual Percentage Support (pct) for Kamala Harris in the Bayesian Model"
+#| fig-cap: "The MLR Model Accruatley Predicts the Testing Dataset"
 #| echo: false
@@
 plot(test_data$pct, colMeans(predictions), 
-     xlab = "Actual PCT Values", 
-     ylab = "Predicted  PCT Values", 
-     main = "Predicted PCT vs Actual PCT")
-abline(0, 1, col = "red")  # Add a 45-degree reference line
+     xlab = "Actual Voter Percentage Values", 
+     ylab = "Predicted Voter Percentage Values", 
+     main = "Multiple Predictors Kamala Harris Bayesian Model")
+abline(0, 1, col = "red")
 ```
@@
 #| label: fig-bayesian-model-residual
-#| fig-cap: "Residual Plot for the Bayesian Hierarchical Model Predictions"
+#| fig-cap: "MLR Residuals Appear to Show No Trends"
 #| echo: false
@@
      ylab = "Residuals", 
-     main = "Residual Plot")
+     main = "Residual for the Bayesian Hierarchical Model Predictions")
 abline(h = 0, col = "red")
@@
 #| label: fig-bayesian-model-CI
-#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
+#| fig-cap: "Kamala Harris is Predicted to get Approximatley 52% of Popular Vote"
 #| echo: false
@@
 # Plot random effects
-plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95, 
+     xlab = "Kamala Harris Popular Vote Percentage",
+     title = "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model")
 ```
@@
 
-kable(head(result_summary, 10))
+kable(head(result_summary, 10), 
+      col.names = c("Acutal Kamala Percentage", 
+                    "Predicted Percentage", 
+                    "Lower Bound Credible Interval",
+                    "Upper bound Credible Interval"))
 ```
@@
 #| label: fig-bayesian-model-result-CI
-#| fig-cap: "U.S. Election Prediction Results"
+#| fig-cap: "The Model Shows a Strong Alignment Between Actual and Predicted Vote Percentages"
 #| echo: false
@@
     title = "U.S. Election Prediction Results",
-    x = "Actual PCT",
-    y = "Predicted PCT"
+    x = "Actual Percentage",
+    y = "Predicted Percentage"
   )
@@
 
-## Trump Voter Prediction Model
+# Trump Voter Prediction Model
 
@@
 Likewise, the Trump dataset is also split into training and testing data.
-The model outputs are below.
+The model results are similar to that of Harris's MLR model.
 
@@
 ```
+
+```{r}
+#| label: fig-model-trumpmlr-res
+#| fig-cap: "Bayesian Hierarchical MLR Model of Trump Voting Percentage Captures Variability Without Overfitting"
+#| fig-subcap: ["Trump Predictions (Test Data)", "Residuals for Predictions"]
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+
+predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
+plot(test_data_trump$pct, colMeans(predictions2), 
+     xlab = "Actual Voter Percentage Values", 
+     ylab = "Predicted Voter Percentage Values")
+abline(0, 1, col = "red")
+
+residuals <- test_data$pct - colMeans(predictions)
+plot(colMeans(predictions), residuals, 
+     xlab = "Predicted Values", 
+     ylab = "Residuals")
+abline(h = 0, col = "red")
+```
+
+Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
+This suggests that model can generalize outside data.
+Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
+# Trump Model Results
+
+```{r}
+#| label: fig-trumpmodelci
+#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
+#| echo: false
+#| warning: false
+plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
+```
+
+Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% credible interval ranges from around 42.5% to 47.2%.
+This credible interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.
 
@@
   labs(
-    title = "Actual vs Predicted Percentage For Trump MLR Model (All Trump Data)",
+    title = "Actual vs Predicted Percentage For Trump MLR Model",
     x = "Actual PCT",
@@
 
-kable(head(result_summary, 10))
+kable(head(result_summary, 10), col.names = c("Acutal Harris Percentage", 
+                    "Predicted Percentage", 
+                    "Lower Bound Credible Interval",
+                    "Upper bound Credible Interval"))
 
@@
 
-```{r}
-#| label: fig-model-trumpmlr-res
-#| fig-cap: "MLR Trump Model does not Appear to Overfit"
-#| echo: false
-#| warning: false
-predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
-plot(test_data_trump$pct, colMeans(predictions2), 
-     xlab = "Actual PCT Values", 
-     ylab = "Predicted  PCT Values", 
-     main = "Predicted PCT vs Actual PCT (Test Data)")
-abline(0, 1, col = "red")  # Add a 45-degree reference line
-```
-
-Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
-This suggests that model can generalize outside data.
-Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
-```{r}
-#| label: fig-trumpmodelci
-#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
-#| echo: false
-#| warning: false
-plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
-```
-
-Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% confidence interval ranges from around 42.5% to 47.2%.
-This confidence interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.
+\newpage
 
@@
 
-Based on the figures in @sec-data, the team identified three sources of bias beyond sample size: methodology, the timing of polls relative to the election date, and poll scores.
+Based on the figures in @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores.
 From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability.
@@
 
-#The Ideal Survey
+# The Ideal Survey
 
@@
 Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes.
+County proportion is used as the stratification variable because it is easily accessible.
+Additionally, counties possess inherent demographic and cultural characteristics that can influence their alignment with particular political parties.
 Ideally, stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for.

### 2024-11-04 paper/paper.qmd
update paper table and graph
made text understandability edits

@@
 Our results show that this model provides a more stable and accurate forecast than traditional poll aggregation alone.
+With this model, we generate a prediction that Kamala Harris will win the popular vote in 2024 and win the electoral college.
 
@@
 @sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model.
-We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.
+We then present our results in @sec-result and discuss the implications, limitations, then we predict the election in @sec-prediction, and future research directions in @sec-discussion.
 
@@
 The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump.
-The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.
+The prediction is based on quantifying various polling factors, including sample size, poll scores, and poll methodology, which are used as predictors.
 
@@
 
-- pct: The percentage of support or vote share that each candidate (Kamala Harris or Donald Trump) received in the poll.
+- pct: The percentage of support or vote share that each candidate (Kamala Harris, Donald Trump, or third party alternatives) received in the poll.
 
@@
 - party: The political party of the candidate (e.g., DEM for Democrat, REP for Republican).
-- candidate_name: The name of the candidate, either Kamala Harris or Donald Trump.
+- candidate_name: The name of the candidate (i.e. Kamala Harris, Donald Trump, or third party candidates).
 - pollster: The name of the polling organization that conducted the poll.
@@
 
-We then create a new attribute, days_taken_from_election, which represents the time gap between the poll's start date and the actual U.S.
+We then create a new attribute, days taken from election, which represents the time gap between the poll's start date and the actual U.S.
 election date.
@@
 
-This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix).
+This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in @sec-electoral-college).
 The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.
@@
 
-### Basic Statistics Summary for Cleaned Data
+### Relationship Between Variables
+
+#### State and Candidate Percentage
 
@@
   theme(legend.position = "bottom") + 
-  theme(axis.text.y = element_text(size = 8))
-
-```
-
-In @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump.
+  theme(axis.text.y = element_text(size = 5))
+
+```
+
+In @fig-avg-pct-state, historically Democratic states are polling for Kamala and historically Republican states are polling for Trump.
 Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.
+
+#### Poll Methodology and Candidate Percentage
 
@@
 @fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.
-
-### Relationship Between Variables
+It is important to note that statistical outliers are not visible in the graph to ensure better visibility.
+
+#### Poll Score and Sample Size
 
@@
 #| label: fig-pollscore-samplesize
-#| fig-cap: More Reliable Pollsters Have Larger Sample Sizes in their Polls
+#| fig-cap: Pollsters with Higher Pollscores Have Larger Sample Sizes in their Polls
 #| echo: false
@@
 It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.
+
+#### Sample Size and Voter Percentage
 
@@
 
-Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the percentage of a candidate.
+Based on @fig-pollsamplesize-canidate-pct, there does not seem to be a relationship between the sample size of a poll the voter percentage of a candidate.
+
+#### Poll Score and Voter Percentage
 
@@
 This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.
+
+#### Days from Election and Voter Percentage
 
@@
 We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots.
-As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, the relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy.
-The plot of pct versus poll score shows a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct.
-Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value.
+As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, the relationship between percentage and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy.
+The plot of percentage versus poll score shows a negative trend for Donald Trump and a positive trend for Kamala Harris, suggesting that poll scores are consistently influential in explaining variations in percentage.
+Additionally, the percentage versus days taken plot indicates that the proximity to the election affects the percentage, with a slight boost towards major candidates.
 These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.
@@
 By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.
+
+The team also built a Bayesian MLR using the same approach and formula for Trump. It is in section @sec-trumpmodel.
 
@@
 
-# Trump Voter Prediction Model
-
-The multiple linear regression model (MLR) for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris.
-Likewise, the Trump dataset is also split into training and testing data.
-The model results are similar to that of Harris's MLR model.
-
+The team has provided results for a Trump Bayesian MLR model. It is in section @sec-trumpmodelresults.
+
+\newpage
+
+# Election Prediction {#sec-prediction}
+
+Our prediction process consists of two primary components.
+First, we develop models for both Trump and Harris based on the variables outlined in @sec-model and @sec-trumpmodel.
+This involves partitioning the dataset into training and testing subsets.
+Next, we further divide the testing dataset by state.
+We then input this test data into the respective models to generate predictions.
+By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state.
+The candidate with a higher percentage is deemed the winner for that state.
+
+We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa.
+Winners for other states were determined based on historical trends and predictions from sources like @cnn_electoral_nodate.
+Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
 ```{r}
@@
 ```
-
-```{r}
-#| label: fig-model-trumpmlr-res
-#| fig-cap: "Bayesian Hierarchical MLR Model of Trump Voting Percentage Captures Variability Without Overfitting"
-#| fig-subcap: ["Trump Predictions (Test Data)", "Residuals for Predictions"]
-#| echo: false
-#| warning: false
-#| layout-ncol: 2
-
-predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
-plot(test_data_trump$pct, colMeans(predictions2), 
-     xlab = "Actual Voter Percentage Values", 
-     ylab = "Predicted Voter Percentage Values")
-abline(0, 1, col = "red")
-
-residuals <- test_data$pct - colMeans(predictions)
-plot(colMeans(predictions), residuals, 
-     xlab = "Predicted Values", 
-     ylab = "Residuals")
-abline(h = 0, col = "red")
-```
-
-Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
-This suggests that model can generalize outside data.
-Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
-
-# Trump Model Results
-
-```{r}
-#| label: fig-trumpmodelci
-#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
-#| echo: false
-#| warning: false
-plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
-```
-
-Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% credible interval ranges from around 42.5% to 47.2%.
-This credible interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.
-
-```{r}
-#| label: fig-model-trumpmlr
-#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
-#| echo: false
-#| warning: false
-just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
-predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
-predicted_means = colMeans(predictions2)
-predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
-result_summary = data.frame(
-  Actual_PCT = just_trump$pct,
-  Predicted_PCT = predicted_means,
-  Lower_CI = predicted_intervals[1, ],
-  Upper_CI = predicted_intervals[2, ]
-)
-ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
-  geom_point(alpha = 0.7) +
-  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
-  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
-  labs(
-    title = "Actual vs Predicted Percentage For Trump MLR Model",
-    x = "Actual PCT",
-    y = "Predicted PCT"
-  )
-```
-
-```{r}
-#| label: tbl-model-trumpmlr
-#| tbl-cap: "Head of Bayesian Model Result Summary"
-#| echo: false
-#| warning: false
-
-kable(head(result_summary, 10), col.names = c("Acutal Harris Percentage", 
-                    "Predicted Percentage", 
-                    "Lower Bound Credible Interval",
-                    "Upper bound Credible Interval"))
-
-```
-
-From @fig-model-trumpmlr and @tbl-model-trumpmlr the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line).
-Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
-
-\newpage
-
-## Election Prediction {#sec-prediction}
-
-Our prediction process consists of two primary components.
-First, we develop models for both Trump and Harris based on the variables outlined in @sec-model.
-This involves partitioning the dataset into training and testing subsets.
-Next, we further divide the testing dataset into swing states and other competitive races.
-We then input this test data into the respective models to generate predictions.
-By averaging these predictions, we can calculate the expected voter percentage for each candidate in each state.
-The candidate with a higher percentage is deemed the winner for that state.
-
-We generated predictions for the following states: Arizona, Nevada, Georgia, Pennsylvania, Michigan, Minnesota, Wisconsin, Florida, Texas, Maine CD-2, Nebraska CD-2, New Hampshire, Ohio, Virginia, North Carolina, and Iowa.
-Winners for other states were determined based on historical trends and predictions from sources like @cnn_electoral_nodate.
-Most states without predictions are strongly Republican or Democratic, so their absence is not expected to significantly impact prediction validity.
 
@@
 This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys.
-Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias.
+Additionally, using multiple communication methods provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias.
 Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.
@@
 Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes.
-This trend may be worsened by decreased participation in polls as the election date approaches.
+This trend may be worsened by decreased participation in polls further from the election day.
 As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates.
@@
 Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies.
-This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.
+This potential undervaluation of Trump’s support could benefit the Harris campaign, by providing a warning to not repeat previous election mistakes.
 
@@
 
+Furthermore, the incorporation of both poll score and sample size into the model may lead to the potential for double counting the sample size.
+This concern was raised in Nate Silver's article [@silver_polls_2008].
+As illustrated in @fig-pollscore-samplesize, there is a positive correlation between poll score and sample size.
+However, this correlation is weak.
+Additionally, given that pollsters with negative poll scores are generally more reliable than those with positive scores, we would expect the opposite relationship to hold true for sample size to even have a possibility of being double counted.
+
 Moreover, errors and offsets inherent in polling data, such as response bias, nonresponse adjustments, and sampling variability, can propagate into the model’s results.
@@
 
-In @sec-prediction, the team predicted that Kamala Harris would secure 298 delegates while Donald Trump would obtain 240 delegates.
+In @sec-prediction, the team predicted that Kamala Harris would win 298 delegates while Donald Trump would get 240 delegates.
 However, this prediction has several limitations.
-Firstly, the model used to generate these predictions inherits the constraints outlined in @sec-model-limitations.
+First, the prediction will inherit all of the limitations of the model it (model limitations in @sec-model-limitations).
 Additionally, the predictions are based on averaging results from polling data for each state, which can lead to propagated prediction errors.
 
-Each state has a limited number of polls available; while the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states contributing only up to 27.
-Moreover, as noted in Section @sec-prediction, the analysis focused solely on contested races, specifically swing states.
-Consequently, many states were assigned to a candidate based on strong historical trends rather than direct predictions.
-Although this approach is reasonable, it may overlook emerging trends that could potentially shift a state's allegiance from Republican to Democrat or vice versa.
+Also, each state has a limited number of polls available.
+While the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states having only 27.
+Moreover, as noted in section @sec-prediction, the analysis focused solely on contested races, specifically swing states.
+Therefore, many states were assigned to a candidate based on strong historical trends and not from predictions.
+Although this approach is reasonable, it may overlook current trends that could potentially shift a state's vote from Republican to Democrat or vice versa.
 
@@
 
-The American federal or national government is split into three branches, executive., legislative, and judicial the executive branch includes the president and the military.
+The American federal or national government is split into three branches, executive, legislative, and judicial.
+The executive branch includes the president and the military.
 The legislative branch includes two subgroups, the House of Representatives, and the Senate.
@@
 Every state has two Senators and one Representative per approximately 750,000 people.
-The judicial branch is the court system.
+Finally, the judicial branch is the court system.
 
@@
 Instead of a direct popular vote, each state is allocated a certain number of electors based on its representation in Congress (the total of its Senators and Representatives).
-When voters vote, they are voting for a slate of electors pledged to a candidate.
+When voters vote, they are voting for what candidate they want their electors to vote for.
 The candidate who receives the most electoral votes (270 out of 538) wins the presidency.
-This system means that winning the popular vote in a state result in winning all that state's electoral votes.
+This system means that winning the popular vote in a state results in winning all that state's electoral votes.
 The only exception are the states of Maine and Nebraska, who award electoral votes by congressional district, with two additional votes given to the statewide winner.
 
-The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting pasterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts.
+The Electoral College results in some states being unnecessary to campaign in, as their strong historical voting patterns towards either Democrats or Republicans make them unlikely to change, regardless of campaign efforts.
 Therefore, for statisticians, polling information from these states may not be that useful when trying to predict the outcome of an election.
@@
 In Nebraska, the state overall is projected to lean Republican, with the first and third districts also strongly favoring Republican candidates, while the second district leans Democratic.
-In Maine, the overall expectation is a Democratic leaning, consistent with its first district, though the second district leans Republican.
+In Maine, the overall state is a Democratic leaning, and so is its first district, but the second district leans Republican.
 Furthermore, Alaska and Hawaii are not on the map.
@@
 
-The seven agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, and Minnesota as the next closest races.
+The seven agreed upon swing states are Pennsylvania, North Carolina, Georgia, Arizona, Nevada, Wisconsin, and Michigan with Texas, Florida, Nebraska District Two, Maine District 2, Iowa, and Minnesota as the next closest races.
 
@@
 
+# Trump Voter Prediction Model {#sec-trumpmodel}
+
+The multiple linear regression model (MLR) Bayesian hierarchical model for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris.
+Likewise, the Trump dataset is also split into training and testing data.
+The model results are similar to that of Harris's MLR model.
+
+```{r}
+#| echo: false
+#| warning: false
+set.seed(123)
+train_data_trump <- read_parquet(here("data/02-analysis_data/train_data_trump.parquet"))
+test_data_trump <- read_parquet(here("data/02-analysis_data/test_data_trump.parquet"))
+bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))
+```
+
+```{r}
+#| label: fig-model-trumpmlr-res
+#| fig-cap: "Bayesian Hierarchical MLR Model of Trump Voting Percentage Captures Variability Without Overfitting"
+#| fig-subcap: ["Trump Predictions (Test Data)", "Residuals for Predictions"]
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+
+predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
+plot(test_data_trump$pct, colMeans(predictions2), 
+     xlab = "Actual Voter Percentage Values", 
+     ylab = "Predicted Voter Percentage Values")
+abline(0, 1, col = "red")
+
+residuals <- test_data$pct - colMeans(predictions)
+plot(colMeans(predictions), residuals, 
+     xlab = "Predicted Values", 
+     ylab = "Residuals")
+abline(h = 0, col = "red")
+```
+
+Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values.
+This suggests that model has predictability strength.
+Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
+# Trump Model Results {#sec-trumpmodelresults}
+
+```{r}
+#| label: fig-trumpmodelci
+#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
+#| echo: false
+#| warning: false
+plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
+```
+
+Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% credible interval ranges from around 42.5% to 47.2%.
+This credible interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.
+
+```{r}
+#| label: fig-model-trumpmlr
+#| fig-cap: "Bayesian MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
+#| echo: false
+#| warning: false
+just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
+predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
+predicted_means = colMeans(predictions2)
+predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+result_summary = data.frame(
+  Actual_PCT = just_trump$pct,
+  Predicted_PCT = predicted_means,
+  Lower_CI = predicted_intervals[1, ],
+  Upper_CI = predicted_intervals[2, ]
+)
+ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
+  geom_point(alpha = 0.7) +
+  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "Actual vs Predicted Percentage For Trump MLR Model",
+    x = "Actual PCT",
+    y = "Predicted PCT"
+  )
+```
+
+```{r}
+#| label: tbl-model-trumpmlr
+#| tbl-cap: "Head of Bayesian Model Result Summary"
+#| echo: false
+#| warning: false
+
+kable(head(result_summary, 10), col.names = c("Acutal Harris Percentage", 
+                    "Predicted Percentage", 
+                    "Lower Bound Credible Interval",
+                    "Upper bound Credible Interval"))
+
+```
+
+From @fig-model-trumpmlr and @tbl-model-trumpmlr the model accounts for large amount of the variance in Trump's voter percentage as the data points appear close to the prediction (red line).
+Furthermore, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.
+
+\newpage
+
 # The Ideal Survey
@@
 Given the monetary constraint, the team's first decision is to mostly include swing states and districts.
-These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, and Georgia.
+These are Nevada, Arizona, Wisconsin, Michigan, Pennsylvania, North Carolina, Georgia, Nebraska District 2, and Maine District 2.
 Swing states are the primary focus of the poll because they disproportionately affect the outcome of the election based on the explanation in @sec-electoral-college.
-However, the poll will also include Texas, Florida, Minnesota, Nebraska District 2, and Maine District 2, as they have the potential to one by either party but are not as likely to change as the previously mentioned swing states.
-Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to Texas, Florida, and Minnesota.
+However, the poll will also include other close races like Texas, Florida, and Minnesota, as they have the potential to be won by either party but are not as likely to change as the previously mentioned swing states.
+Nevertheless, it is important to note that to save costs, the team will focus on polling true swing states compared to states like Texas, Florida, and Minnesota.
 
@@
 
-The team will then find and use multiple databases, such as @usps_customer_nodate, to find addresses, telephones, and emails of potential voters in those states.
+The team will then find and use multiple databases, such as U.S.
+Postal Service’s residential address file [@usps_customer_nodate], to find addresses, telephones, and emails of potential voters in those states.
 According to @pew_research_center_us_nodate, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums.
 Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email.
-Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.
+Additionally, to encourage survey completion, one in every 50 participants will have a chance at winning $20.
 
@@
 Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes.
-County proportion is used as the stratification variable because it is easily accessible.
+County proportion is used as the stratification variable because that information is easily accessible.
 Additionally, counties possess inherent demographic and cultural characteristics that can influence their alignment with particular political parties.
@@
 This sample size is large enough to provide reliable conclusions but not so large that it resembles sampling with replacement.
-For true swing states, the team aims to survey 1,000 individuals.
-However, if the final sample size falls short of this target, it is not considered a significant issue.
+For true swing states and other close races, the team aims to survey 1,000 individuals.
+However, for non-swing states, if the final sample size is less than 1,000, it is not a significant issue.
 
@@
 An example of this is question number 7.
+The idea to use a worthless question was inspired by @pew_research_center_us_nodate.
 
@@
 
-Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently.
+Answer bias, like order bias, occurs when the order of answer choices influences participants' selections.
+For instance, the first option can often be chosen more frequently.
 To prevent this, answer choices should be randomized upon survey entry.
@@
 While the team's methodology and survey create a robust system, there are potential issues.
-The reliance of weighting results based on a candidate's demographics can lead to error propagation.
-For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have impact on the polls prediction of what candidate will win the state.
+The usage of weighting results based on a candidate's demographics can lead to error propagation.
+For instance, if a certain racial demographic population is only captured limitedly and that limited sample is far from representative, a few individuals in the population can have major impact on the polls prediction of what candidate will win the state.
 There is also a selection bias in terms of the monetary reward.
-Potentially, people who like monetary rewards would be more likely to engage in the survey and could therefore exhibit certain voting or demographic characteristics that create an unrepresentative sample.
+Potentially, people who like monetary rewards would be more likely to engage in the survey and could therefore exhibit certain voting characteristics that create an unrepresentative sample.
 

### 2024-11-04 scripts/00-simulate_data.R
made cleaerer comments

@@ -54,7 +54,6 @@ swing_states <- c(
 
 
 
-# Generate Poll IDs and States
 unique_poll_ids <- sample(10000:99999, length(states))
 unique_question_ids <- sample(10000:999999, length(states))
 poll_ids <- rep(unique_poll_ids, each = 4)
@@ -67,7 +66,6 @@ sample_size <- rep(sample_size, each = 4)
 pollster_column <- sample(pollster, length(states), replace = TRUE)
 pollster_column <- rep(pollster_column, each = 4)
 
-# Function to generate a set of 4 random percentages that sum to 100%
 generate_percentages <- function(state) {
   jill_pct <- max(rnorm(1, 0.02, 0.01), 0.0)
   rfk_pct <- max(rnorm(1, 0.02, 0.01), 0.0)
@@ -95,7 +93,6 @@ generate_percentages <- function(state) {
 }
 
 
-# Create a dataset by randomly assigning states and parties to divisions
 analysis_data <- tibble(
   pollster = pollster_column,
   poll_id = poll_ids,

### 2024-11-04 paper/paper.qmd
fixed figure formatting
Update paper.qmd

@@
 title: "Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality"
-subtitle: "A Hybrid Model Predicts a Narrow Lead for Kamala Harris in Swing States as Election Day Approaches"
+subtitle: "A Hybrid Model Predicts a Lead for Kamala Harris in Swing States as Election Day Approaches"
 author: 
@@
 However, the polls in general show a slight lead for Harris throughout the last 100 days.
+
+\newpage
 

### 2024-11-04 scripts/03-clean_data.R
Update 03-clean_data.R

@@ -59,24 +59,7 @@ cleaned_data_state <- cleaned_data |>
 
 cleaned_data <- cleaned_data |> na.omit()
 
-
-# Delete NA columns
-# cleaned_data <- raw_data |>
-#   select(-sponsor_ids,
-#     -sponsors,
-#     -sponsor_candidate_id,
-#     -sponsor_candidate,
-#     -sponsor_candidate_party,
-#     -endorsed_candidate_id,
-#     -endorsed_candidate_name,
-#     -endorsed_candidate_party,
-#     -subpopulation, -tracking,
-#     -notes, -url_article, -url_topline, -url_crosstab,
-#     -source, -internal, -partisan, -seat_name,
-#     -ranked_choice_round) |> na.omit()
-
 #### Save data ####
 write_parquet(cleaned_data, "data/02-analysis_data/cleaned_data.parquet")
 write_parquet(cleaned_data_national, "data/02-analysis_data/cleaned_data_national.parquet")
 write_parquet(cleaned_data_state, "data/02-analysis_data/cleaned_data_state.parquet")
-write_parquet(morningconsult_data, "data/02-analysis_data/mc_data.parquet")

### 2024-11-04 paper/paper.qmd
fixed spelling mistakes

@@
 
-The team also built a Bayesian MLR using the same approach and formula for Trump. It is in section @sec-trumpmodel.
+The team also built a Bayesian MLR using the same approach and formula for Trump. It is in @sec-trumpmodel.
 
@@
 
-The team has provided results for a Trump Bayesian MLR model. It is in section @sec-trumpmodelresults.
+The team has provided results for a Trump Bayesian MLR model. It is in @sec-trumpmodelresults.
 
@@
 
-According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States.
+According to @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States.
 There are also a few states with predictions not visible on the map, we will describe those predictions below.
@@
 Based on the figures in @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores.
-From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability.
+From the analysis in @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability.
 This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys.
@@
 Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic.
-However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.
-
-Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls.
+However, we find the earlier reasons more compelling, as the data in @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.
+
+Finally, as illustrated in @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls.
 It is important to note that the FiveThirtyEight poll score system is partly based on historical accuracy.
@@
 While the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states having only 27.
-Moreover, as noted in section @sec-prediction, the analysis focused solely on contested races, specifically swing states.
+Moreover, as noted in @sec-prediction, the analysis focused solely on contested races, specifically swing states.
 Therefore, many states were assigned to a candidate based on strong historical trends and not from predictions.

### 2024-11-06 paper/paper.qmd
fix typo and update decimal

@@
 
-### States included in anaylsis
+### States included in Analysis
 
@@
 #| label: fig-avg-pct-state
-#| fig-cap: Historical State Voting Trends Are Maintined in 2024
+#| fig-cap: Historical State Voting Trends Are Maintained in 2024
 #| echo: false
@@
 #| label: fig-polldate-canidate-pct
-#| fig-cap: Both Canidate Percentages Are Gaining Support Closer to the Election
+#| fig-cap: Both Candidate Percentages Are Gaining Support Closer to the Election
 #| echo: false
@@
 
-Next, we set up a bayesian model as the following in accordance to @sec-modelsetup:
+Next, we set up a Bayesian model as the following in accordance to @sec-modelsetup:
 
@@
 #| label: fig-bayesian-model-actual-predict
-#| fig-cap: "The MLR Model Accruatley Predicts the Testing Dataset"
+#| fig-cap: "The MLR Model Accurately Predicts the Testing Dataset"
 #| echo: false
@@
 #| label: fig-bayesian-model-CI
-#| fig-cap: "Kamala Harris is Predicted to get Approximatley 52% of Popular Vote"
+#| fig-cap: "Kamala Harris is Predicted to get Approximately 52% of the Popular Vote"
 #| echo: false
@@
 
-predicted_means = colMeans(predictions2)
-predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+predicted_means = round(colMeans(predictions2), 2)
+predicted_intervals = round(apply(predictions2, 2, quantile, probs = c(0.025, 0.975)), 2)
 
@@
 
-  avg_predicted_pct_state_harris <- mean(predictions_state_harris)
-  avg_predicted_pct_state_trump <- mean(predictions_state_trump)
+  avg_predicted_pct_state_harris <- round(mean(predictions_state_harris), 2)
+  avg_predicted_pct_state_trump <- round(mean(predictions_state_trump), 2)
 
@@
 
-# Overview of American Electon System
+# Overview of American Election System
 
@@
 #| label: fig-trumpmodelci
-#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
+#| fig-cap: "Donald Trump Expected to Receive Approximately 45% of the Vote"
 #| echo: false
@@
 predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
-predicted_means = colMeans(predictions2)
-predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
+predicted_means = round(colMeans(predictions2), 2)
+predicted_intervals = round(apply(predictions2, 2, quantile, probs = c(0.025, 0.975)), 2)
 result_summary = data.frame(
@@
 
-kable(head(result_summary, 10), col.names = c("Acutal Harris Percentage", 
+kable(head(result_summary, 10), 
+      col.names = c("Actual Trump Percentage", 
                     "Predicted Percentage", 
                     "Lower Bound Credible Interval",
-                    "Upper bound Credible Interval"))
-
+                    "Upper Bound Credible Interval"))
 ```

### 2024-11-06 paper/paper.qmd
update date

@@
 thanks: "Code and data are available at: [https://github.com/yulexun/uselection](https://github.com/yulexun/uselection)."
-date: today
+date: November 4, 2024
 date-format: long
